### 2017-03-19 paper.tex
well-devleoped draft

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%2345678901234567890123456789012345678901234567890123456789012345678901234567890
%        1         2         3         4         5         6         7         8

\documentclass[letterpaper, 10 pt, conference]{ieeeconf}

\IEEEoverridecommandlockouts                              % This command is only
                                                          % needed if you want to
                                                          % use the \thanks command
\overrideIEEEmargins
% See the \addtolength command later in the file to balance the column lengths
% on the last page of the document

% The following packages can be found on http:\\www.ctan.org
%\usepackage{graphics} % for pdf, bitmapped graphics files
%\usepackage{epsfig} % for postscript graphics files
%\usepackage{mathptmx} % assumes new font selection scheme installed
%\usepackage{times} % assumes new font selection scheme installed
%\usepackage{amsmath} % assumes amsmath package installed
%\usepackage{amssymb}  % assumes amsmath package installed
\usepackage{xcolor}

\title{\LARGE \bf
A Formal Description of Bitcoin's Script Language
}

\author{ \parbox{4 in}{
                        \centering Domenic Cianfichi, Hugo Mailhot and Joseph McGee
                        Department of Computer Science\\
                        University of California, Davis\\
                        1 Shields Ave, Davis, CA 95616, USA
                      }
}

\begin{document}

\maketitle
\thispagestyle{empty}
\pagestyle{empty}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}

Bitcoin transactions are validated using a specialized language called Script. Despite the critical role it plays in the Bitcoin network, so far no formal description of the language has been developed to allow reasoning about the expressive power of the language. The recent years have shown that a lack of understanding of Script made it possible to create valid programs with undesirable behavior. Furthermore, the Bitcoin community acknowledges that standard transaction script formats, and maybe Script itself, might be changed in the future. In a first step to allow rigorous reasoning about Script's expressivity and limitations, we present a formal description of the language, along with example analyses that illustrate the usefulness of such a description.

\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{INTRODUCTION}

\textcolor{red}{\textit{To describe and motivate the problem or your survey topic }}\\

Script is what Bitcoin uses to encode transactions over the network. It is purposefully restricted, and most notably lacks a mechanism for looping and guarantees termination. To properly learn the language and understand its semantics, one must read through the catalog of \textit{opcodes} (hex values used as operators) and analyze its C++ reference implementation.

As a postfix language written in sequences of bytes, Script is not an easy language to pick up and understand. It is even harder to assess the effect that proposed modifications to the language might have on the guarantees that it must provide. In the past, Script has been progressively restricted in response to discovered vulnerabilities in the language. The restrictions eliminated the vulnerabilities, at the cost of reducing the expressiveness of the language. It is not clear that all of the restrictions added were necessary to produce the desired end of securing the language. A formal description and eventually specification would help in such determinations, allowing more objective and precise decisions.

As Bitcoin's economic presence grows, formal specifications, abstracted away from current current implementations of the Bitcoin protocol, will become increasingly essential in mediating conflict over what Bitcoin is or should be, allowing participants in the Bitcoin community to reason formally and precisely about their opinions and proposals.

\textcolor{blue}{[Describe history of Script and Bitcoin maybe, say the language was flawed]}

\textcolor{blue}{[Mention [3] when it says that "These limitations are encoded in a function called isStandard(), which defines five types of "standard" transactions. These limitations are temporary and might be lifted by the time you read this."
The transaction programs being subject to change justifies allowing formal reasoning about the language.]}

\section{APPROACH}

\textcolor{red}{\textit{Detailed description of the approach, and highlight the main technical difficulties and novelties. }}\\

\subsection{Definition of Script Grammar}

\subsubsection{Syntax}
\# Bitcoin wiki -> tables describe the format\\
\# transaction is X Y, inputs are such and such\\
\# looked at enums in source code to get full list of defined opcode and their hex values\\
JOSEPH: also explain how the different syntactic groups were broken down\\

\subsubsection{Operational semantics}

We described the big-step semantics and the small-step semantics of the language using a combination of the reference implementation of the language [Ref] and of the Script page of the Bitcoin wiki [Ref].

\subsubsection{Difficulties encountered}

  1) The source code of a script program is just a sequence of bytes and is not human-readable. To make documentation easier, the creators of the language defined an alias (\textit{opcode}) for each command. We first thought that the whole language could be described using only the opcodes, without referring to byte sequences as such, but the very semantics of certain commands made this impossible (see for example the \texttt{OP\_PUSHDATA} commands in Section 2 of the appendix). It took us a few tries to come up with a notation precise enough to describe source code as byte sequences, and yet elegant enough to allow concise expression of the semantics.

\subsection{Application of Script Grammar}

[Describe how grammar was used]

\section{APPLICATIONS AND RESULTS}

 \textcolor{red}{\textit{If appropriate, you should describe your implementation and experimental results. Explain how to interpret your numbers and results if applicable.}}

\textcolor{blue}{[Motivation for the application cases we chose.]}

\subsection{Programs written in the current version of Script halt within time linear in the length of the program}

For the Bitcoin network to function properly, it is important that the network form a single connected component. A type of attack against the network is to try and cut some nodes from the rest of the network, and then ``double-spend'' the same amount of money on the two disconnected components, which will both accept the transactions. Such a cut can be achieved by performing a DoS attack against a group of nodes that collectively ensure connectivity between two otherwise separated subnetworks. One way to perform a DoS attack could be to have a node attempt to evaluate a Script program with an unreasonable execution time. It is therefore important to guarantee that a Script program will terminate within a certain amount of steps.

Using the small-step semantics defined in Section 3 of the appendix, we see that no reduction rule yields its entire input, as was the case with the \texttt{while} local reduction rule in IMP. That is to say, no recursion is allowed. At most a proper subset of the input is conserved after applying the rule, when it is not completely replaced by \texttt{OP\_NOP}, Script's equivalent of the \texttt{skip} command. This guarantees that every small-step reduction progresses towards termination of the program. Furthermore, the global reduction rule trivially entails that for any program that is not only \texttt{OP\_NOP}, there exists a unique context reducible by a local reduction rule. Thus we have that every Script program halts.

We also know, from our description of Script's big-step semantics in Section 2 of the appendix, that every command requires a constant number of checks and operations on the stack.

So we have that a Script program will always halt, and that the number of operations required to evaluate it is at most linear in the amount of commands it contains.

\subsection{Previous versions of Script allowed undesirable programs}
Here we prove how previous states were problematic

\subsection{The Script command set could be reduced without loss of expressiveness}
Here we prove that some commands are redundant. Must exaplin why this can be desirable sometimes, but also why some of them are clearly uselessly redundant.

\section{RELATED WORK}

\textcolor{red}{\textit{Detailed discussion of related work. Should stress how these relate to your work. Simply list and describe what other people did is not sufficient.}}\\

\textcolor{blue}{COMPLETELY REVISE THIS}

The best documentation for Script exists on the Bitcoin wiki [2].  This documentation is written in plain English, and not in a logical formalism that allows mathematical reasoning. To our knowledge, no previous effort has been made to provide a formal specification of Script. If we turn to Forth to find something related, we can find standard specifications like the Forth-83 standard (http://forth.sourceforge.net/standard/fst83/fst83-12.htm). However even these specifications are also given in plain English.

\section{CONCLUSIONS}

\textcolor{red}{\textit{Again summarize the work and discuss possible future work.}}\\

\addtolength{\textheight}{-12cm}   % This command serves to balance the column lengths
                                  % on the last page of the document manually. It shortens
                                  % the textheight of the last page by a suitable amount.
                                  % This command does not take effect until the next page
                                  % so it should come on the page before the last. Make
                                  % sure that you do not shorten the textheight too much.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{thebibliography}{99}
\bibitem{c1} Bitcoin Core. (March 2017). \textit{Bitcoin Core} [Online; Accessed: 15-March-2017]. Available: https://github.com/bitcoin/bitcoin.

\bibitem{c2} Bitcoin Wiki. (March 2017). \textit{Script} [Online; Accessed: 15-March-2017]. Available: https://en.bitcoin.it/wiki/Script.

\bibitem{c3} Andreas M. Antonopoulos. (October 18 2016). \textit{Mastering Bitcoin - Unlocking digital currencies} [Online; Accessed: 15-March-2017]. Available: https://github.com/bitcoinbook/bitcoinbook.

\end{thebibliography}

\end{document}

### 2017-03-19 scriptGrammar.tex

\documentclass{article}
\usepackage{fullpage} % make the margins a reasonable size
\usepackage{amsmath}  % some handy math symbols
\usepackage{amssymb}  % more handy math symbols
\usepackage{amsthm}   % this is necessary to create theorems/lemmas/etc.
\usepackage{mathtools}
\usepackage[utf8]{inputenc}
\usepackage{mathpartir}
\usepackage{hyperref}
\usepackage{pdfpages}
\usepackage{multicol}
\usepackage{ulem}

\begin{document}

\IfFileExists{main.pdf}{\includepdf{main.pdf}}{}

\title{Script Grammar}
\date{}
\maketitle

\section{Intdoduction}

Script is a constrained stack-based language inspired by Forth.

Script programs (scripts) exist within Bitcoin transactions, which eventually are finalized or irrevocably committed to the global state of the Bitcoin network by their inclusion in a block which exists in the block-chain with the most proof of work expended.

Transactions consist of inputs and outputs, both of which include scripts. A transaction input script \textit{unlocks} bitcoins; a transaction output script \textit{locks} bitcoins. An input references the output of a previous transaction and provides a script which yields $\sigma_0$ when run. In order for the input to successfully unlock the bitcoins locked up in the referenced output, said output's script must yield $\sigma_1 = \sigma[V = valid]$ when run with it's state initialized to the state $\sigma_0$--yielded by the unlocking/input/spending script.

Script program syntax is defined in terms of byte-commands. Each command is one byte long. Not all bytes are defined as valid commands and many bytes which are defined have syntactic restrictions on the following program bytes which must hold in order for the program to be well-formed. Script defines 173 such commands, each corresponding to a particular byte, also called \textit{words} or \textit{opcodes}. For example, the command that adds 1 to the value on top of the stack, is written \texttt{0x8b}. These opcodes have human-readable aliases to make description of the language easier. For example, the opcode \texttt{0x8b} is referred to as \texttt{OP\_1ADD} in the reference implementation's source code and documentation.

Script is intentionally constrainted: There language does not allow for the expression of loops or recursion. This is desirable, guaranteeing that the runtime of a Script program is linear with the size of the program.

Script is defined by the specification provided by the reference implementation. 
Script is  as sequences of bytecodes. In describing its semantics, we will generally not refer to the format of the values, but some commands in the language only have meaning if we take into account the fact that any part of a Script program is a byte vector. For this reason, we define byte vectors as the following, and will reserve \texttt{b} and \texttt{B} to express a byte and a byte vector, respectively.

\hyperlink{OPSEM}{Section 2} presents a consistent and exhaustive BNF describing Script's syntax.

\hyperlink{OPSEM}{Section 3} specifies the operational semantics of the syntactic constructs defined in \hyperlink{OPSEM}{Section 2}.
\section{Syntax}

\subsection{Fundamental data types}
\noindent
\texttt{b ::= 0x00 | 0x01 | ... | 0xFF} \\
\texttt{B ::= b | b B} \\
\texttt{b$_i$ ::= b} \\
\texttt{B$^n$ ::= b$_1$ b$_2$ ... b$_n$} \\
\texttt{varint ::= }

\subsection{Transactions}
\noindent
\texttt{tx ::= tx\_version tx\_numin tx\_inputs tx\_numout tx\_outputs tx\_locktime  } \\
\texttt{tx\_version ::= B$^4$  } \\
\texttt{tx\_numin ::= varint  } \\
\texttt{tx\_inputs ::= txin | txin tx\_inputs   } \\
\texttt{tx\_numout ::= varint  } \\
\texttt{tx\_outputs ::= txout | txout tx\_outputs  } \\
\texttt{tx\_locktime ::= B$^4$  }

\subsubsection{Transaction Inputs}
\noindent
\texttt{txin ::= txin\_prevhash txin\_index txin\_scriptlen txin\_script txin\_seqno} \\
\texttt{txin\_prevhash ::= B$^{256}$} \\
\texttt{txin\_index ::= B$^4$} \\
\texttt{txin\_scriptlen ::= varint} \\
\texttt{txin\_script ::= script} \\
\texttt{txin\_seqno ::= B$^4$}

\subsubsection{Transaction Outputs}
\noindent
\texttt{txout ::= txout\_value txout\_scriptlen txout\_script  } \\
\texttt{txout\_value ::= B$^8$} \\
\texttt{txout\_scriptlen ::= varint} \\
\texttt{txout\_script ::= script}

\subsection{Script}
\noindent
\texttt{script ::= com | com script} \\
\texttt{com ::= scom | mcom}

\subsubsection{Single-Word Commands}
\noindent \hangindent=0.7cm
\texttt{scom ::=
  scom\_push | scom\_control | scom\_stack | scom\_splice | scom\_bitlogic |
  scom\_numeric | scom\_crypto | scom\_expansion | scom\_template} \\

\noindent \hangindent=0.7cm
\texttt{scom\_push ::=
  OP\_0 | OP\_FALSE | OP\_1NEGATE | \sout{OP\_RESERVED} | OP\_1 | OP\_TRUE |
  OP\_2 | OP\_3 | OP\_4 | OP\_5 | OP\_6 | OP\_7 | OP\_8 | OP\_9 | OP\_10 |
  OP\_11 | OP\_12 | OP\_13 | OP\_14 | OP\_15 | OP\_16} \\

\noindent \hangindent=0.7cm
\texttt{scom\_control ::=
 OP\_NOP | \sout{OP\_VER} | \sout{OP\_VERIF} | \sout{OP\_VERNOTIF} | OP\_VERIFY
 | OP\_RETURN} \\

\noindent \hangindent=0.7cm
\texttt{scom\_stack ::=
  OP\_TOALTSTACK | OP\_FROMALTSTACK | OP\_2DROP | OP\_2DUP | OP\_3DUP |
  OP\_2OVER | OP\_2ROT | OP\_2SWAP | OP\_IFDUP | OP\_DEPTH | OP\_DROP | OP\_DUP
  | OP\_NIP | OP\_OVER | OP\_PICK | OP\_ROLL | OP\_ROT | OP\_SWAP | OP\_TUCK}
  \\

\noindent \hangindent=0.7cm
\texttt{scom\_splice ::=
  \sout{OP\_CAT }| \sout{OP\_SUBSTR} | \sout{OP\_LEFT} | \sout{OP\_RIGHT} |
  OP\_SIZE} \\

\noindent \hangindent=0.7cm
\texttt{scom\_bitlogic ::=
  \sout{OP\_INVERT} | \sout{OP\_AND} | \sout{OP\_OR} | \sout{OP\_XOR} |
  OP\_EQUAL | OP\_EQUALVERIFY | \sout{OP\_RESERVED1} | \sout{OP\_RESERVED2}} \\

\noindent \hangindent=0.7cm
\texttt{scom\_numeric ::=
  OP\_1ADD | OP\_1SUB | \sout{OP\_2MUL} | \sout{OP\_2DIV} | OP\_NEGATE |
  OP\_ABS | OP\_NOT | OP\_0NOTEQUAL | OP\_ADD | OP\_SUB | \sout{OP\_MUL} |
  \sout{OP\_DIV} | \sout{OP\_MOD} | \sout{OP\_LSHIFT} | \sout{OP\_RSHIFT} |
  OP\_BOOLAND | OP\_BOOLOR | OP\_NUMEQUAL | OP\_NUMEQUALVERIFY |
  OP\_NUMNOTEQUAL | OP\_LESSTHAN | OP\_GREATERTHAN | OP\_LESSTHANOREQUAL |
  OP\_GREATERTHANOREQUAL | OP\_MIN | OP\_MAX | OP\_WITHIN} \\

\noindent \hangindent=0.7cm
\texttt{scom\_crypto ::=
  OP\_RIPEMD160 | OP\_SHA1 | OP\_SHA256 | OP\_HASH160 | OP\_HASH256 |
  OP\_CODESEPARATOR | OP\_CHECKSIG | OP\_CHECKSIGVERIFY | OP\_CHECKMULTISIG |
  OP\_CHECKMULTISIGVERIFY} \\

\noindent \hangindent=0.7cm
\texttt{scom\_expansion ::=
  \sout{OP\_NOP1} | OP\_CHECKLOCKTIMEVERIFY | OP\_NOP2 |
  OP\_CHECKSEQUENCEVERIFY | OP\_NOP3 | \sout{OP\_NOP4} | \sout{OP\_NOP5} |
  \sout{OP\_NOP6} | \sout{OP\_NOP7} | \sout{OP\_NOP8} | \sout{OP\_NOP9} |
  \sout{OP\_NOP10}} \\

\noindent \hangindent=0.7cm
\texttt{scom\_template ::=
  \sout{OP\_SMALLINTEGER} | \sout{OP\_PUBKEYS} | \sout{OP\_PUBKEYHASH} |
  \sout{OP\_PUBKEY} | \sout{OP\_INVALIDOPCODE}}

\subsubsection{Multiple-Word Commands}
\noindent \hangindent=0.7cm
\texttt{mcom ::= mcom\_if | mcom\_notif | mcom\_push} \\

% compund control commands
\noindent
\texttt{mcom\_if ::= OP\_IF script OP\_ENDIF | OP\_IF script mcom\_else OP\_ENDIF} \\
\texttt{mcom\_notif ::= OP\_NOTIF script OP\_ENDIF | OP\_NOTIF script mcom\_else OP\_ENDIF} \\
\texttt{mcom\_else ::= OP\_ELSE script | OP\_ELSE script mcom\_else } \\

% compound push commands
\noindent \hangindent=0.7cm
\texttt{mcom\_push ::=
  OP\_PUSHBYTES$_N$ B$^N$ | OP\_PUSHDATA1 B$^1$ B |OP\_PUSHDATA2 B$^2$ B
  | OP\_PUSHDATA4 B$^4$ B}

\subsubsection{Disabled Commands}
\noindent \hangindent=0.7cm
\texttt{dcom ::=
  dcom\_push | dcom\_control | dcom\_stack | dcom\_splice | dcom\_bitlogic |
  dcom\_numeric | dcom\_crypto | dcom\_expansion | dcom\_template} \\

\noindent \hangindent=0.7cm
\texttt{dcom\_push ::= \sout{OP\_RESERVED}} \\

\noindent \hangindent=0.7cm
\texttt{scom\_control ::=
 \sout{OP\_VER} | \sout{OP\_VERIF} | \sout{OP\_VERNOTIF}} \\

\noindent \hangindent=0.7cm
\texttt{dcom\_stack ::= } \\

\noindent \hangindent=0.7cm
\texttt{dcom\_splice ::=
  \sout{OP\_CAT }| \sout{OP\_SUBSTR} | \sout{OP\_LEFT} | \sout{OP\_RIGHT}
  } \\

\noindent \hangindent=0.7cm
\texttt{dcom\_bitlogic ::=
  \sout{OP\_INVERT} | \sout{OP\_AND} | \sout{OP\_OR} | \sout{OP\_XOR} |
  \sout{OP\_RESERVED1} | \sout{OP\_RESERVED2}} \\

\noindent \hangindent=0.7cm
\texttt{dcom\_numeric ::=
  \sout{OP\_2MUL} | \sout{OP\_2DIV} | \sout{OP\_MUL} |
  \sout{OP\_DIV} | \sout{OP\_MOD} | \sout{OP\_LSHIFT} | \sout{OP\_RSHIFT}} \\

\noindent \hangindent=0.7cm
\texttt{dcom\_crypto ::=} \\

\noindent \hangindent=0.7cm
\texttt{dcom\_expansion ::=
  \sout{OP\_NOP1} | \sout{OP\_NOP4} | \sout{OP\_NOP5} |
  \sout{OP\_NOP6} | \sout{OP\_NOP7} | \sout{OP\_NOP8} | \sout{OP\_NOP9} |
  \sout{OP\_NOP10}} \\

\noindent \hangindent=0.7cm
\texttt{dcom\_template ::=
  \sout{OP\_SMALLINTEGER} | \sout{OP\_PUBKEYS} | \sout{OP\_PUBKEYHASH} |
  \sout{OP\_PUBKEY} | \sout{OP\_INVALIDOPCODE}}

\pagebreak

\subsubsection{OP Codes}

\begin{multicols}{3}

\textbf{\centering push value} \\
\texttt{OP\_0 ::= 0x00  } \\
\texttt{OP\_FALSE ::= OP\_0  } \\
\texttt{{OP\_PUSHBYTES$^N$} ::= 0x01-0x4b} \\
\texttt{OP\_PUSHDATA1 ::= 0x4c  } \\
\texttt{OP\_PUSHDATA2 ::= 0x4d  } \\
\texttt{OP\_PUSHDATA4 ::= 0x4e  } \\
\texttt{OP\_1NEGATE ::= 0x4f  } \\
\texttt{OP\_RESERVED ::= 0x50  } \\
\texttt{OP\_1 ::= 0x51  } \\
\texttt{OP\_TRUE::=OP\_1  } \\
\texttt{OP\_2 ::= 0x52  } \\
\texttt{OP\_3 ::= 0x53  } \\
\texttt{OP\_4 ::= 0x54  } \\
\texttt{OP\_5 ::= 0x55  } \\
\texttt{OP\_6 ::= 0x56  } \\
\texttt{OP\_7 ::= 0x57  } \\
\texttt{OP\_8 ::= 0x58  } \\
\texttt{OP\_9 ::= 0x59  } \\
\texttt{OP\_10 ::= 0x5a  } \\
\texttt{OP\_11 ::= 0x5b  } \\
\texttt{OP\_12 ::= 0x5c  } \\
\texttt{OP\_13 ::= 0x5d  } \\
\texttt{OP\_14 ::= 0x5e  } \\
\texttt{OP\_15 ::= 0x5f  } \\
\texttt{OP\_16 ::= 0x60  } \\

\textbf{\centering control} \\
\texttt{OP\_NOP ::= 0x61  } \\
\texttt{OP\_VER ::= 0x62  } \\
\texttt{OP\_IF ::= 0x63  } \\
\texttt{OP\_NOTIF ::= 0x64  } \\
\texttt{OP\_VERIF ::= 0x65  } \\
\texttt{OP\_VERNOTIF ::= 0x66  } \\
\texttt{OP\_ELSE ::= 0x67  } \\
\texttt{OP\_ENDIF ::= 0x68  } \\
\texttt{OP\_VERIFY ::= 0x69  } \\
\texttt{OP\_RETURN ::= 0x6a  } \\

\textbf{\centering stack ops} \\
\texttt{OP\_TOALTSTACK ::= 0x6b  } \\
\texttt{OP\_FROMALTSTACK ::= 0x6c  } \\
\texttt{OP\_2DROP ::= 0x6d  } \\
\texttt{OP\_2DUP ::= 0x6e  } \\
\texttt{OP\_3DUP ::= 0x6f  } \\
\texttt{OP\_2OVER ::= 0x70  } \\
\texttt{OP\_2ROT ::= 0x71  } \\
\texttt{OP\_2SWAP ::= 0x72  } \\
\texttt{OP\_IFDUP ::= 0x73  } \\
\texttt{OP\_DEPTH ::= 0x74  } \\
\texttt{OP\_DROP ::= 0x75  } \\
\texttt{OP\_DUP ::= 0x76  } \\
\texttt{OP\_NIP ::= 0x77  } \\
\texttt{OP\_OVER ::= 0x78  } \\
\texttt{OP\_PICK ::= 0x79  } \\
\texttt{OP\_ROLL ::= 0x7a  } \\
\texttt{OP\_ROT ::= 0x7b  } \\
\texttt{OP\_SWAP ::= 0x7c  } \\
\texttt{OP\_TUCK ::= 0x7d  } \\

\textbf{\centering splice ops} \\
\texttt{OP\_CAT ::= 0x7e  } \\
\texttt{OP\_SUBSTR ::= 0x7f  } \\
\texttt{OP\_LEFT ::= 0x80  } \\
\texttt{OP\_RIGHT ::= 0x81  } \\
\texttt{OP\_SIZE ::= 0x82  } \\

\textbf{\centering bit logic} \\
\texttt{OP\_INVERT ::= 0x83  } \\
\texttt{OP\_AND ::= 0x84  } \\
\texttt{OP\_OR ::= 0x85  } \\
\texttt{OP\_XOR ::= 0x86  } \\
\texttt{OP\_EQUAL ::= 0x87  } \\
\texttt{OP\_EQUALVERIFY ::= 0x88  } \\
\texttt{OP\_RESERVED1 ::= 0x89  } \\
\texttt{OP\_RESERVED2 ::= 0x8a  } \\

\textbf{\centering numeric} \\
\texttt{OP\_1ADD ::= 0x8b  } \\
\texttt{OP\_1SUB ::= 0x8c  } \\
\texttt{OP\_2MUL ::= 0x8d  } \\
\texttt{OP\_2DIV ::= 0x8e  } \\
\texttt{OP\_NEGATE ::= 0x8f  } \\
\texttt{OP\_ABS ::= 0x90  } \\
\texttt{OP\_NOT ::= 0x91  } \\
\texttt{OP\_0NOTEQUAL ::= 0x92  } \\
\texttt{OP\_ADD ::= 0x93  } \\
\texttt{OP\_SUB ::= 0x94  } \\
\texttt{OP\_MUL ::= 0x95  } \\
\texttt{OP\_DIV ::= 0x96  } \\
\texttt{OP\_MOD ::= 0x97  } \\
\texttt{OP\_LSHIFT ::= 0x98  } \\
\texttt{OP\_RSHIFT ::= 0x99  } \\
\texttt{OP\_BOOLAND ::= 0x9a  } \\
\texttt{OP\_BOOLOR ::= 0x9b  } \\
\texttt{OP\_NUMEQUAL ::= 0x9c  } \\
\texttt{OP\_NUMEQUALVERIFY ::= 0x9d  } \\
\texttt{OP\_NUMNOTEQUAL ::= 0x9e  } \\
\texttt{OP\_LESSTHAN ::= 0x9f  } \\
\texttt{OP\_GREATERTHAN ::= 0xa0  } \\
\texttt{OP\_LESSTHANOREQUAL ::= 0xa1  } \\
\texttt{OP\_GREATERTHANOREQUAL ::= 0xa2  } \\
\texttt{OP\_MIN ::= 0xa3  } \\
\texttt{OP\_MAX ::= 0xa4  } \\
\texttt{OP\_WITHIN ::= 0xa5  } \\

\textbf{\centering crypto} \\
\texttt{OP\_RIPEMD160 ::= 0xa6  } \\
\texttt{OP\_SHA1 ::= 0xa7  } \\
\texttt{OP\_SHA256 ::= 0xa8  } \\
\texttt{OP\_HASH160 ::= 0xa9  } \\
\texttt{OP\_HASH256 ::= 0xaa  } \\
\texttt{OP\_CODESEPARATOR ::= 0xab  } \\
\texttt{OP\_CHECKSIG ::= 0xac  } \\
\texttt{OP\_CHECKSIGVERIFY ::= 0xad  } \\
\texttt{OP\_CHECKMULTISIG ::= 0xae  } \\
\texttt{OP\_CHECKMULTISIGVERIFY ::= 0xaf  } \\

\textbf{\centering expansion} \\
\texttt{OP\_NOP1 ::= 0xb0  } \\
\texttt{OP\_CHECKLOCKTIMEVERIFY ::= OP\_NOP2 ::= 0xb1  } \\
\texttt{OP\_CHECKSEQUENCEVERIFY ::= OP\_NOP3 ::= 0xb2  } \\
\texttt{OP\_NOP4 ::= 0xb3  } \\
\texttt{OP\_NOP5 ::= 0xb4  } \\
\texttt{OP\_NOP6 ::= 0xb5  } \\
\texttt{OP\_NOP7 ::= 0xb6  } \\
\texttt{OP\_NOP8 ::= 0xb7  } \\
\texttt{OP\_NOP9 ::= 0xb8  } \\
\texttt{OP\_NOP10 ::= 0xb9  } \\

\textbf{\centering template matching params }\\
\texttt{OP\_SMALLINTEGER ::= 0xfa  } \\
\texttt{OP\_PUBKEYS ::= 0xfb  } \\
\texttt{OP\_PUBKEYHASH ::= 0xfd  } \\
\texttt{OP\_PUBKEY ::= 0xfe  } \\
\texttt{OP\_INVALIDOPCODE ::= 0xff  } \\

\end{multicols}

\pagebreak[2]

\hypertarget{OPSEM}{\section{Operational Semantics -- Big Step Semantics}}

At any given point of a program's execution, the state is entirely described by the following elements:

\begin{itemize}
	\item The stack $S$, whose elements $S_i$ for $1 \leq i \leq |S|$ are indexed starting from the bottom;
    \item The alt-stack $AS$, an auxiliary stack indexed in the same way;
    \item The validity variable $V$, equal to \texttt{true} by default.
\end{itemize}

Any state in which $V = \texttt{false}$ causes the program to immediately terminate with a special exception, handled by the environment in which the program was being run.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Utility rules
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\hypertarget{OPGEN}{\subsection{Generic rules}}
In order to more succinctly define the operational semantics of Script, we define generic rules not tied to any command in the language that express operations common in a stack-based language.

% PUSH
\subsubsection{PUSH}
\inferrule
{
	|S| = L
}
{	
	<\texttt{PUSH } B, \sigma> \Downarrow \sigma[|S| = L+1, S_{|S|} = B]
}
\vspace{3mm}

% DROP
\subsubsection{DROP}
\inferrule
{
	\sigma(|S|) = L \hspace{3mm} 
    B \Downarrow n \hspace{3mm} 
	L \geq n
}
{
	<\texttt{DROP } B, \sigma> \Downarrow \sigma[S_i = S_{i+1} \hspace{3mm} \forall i. \hspace{1mm} n \leq i < L]
}
\vspace{3mm}

% PUSH_ALT
\subsubsection{PUSH\_ALT}
\inferrule
{|AS| = L}
{<\texttt{PUSH\_ALT } B, \sigma> \Downarrow \sigma[|AS| = L+1, AS_{|AS|} = B]}
\vspace{3mm}

% DROP_ALT
\subsubsection{DROP\_ALT}
\inferrule
{
	\sigma(|AS|) = L \hspace{3mm} \sigma(|AS|) \geq B
}
{
	<\texttt{DROP\_ALT } B, \sigma> \Downarrow \sigma[AS_i = AS_{i+1} \hspace{3mm} \forall i. \hspace{1mm} B \leq i < L]
}
\vspace{3mm}

% TEST
\subsubsection{TEST}
\inferrule
{	
	\sigma(|S|) \geq x \hspace{3mm}
	\sigma(S_x) \Downarrow t \hspace{3mm} 
    t = 0
}
{
	<\texttt{TEST $x$}, \sigma> \Downarrow FALSE
}
\vspace{3mm}

\inferrule
{	
	\sigma(|S|) \geq x \hspace{3mm}
	\sigma(S_x) \Downarrow t \hspace{3mm} 
    t \neq 0
}
{
	<\texttt{TEST $x$}, \sigma> \Downarrow TRUE
}
\vspace{3mm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Constants}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Constants: OP Code Definitions

% OP_0, OP_FALSE
\subsubsection{OP\_0, OP\_FALSE}
Push the byte-vector representing 0 onto the stack. \\

\inferrule
{
	\sigma(|S|) = L \hspace{3mm} 
    B \Downarrow 0
}
{
	<\texttt{OP\_0}, \sigma> \Downarrow \sigma[|S|=L+1, S_{L+1}=B]
}

% OP_N
\subsubsection{OP\_N, \textit{OP\_PUSHNBYTES}}
Push the next $N$ bytes onto the stack. \\

\inferrule
{
	\sigma(|S|) = L
}
{
	<\texttt{OP\_N $b_1 \ldots b_N$}, \sigma> \Downarrow 
	\sigma[|S|=L+1, S_{L+1}=<b_1 \ldots b_N>]
}

% OP_PUSHDATA1
\subsubsection{OP\_PUSHDATA1}
The next byte specifies how many bytes to push onto the stack as a byte-vector. \\

\inferrule
{
    B_1 \Downarrow k \hspace{3mm} 
	\sigma(|S|) = L \hspace{3mm} 
}
{
	<\texttt{OP\_PUSHDATA1 $B_1^1$ $B_2^k$}, \sigma> \Downarrow 
	\sigma[|S|=L+1, S_{L+1}=B_2]
}

%\vspace{3mm}
%\inferrule
%{
%	x = |$B$| \hspace{3mm}
%    b_1 \Downarrow k \hspace{3mm} 
%    x < k 
%}
%{	
%	<\texttt{OP\_PUSHDATA1 $b_1$ $B$ }, \sigma> \Downarrow 
%	\sigma[V = false]
%}

% OP_PUSHDATA2
\subsubsection{OP\_PUSHDATA2}
The next two bytes specify how many bytes to push onto the stack as a byte-vector. \\

\inferrule
{
    B_1 \Downarrow k \hspace{3mm} 
	\sigma(|S|) = L \hspace{3mm} 
}
{
	<\texttt{OP\_PUSHDATA2 $B_1^2$ $B_2^k$}, \sigma> \Downarrow 
	\sigma[|S|=L+1, S_{L+1}=B_2]
}

%\vspace{3mm}
%\inferrule
%{
%	x = |$B$| \hspace{3mm}
%    (b_1 b_2) \Downarrow k \hspace{3mm} 
%    x < k 
%}
%{	
%	<\texttt{OP\_PUSHDATA2 $b_1$ $b_2$ $B$ }, \sigma> \Downarrow 
%	\sigma[V = false]
%}

% OP_PUSHDATA4
\subsubsection{OP\_PUSHDATA4}
The next four bytes specify how many bytes to push onto the stack as a byte-vector. \\

\inferrule
{
    B_1 \Downarrow k \hspace{3mm} 
	\sigma(|S|) = L \hspace{3mm} 
}
{
	<\texttt{OP\_PUSHDATA4 $B_1^4$ $B_2^k$}, \sigma> \Downarrow
	\sigma[|S|=L+1, S_{L+1}=B_2]
}

%\vspace{3mm}
%\inferrule
%{
%	x = |$B$| \hspace{3mm}
%    (b_1 b_2 b_3 b_4) \Downarrow k \hspace{3mm} 
%    x < k 
%}
%{	
%	<\texttt{OP\_PUSHDATA4 $b_1$ $b_2$ $b_3$ $b_4$ $B$ }, \sigma> \Downarrow 
%	\sigma[V = false]
%}

% OP_1NEGATE
\subsubsection{OP\_1NEGATE}
Push the byte-vector representing -1 onto the stack.\\

\inferrule
{
	\sigma(|S|) = L \hspace{3mm} 
    B \Downarrow -1
}
{<\texttt{OP\_0}, \sigma> \Downarrow 
	\sigma[|S|=l+1, S_{l+1}=B]}

% OP_1, OP_TRUE
\subsubsection{OP\_1, OP\_TRUE}
Push the byte-vector representing 1 onto the stack. \\
\inferrule
{\sigma(|S|) = L \hspace{3mm} B \Downarrow 1}
{<\texttt{OP\_0}, \sigma> \Downarrow 
	\sigma[|S|=L+1, S_{L+1}=B]}

% OP_2-OP_16
\subsubsection{OP\_2-OP\_16, \textit{OP\_PUSHN}}
Push the byte-vector representing the number specified in the word name onto the stack. \\

\inferrule
{\sigma(|S|) = L \hspace{3mm} B \Downarrow N}
{<\texttt{OP\_PUSHN}, \sigma> \Downarrow 
	\sigma[|S|=L+1, S_{L+1}=B]}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Flow Control}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Flow Control: OP Code Definitions

% OP_NOP
\subsubsection{OP\_NOP}
Do nothing. \\

\inferrule
{\quad}
{
	<\texttt{OP\_NOP},\sigma> \Downarrow \sigma
}

% OP_IF, OP_ELSE, OP_ENDIF
\subsubsection{OP\_IF, OP\_ELSE, OP\_ENDIF}
Run $C_1$ if top of stack is present and evaluates to true. Otherwise run $C_2$ if stack is present and evaluates to false. \\

\inferrule
{	
	\sigma(|S|)=L \hspace{3mm}
	<\texttt{TEST $L$},\sigma> \Downarrow \texttt{TRUE} \hspace{3mm}
	<\texttt{DROP $L$},\sigma> \Downarrow \sigma_1 \hspace{3mm}
    <C_1, \sigma_1> \Downarrow \sigma '
}
{	
	<\texttt{OP\_IF $C_1$ OP\_ENDIF}> \Downarrow \sigma '
}
\vspace{3mm}

\inferrule
{	
	\sigma(|S|)=L \hspace{3mm}
	<\texttt{TEST $L$},\sigma> \Downarrow \texttt{FALSE} \hspace{3mm}
	<\texttt{DROP $L$},\sigma> \Downarrow \sigma ' \hspace{3mm}
}
{	
	<\texttt{OP\_IF $C_1$ OP\_ENDIF}> \Downarrow \sigma '
}
\vspace{3mm}

\inferrule
{	\sigma(|S|)=L \hspace{3mm}
	<\texttt{TEST $L$},\sigma> \Downarrow \texttt{TRUE} \hspace{3mm}
	<\texttt{DROP $L$},\sigma> \Downarrow \sigma_1 \hspace{3mm}
    <C_1, \sigma_1> \Downarrow \sigma '
}
{	<\texttt{OP\_IF $C_1$ OP\_ELSE $C_2$ OP\_ENDIF}> \Downarrow \sigma '
}
\vspace{3mm}

\inferrule
{	\sigma(|S|)=L \hspace{3mm}
	<\texttt{TEST $L$},\sigma> \Downarrow \texttt{FALSE} \hspace{3mm}
	<\texttt{DROP $L$},\sigma> \Downarrow \sigma_1 \hspace{3mm}
    <C_2, \sigma_1> \Downarrow \sigma '
}
{	<\texttt{OP\_IF $C_1$ OP\_ELSE $C_2$ OP\_ENDIF}> \Downarrow \sigma '
}
\vspace{3mm}

% OP_NOTIF, OP_ELSE, OP_ENDIF
\subsubsection{OP\_NOTIF, OP\_ELSE, OP\_ENDIF}
Run $C_1$ if top of stack is present and evaluates to false. Otherwise run $C_2$ if stack is present and evaluates to true. \\

\inferrule
{	\sigma(|S|)=L \hspace{3mm}
	<\texttt{TEST $L$},\sigma> \Downarrow \texttt{FALSE} \hspace{3mm}
	<\texttt{DROP $L$},\sigma> \Downarrow \sigma_1 \hspace{3mm}
    <C_1, \sigma_1> \Downarrow \sigma '
}
{	<\texttt{OP\_NOTIF $C_1$ OP\_ENDIF}> \Downarrow \sigma '
}
\vspace{3mm}

\inferrule
{	\sigma(|S|)=L \hspace{3mm}
	<\texttt{TEST $L$},\sigma> \Downarrow \texttt{TRUE} \hspace{3mm}
	<\texttt{DROP $L$},\sigma> \Downarrow \sigma ' \hspace{3mm}
}
{	<\texttt{OP\_NOTIF $C_1$ OP\_ENDIF}> \Downarrow \sigma '
}
\vspace{3mm}

\inferrule
{	\sigma(|S|)=L \hspace{3mm}
	<\texttt{TEST $L$},\sigma> \Downarrow \texttt{FALSE} \hspace{3mm}
	<\texttt{DROP $L$},\sigma> \Downarrow \sigma_1 \hspace{3mm}
    <C_1, \sigma_1> \Downarrow \sigma '
}
{	<\texttt{OP\_NOTIF $C_1$ OP\_ELSE $C_2$ OP\_ENDIF}> \Downarrow \sigma '
}
\vspace{3mm}

\inferrule
{	\sigma(|S|)=L \hspace{3mm}
	<\texttt{TEST $L$},\sigma> \Downarrow \texttt{TRUE} \hspace{3mm}
	<\texttt{DROP $L$},\sigma> \Downarrow \sigma_1 \hspace{3mm}
    <C_2, \sigma_1> \Downarrow \sigma '
}
{	<\texttt{OP\_NOTIF $C_1$ OP\_ELSE $C_2$ OP\_ENDIF}> \Downarrow \sigma '
}
\vspace{3mm}

% OP_VERIFY
\subsubsection{OP\_VERIFY := 0x69}
If top of stack is present and evaluates to true then remove top of stock and mark transaction as valid; otherwise mark transaction as invalid. \\

\inferrule
{	\sigma(|S|)=L \hspace{3mm}
	<\texttt{TEST $L$},\sigma> \Downarrow \texttt{TRUE} \hspace{3mm}
	<\texttt{DROP $L$},\sigma> \Downarrow \sigma ' \hspace{3mm}
    \sigma'(V)=\texttt{invalid}
}
{	<\texttt{OP\_VERIFY}, \sigma> \Downarrow \sigma '
}

% OP_RETURN
\subsubsection{OP\_RETURN := 0x6a}
Unconditionally mark transaction as invalid. \\

\inferrule
{	\quad
}
{	<\texttt{OP\_RETURN}, \sigma> \Downarrow \sigma[V=\texttt{invalid}]
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Stack}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% OP_TOALTSTACK
\subsubsection{OP\_TOALTSTACK}

\inferrule
{	
	|S| = L \hspace{3mm}
    S_L = x \hspace{3mm}
    <\texttt{DROP } L, \sigma> \Downarrow \sigma_1
    <\texttt{PUSHALT } x, \sigma_1> \Downarrow \sigma'
}
{
	<\texttt{OP\_TOALTSTACK}, \sigma > \Downarrow \sigma
}
\vspace{3mm}

% OP_FROMALTSTACK
\subsubsection{OP\_FROMALTSTACK}

\inferrule
{	
	|AS| = L \hspace{3mm}
    AS_L = x \hspace{3mm}
    <\texttt{DROPALT } L, \sigma> \Downarrow \sigma_1
    <\texttt{PUSH } x, \sigma_1> \Downarrow \sigma'
}
{
	<\texttt{OP\_FROMALTSTACK}, \sigma > \Downarrow \sigma
}
\vspace{3mm}

%  OP_IFDUP
\subsubsection{OP\_IFDUP}
If top of stack is present and non-zero, duplicate the top of the stack. \\

\inferrule
{
	\sigma(S_{|S|}) = 0
}
{	
	<\texttt{OP\_IFDUP}, \sigma > \Downarrow \sigma
}
\vspace{3mm}

\inferrule
{
	\sigma(S_{|S|}) \neq 0 \hspace{3mm}
    <\texttt{PUSH } S_{|S|} \Downarrow \sigma ' >
} 
{
	<\texttt{OP\_IFDUP}, \sigma > \Downarrow \sigma'
}
\vspace{3mm}

% OP_DEPTH
\subsubsection{OP\_DEPTH}
Push the byte-vector representing the depth of the stack onto the stack. \\

\inferrule
{
	\sigma(|S|) = L \hspace{3mm} 
    < \texttt{PUSH } L, \sigma > \Downarrow \sigma '
}
{
	<\texttt{OP\_DEPTH}, \sigma > \Downarrow \sigma '
}
\vspace{3mm}

% OP_DROP
\subsubsection{OP\_DROP}
Remove the top of the stack. \\

\inferrule
{
	\sigma(|S|) = L \hspace{3mm} 
    L > 0 \hspace{3mm}
    <\texttt{DROP } L, \sigma > \Downarrow \sigma'
}
{	
	<\texttt{OP\_DROP}, \sigma > \Downarrow \sigma'
}
\vspace{3mm}

% OP_DUP
\subsubsection{OP\_DUP}
Duplicate the top of the stack. \\

\inferrule
{
	\sigma(|S|) = L \hspace{3mm} 
    L > 0 \hspace{3mm}
    <\texttt{PUSH }S_L, \sigma > \Downarrow \sigma'
}
{
	<\texttt{OP\_DUP}, \sigma > \Downarrow \sigma'
}
\vspace{3mm}

% OP_NIP
\subsubsection{OP\_NIP}
Remove the byte-vector second from the top of the stack. \\

\inferrule
{
	\sigma(|S|) = L \hspace{3mm} 
    L > 1 \hspace{3mm} 
    <\texttt{DROP } L-1, \sigma > \Downarrow \sigma'
}
{
	<\texttt{OP\_NIP}, \sigma > \Downarrow \sigma'
}
\vspace{3mm}

% OP_OVER
\subsubsection{OP\_OVER}
Copy the byte-vector second from the top onto the top of the stack. \\

\inferrule
{
	\sigma(|S|) = L \hspace{3mm} 
    L > 1 \hspace{3mm} 
    <\texttt{PUSH } S_{L-1}, \sigma > \Downarrow \sigma'
}
{
	<\texttt{OP\_OVER}, \sigma > \Downarrow \sigma'
}
\vspace{3mm}

% OP_PICK
\subsubsection{OP\_PICK}
Copy the byte-vector $n$ from the top onto the top of the stack, not counting the top element $n$. \\

\inferrule
{
	\sigma(|S|) = L \hspace{3mm}
    \sigma(S_L) = n \hspace{3mm}
    <\texttt{DROP } L, \sigma> \Downarrow \sigma_1 \hspace{3mm}  
    <\texttt{PUSH } S_{L-n-1}, \sigma_1> \Downarrow \sigma'
}
{
	<\texttt{OP\_PICK}, \sigma > \Downarrow \sigma'
}
\vspace{3mm}

% OP_ROLL
\subsubsection{OP\_ROLL}
Move the byte-vector $n$ from the top onto the top of the stack, not counting the top element $n$. \\

\inferrule
{
	\sigma(|S|) = L \hspace{3mm}
    \sigma(S_L) = n \hspace{3mm}
    <\texttt{DROP } L, \sigma> \Downarrow \sigma_1 \hspace{3mm}  
    <\texttt{PUSH } S_{L-n-1}, \sigma_1> \Downarrow \sigma_2
    <\texttt{DROP } L-n-1, \sigma_2> \Downarrow \sigma'
}
{
	<\texttt{OP\_ROLL}, \sigma > \Downarrow \sigma'
}
\vspace{3mm}

% OP_ROT
\subsubsection{OP\_ROT}
The top three items on the stack are rotated. Equivalently, move the byte-vector third from the top to the top of the stack. \\

\inferrule
{
	\sigma(|S|) = L \hspace{3mm}
    <\texttt{PUSH } S_{L-2}, \sigma_1> \Downarrow \sigma_2
    <\texttt{DROP } L-2, \sigma_2> \Downarrow \sigma'
}
{
	<\texttt{OP\_ROT}, \sigma > \Downarrow \sigma'
}
\vspace{3mm}

% OP_SWAP
\subsubsection{OP\_SWAP}
The top two items on the stack are swapped. \\

\inferrule
{
	\sigma{|S|} = L \hspace{3mm}
    <\texttt{PUSH } S_{L-1}, \sigma_1> \Downarrow \sigma_2
    <\texttt{DROP } {L-1}, \sigma_2> \Downarrow \sigma'
}
{<\texttt{OP\_SWAP}, \sigma > \Downarrow \sigma'}
\vspace{3mm}

% OP_TUCK
\subsubsection{OP\_TUCK}
The item at the top of the stack is copied and inserted before the second-to-top item. \\

\inferrule
{	
	\sigma(S_{|S|}) = x \hspace{3mm} 
	\sigma(S_{|S|-1}) = y \hspace{3mm} 
	<\texttt{DROP } |S|, \sigma> \Downarrow \sigma_1 \hspace{3mm}
	<\texttt{DROP } |S|, \sigma_1> \Downarrow \sigma_2 \hspace{3mm} \\
    <\texttt{PUSH } x, \sigma_2> \Downarrow \sigma_3 \hspace{3mm}
	<\texttt{PUSH } y, \sigma_3> \Downarrow \sigma_4 \hspace{3mm} 
    <\texttt{PUSH } x, \sigma_4> \Downarrow \sigma'
}
{	
	<\texttt{OP\_TUCK}, \sigma > \Downarrow \sigma'
}
\vspace{3mm}

% OP_2DROP
\subsubsection{OP\_2DROP}
Removes the top two stack items. \\

\inferrule
{	
	<\texttt{DROP } |S|, \sigma> \Downarrow \sigma_1 \hspace{3mm}
	<\texttt{DROP } |S|, \sigma_1> \Downarrow \sigma' \hspace{3mm} \\
}
{	
	<\texttt{OP\_2DROP}, \sigma > \Downarrow \sigma'
}
\vspace{3mm}

% OP_2DUP
\subsubsection{OP\_2DUP}
Duplicates the top two stack items. \\

\inferrule
{	
	\sigma(|S|) = L \hspace{3mm}
	<\texttt{PUSH } S_{L-1}, \sigma> \Downarrow \sigma_1 \hspace{3mm}
    <\texttt{PUSH } S_L, \sigma_1> \Downarrow \sigma' \hspace{3mm}
}
{	
	<\texttt{OP\_2DUP}, \sigma > \Downarrow \sigma'
}
\vspace{3mm}

% OP_3DUP
\subsubsection{OP\_3DUP}
Duplicates the top three stack items. \\

\inferrule
{	
	\sigma(|S|) = L \hspace{3mm}
	<\texttt{PUSH } S_{L-2}, \sigma> \Downarrow \sigma_1 \hspace{3mm}
    <\texttt{PUSH } S_{L-1}, \sigma_1> \Downarrow \sigma_2 \hspace{3mm}
    <\texttt{PUSH } S_L, \sigma_2> \Downarrow \sigma' \hspace{3mm}
}
{	
	<\texttt{OP\_3DUP}, \sigma > \Downarrow \sigma'
}
\vspace{3mm}

% OP_2OVER
\subsubsection{OP\_2OVER}
Copies the pair of items two spaces back in the stack to the front. \\

\inferrule
{	
	\sigma(|S|) = L \hspace{3mm}
	<\texttt{PUSH } S_{L-3}, \sigma> \Downarrow \sigma_1 \hspace{3mm}
    <\texttt{PUSH } S_{L-2}, \sigma_1> \Downarrow \sigma' \hspace{3mm}
}
{	
	<\texttt{OP\_2OVER}, \sigma > \Downarrow \sigma'
}
\vspace{3mm}

% OP_2ROT
\subsubsection{OP\_2ROT}
The fifth and sixth items back are moved to the top of the stack. \\

\inferrule
{	
	\sigma(|S|) = L \hspace{3mm}
    <\texttt{PUSH } S_{L-5}, \sigma_2> \Downarrow \sigma_3 \hspace{3mm}
    <\texttt{PUSH } S_{L-4}, \sigma_3> \Downarrow \sigma' \hspace{3mm}
    <\texttt{DROP } L-5, \sigma> \Downarrow \sigma_1 \hspace{3mm}
    <\texttt{DROP } L-4, \sigma_1> \Downarrow \sigma_2 \hspace{3mm}
}
{	
	<\texttt{OP\_2ROT}, \sigma > \Downarrow \sigma'
}
\vspace{3mm}

% OP_2SWAP
\subsubsection{OP\_2SWAP}
The first and second items are swapped. The third and fourth items are swapped. \\

\inferrule
{	
	\sigma(|S|) = L \hspace{3mm}
    \sigma(S_{L}) = x \hspace{3mm}
    \sigma(S_{L-1}) = y \hspace{3mm}
	<\texttt{DROP } L, \sigma> \Downarrow \sigma_1 \hspace{3mm}
    <\texttt{DROP } L-1, \sigma_1> \Downarrow \sigma_2 \hspace{3mm} \\
    <\texttt{PUSH } S_{L-3}, \sigma_2> \Downarrow \sigma_3 \hspace{3mm}
    <\texttt{DROP } L-3, \sigma_3> \Downarrow \sigma_4 \hspace{3mm}
    <\texttt{PUSH } y, \sigma_4> \Downarrow \sigma_5 \hspace{3mm}
    <\texttt{PUSH } x, \sigma_5> \Downarrow \sigma' \hspace{3mm}
}
{	
	<\texttt{OP\_2SWAP}, \sigma > \Downarrow \sigma'
}
\vspace{3mm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Splice}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%OP_CAT
\subsubsection{OP\_CAT}
\textcolor{red}{\textit{Disabled}}\\
Concatenates two byte strings. \\

\inferrule
{	
	\sigma(|S|) = L \hspace{3mm}
    <\texttt{PUSH } S_{L} S_{L-1}, \sigma> \Downarrow \sigma_1 \hspace{3mm}
    <\texttt{DROP } L, \sigma_1> \Downarrow \sigma_2 \hspace{3mm}
    <\texttt{DROP } L-1, \sigma_2> \Downarrow \sigma' \hspace{3mm}
}
{	
	<\texttt{OP\_CAT}, \sigma > \Downarrow \sigma'
}
\vspace{3mm}

%OP_SUBSTR
\subsubsection{OP\_SUBSTR}
\textcolor{red}{\textit{Disabled}}\\
Returns a subtring of the top element, starting from index given by second element, of size given by third element. \\

\inferrule
{	
	\sigma(|S|) = L \hspace{3mm}
    \sigma(S_L) = b_1 \ldots b_k \hspace{3mm}
    \sigma(S_{L-1}) = start \hspace{3mm}
    \sigma(S_{L-2}) = size \hspace{3mm}
    start < k \hspace{3mm}
    start + size < k \hspace{3mm} \\
    <\texttt{PUSH } b_{start} \ldots b_{start+size}, \sigma> \Downarrow \sigma_1 \hspace{3mm}
    <\texttt{DROP } L, \sigma_1> \Downarrow \sigma_2 \hspace{3mm}
    <\texttt{DROP } L-1, \sigma_2> \Downarrow \sigma_3 \hspace{3mm}
    <\texttt{DROP } L-2, \sigma_3> \Downarrow \sigma' \hspace{3mm}
}
{	
	<\texttt{OP\_SUBSTR}, \sigma > \Downarrow \sigma'
}
\vspace{3mm}

%OP_LEFT
\subsubsection{OP\_LEFT}
\textcolor{red}{\textit{Disabled}}\\
Returns only character left of a specified point in top element, index given by second element. \\

\inferrule
{	
	\sigma(|S|) = L \hspace{3mm}
    \sigma(S_L) = b_1 \ldots b_k \hspace{3mm}
    \sigma(S_{L-1}) = idx \hspace{3mm}
    idx \leq k \hspace{3mm} \\
    <\texttt{PUSH } b_1 b_{idx-1}, \sigma> \Downarrow \sigma_1 \hspace{3mm}
    <\texttt{DROP } L, \sigma_1> \Downarrow \sigma_2 \hspace{3mm}
    <\texttt{DROP } L-1, \sigma_2> \Downarrow \sigma' \hspace{3mm}
}
{	
	<\texttt{OP\_LEFT}, \sigma > \Downarrow \sigma'
}
\vspace{3mm}

%OP_RIGHT
\subsubsection{OP\_RIGHT}
\textcolor{red}{\textit{Disabled}}\\
Returns only character right of a specified point in top element, index given by second element. \\

\inferrule
{	
	\sigma(|S|) = L \hspace{3mm}
    \sigma(S_L) = b_1 \ldots b_k \hspace{3mm}
    \sigma(S_{L-1}) = idx \hspace{3mm}
    idx \leq k \\
    <\texttt{PUSH } b_{idx} \ldots b_k, \sigma> \Downarrow \sigma_1 \hspace{3mm}
    <\texttt{DROP } L, \sigma_1> \Downarrow \sigma_2 \hspace{3mm}
    <\texttt{DROP } L-1, \sigma_2> \Downarrow \sigma' \hspace{3mm}
}
{	
	<\texttt{OP\_RIGHT}, \sigma > \Downarrow \sigma'
}
\vspace{3mm}

\subsubsection{OP\_SIZE}
Pushes the number of bytes of the top element of the stack. \\

\inferrule
{	
	\sigma(|S|) = L \hspace{3mm}
    \sigma(S_L) = B \hspace{3mm}
    <\texttt{PUSH } |B|, \sigma> \Downarrow \sigma'
}
{	
	<\texttt{OP\_SIZE}, \sigma > \Downarrow \sigma'
}
\vspace{3mm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Bitwise Logic}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% OP_INVERT
\subsubsection{OP\_INVERT}
\textcolor{red}{\textit{Disabled}}\\
Flips all bits of the top stack element. \\

\inferrule
{	
	\sigma(|S|) = L \hspace{3mm}
    \sigma(S_L) = b_1 \ldots b_k \hspace{3mm}
    <\texttt{PUSH } \lnot b_1 \lnot b_2 \ldots \lnot b_k, \sigma> \Downarrow \sigma_1 \hspace{3mm}
    <\texttt{DROP } L, \sigma_1> \Downarrow \sigma'
}
{	
	<\texttt{OP\_EQUAL}, \sigma > \Downarrow \sigma'
}
\vspace{3mm}

% OP_AND
\subsubsection{OP\_AND}
\textcolor{red}{\textit{Disabled}}\\
Boolean \textit{and} between bits of first and second stack elements. \\

\inferrule
{	
	\sigma(|S|) = L \hspace{3mm}
    \sigma(S_L) = B_1 \hspace{3mm}
    \sigma(S_{L-1}) = B_2 \hspace{3mm} \\
    <\texttt{PUSH } B_1 \mathrel{\&} B_2, \sigma> \Downarrow \sigma_1 \hspace{3mm}
    <\texttt{DROP } L, \sigma_1> \Downarrow \sigma_2 \hspace{3mm}
    <\texttt{DROP } L-1, \sigma_2> \Downarrow \sigma'
}
{	
	<\texttt{OP\_AND}, \sigma > \Downarrow \sigma'
}
\vspace{3mm}

% OP_OR
\subsubsection{OP\_OR}
\textcolor{red}{\textit{Disabled}}\\
Boolean \textit{and} between bits of first and second stack elements. \\

\inferrule
{	
	\sigma(|S|) = L \hspace{3mm}
    \sigma(S_L) = B_1 \hspace{3mm}
    \sigma(S_{L-1}) = B_2 \hspace{3mm} \\
    <\texttt{PUSH } B_1 \mathrel{|} B_2, \sigma> \Downarrow \sigma_1 \hspace{3mm}
    <\texttt{DROP } L, \sigma_1> \Downarrow \sigma_2 \hspace{3mm}
    <\texttt{DROP } L-1, \sigma_2> \Downarrow \sigma'
}
{	
	<\texttt{OP\_OR}, \sigma > \Downarrow \sigma'
}
\vspace{3mm}

% OP_XOR
\subsubsection{OP\_XOR}
\textcolor{red}{\textit{Disabled}}\\
Boolean \textit{and} between bits of first and second stack elements. \\

\inferrule
{	
	\sigma(|S|) = L \hspace{3mm}
    \sigma(S_L) = B_1 \hspace{3mm}
    \sigma(S_{L-1}) = B_2 \hspace{3mm} \\
    <\texttt{PUSH } B_1 \hspace{1mm} ^\wedge \hspace{1mm} B_2, \sigma> \Downarrow \sigma_1 \hspace{3mm}
    <\texttt{DROP } L, \sigma_1> \Downarrow \sigma_2 \hspace{3mm}
    <\texttt{DROP } L-1, \sigma_2> \Downarrow \sigma'
}
{	
	<\texttt{OP\_XOR}, \sigma > \Downarrow \sigma'
}
\vspace{3mm}

% OP_EQUAL
\subsubsection{OP\_EQUAL}
Returns 1 if the two top elements are equal, 0 otherwise. \\

\inferrule
{	
	\sigma(|S|) = L \hspace{3mm}
    \sigma(S_L) = B_1 \hspace{3mm}
    \sigma(S_{L-1}) = B_2 \hspace{3mm}
    B_1 = B_2 \hspace{3mm}
    <\texttt{PUSH } 1, \sigma> \Downarrow \sigma'
}
{	
	<\texttt{OP\_EQUAL}, \sigma > \Downarrow \sigma'
}
\vspace{3mm}

\inferrule
{	
	\sigma(|S|) = L \hspace{3mm}
    \sigma(S_L) = B_1 \hspace{3mm}
    \sigma(S_{L-1}) = B_2 \hspace{3mm}
    B_1 \neq B_2 \hspace{3mm}
    <\texttt{PUSH } 0, \sigma> \Downarrow \sigma'
}
{	
	<\texttt{OP\_EQUAL}, \sigma > \Downarrow \sigma'
}
\vspace{3mm}

% OP_EQUALVERIFY
\subsubsection{OP\_EQUALVERIFY}
Description. \\

\inferrule
{	
	<\texttt{OP\_EQUAL }, \sigma> \Downarrow \sigma_1
    <\texttt{OP\_VERIFY }, \sigma_1> \Downarrow \sigma'
}
{	
	<\texttt{OP\_EQUALVERIFY}, \sigma > \Downarrow \sigma'
}
\vspace{3mm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Arithmetic}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% OP_1ADD
\subsubsection{OP\_1ADD}
Add 1 to the top of the stack. \\

\inferrule
{	
	\sigma(|S|) = l \hspace{3mm}
    \sigma(S_l) \Downarrow x \hspace{3mm}
    B \Downarrow x+1 \hspace{3mm}
	<\texttt{DROP } l, \sigma> \Downarrow \sigma_1 \hspace{3mm}
    <\texttt{PUSH } B, \sigma_1> \Downarrow \sigma'
}
{	
	<\texttt{OP\_2SWAP}, \sigma > \Downarrow \sigma'
}
\vspace{3mm}

% OP_1SUB
\subsubsection{OP\_1SUB}
Subtract 1 from the top of the stack. \\

\inferrule
{	
	\sigma(|S|) = l \hspace{3mm}
    \sigma(S_l) \Downarrow x \hspace{3mm}
    B \Downarrow x-1 \hspace{3mm}
	<\texttt{DROP } l, \sigma> \Downarrow \sigma_1 \hspace{3mm}
    <\texttt{PUSH } B, \sigma_1> \Downarrow \sigma'
}
{	
	<\texttt{OP\_1SUB}, \sigma> \Downarrow \sigma'
}
\vspace{3mm}

% OP_ABS
\subsubsection{OP\_ABS}
The top of the stack is replaced with the byte-vector representing it's absolute value. \\

\inferrule
{	
	\sigma(|S|) = l \hspace{3mm}
    \sigma(S_l) \Downarrow x \hspace{3mm}
    B \Downarrow |x| \hspace{3mm}
	<\texttt{DROP } l, \sigma> \Downarrow \sigma_1 \hspace{3mm}
    <\texttt{PUSH } B, \sigma_1> \Downarrow \sigma'
}
{	
	<\texttt{OP\_1SUB}, \sigma> \Downarrow \sigma'
}
\vspace{3mm}

% OP_NEGATE
\subsubsection{OP\_NEGATE}
The top of the stack is negated. \\

\inferrule
{	
	\sigma(|S|) = l \hspace{3mm}
    \sigma(S_l) \Downarrow x \hspace{3mm}
    B \Downarrow -x \hspace{3mm}
	<\texttt{DROP } l, \sigma> \Downarrow \sigma_1 \hspace{3mm}
    <\texttt{PUSH } B, \sigma_1> \Downarrow \sigma'
}
{	
	<\texttt{OP\_1SUB}, \sigma> \Downarrow \sigma'
}
\vspace{3mm}

% OP_NOT
\subsubsection{OP\_NOT}
If the top of the stack is 0, replace it with the byte vector representing 1. Otherwise, replace it with the byte-vector representing 0. \\

\inferrule
{	
    <\texttt{OP\_IF PUSH OP\_0 OP\_ELSE PUSH OP\_1 OP\_ENDIF}, \sigma> \Downarrow \sigma'
}
{	
	<\texttt{OP\_NOT}, \sigma> \Downarrow \sigma'
}
\vspace{3mm}

% OP_0NOTEQUAL
\subsubsection{OP\_0NOTEQUAL}
Push the byte-vector representing 0 onto the stack if the top of stack is 0. Otherwise replace the top of stack with the byte-vector representing 1. \\

\inferrule
{	
    <\texttt{OP\_IF PUSH OP\_1 OP\_ELSE PUSH OP\_0 OP\_ENDIF}, \sigma> \Downarrow \sigma'
}
{	
	<\texttt{OP\_0NOTEQUAL}, \sigma> \Downarrow \sigma'
}
\vspace{3mm}

% OP_ADD
\subsubsection{OP\_ADD}
Let the inputs be the top two items of the stack, $a$ and $b$. Push the byte-vector representing $a+b$ onto the stack. \\

\inferrule
{	
	\sigma(|S|) = l \hspace{3mm}
    \sigma(S_l) \Downarrow x \hspace{3mm}
    \sigma(S_{l-1}) \Downarrow y \hspace{3mm}
    B \Downarrow x+y \hspace{3mm}
    <\texttt{OP\_2DROP}, \sigma> \Downarrow \sigma_1 \hspace{3mm}
    <\texttt{PUSH B}, \sigma_1> \Downarrow \sigma' \hspace{3mm}
}
{	
	<\texttt{OP\_ADD}, \sigma> \Downarrow \sigma'
}
\vspace{3mm}

% OP_SUB
\subsubsection{OP\_SUB}
Let the inputs be the top two items of the stack, $a$ and $b$. Push the byte-vector representing $a-b$ onto the stack. \\

\inferrule
{	
	\sigma(|S|) = l \hspace{3mm}
    \sigma(S_l) \Downarrow x \hspace{3mm}
    \sigma(S_{l-1}) \Downarrow y \hspace{3mm}
    B \Downarrow y-x \hspace{3mm}
    <\texttt{OP\_2DROP}, \sigma> \Downarrow \sigma_1 \hspace{3mm}
    <\texttt{PUSH B}, \sigma_1> \Downarrow \sigma' \hspace{3mm}
}
{	
	<\texttt{OP\_SUB}, \sigma> \Downarrow \sigma'
}
\vspace{3mm}

% OP_BOOLAND
\subsubsection{OP\_BOOLAND}
Let the inputs be the top two items of the stack, $a$ and $b$. If both $a$ and $b$ are not 0, then push 1 onto the stack. \\

\inferrule
{	
	\sigma(|S|) = L \hspace{3mm}
    \sigma(S_L) \Downarrow a \hspace{3mm}
    \sigma(S_{L-1}) \Downarrow b \hspace{3mm}
    a \neq 0 \land b \neq 0 \hspace{3mm}
    <\texttt{OP\_2DROP}, \sigma> \Downarrow \sigma_1 \hspace{3mm}
    <\texttt{PUSH OP\_1}, \sigma_1> \Downarrow \sigma' \hspace{3mm}
}
{	
	<\texttt{OP\_BOOLAND}, \sigma> \Downarrow \sigma'
}
\vspace{3mm}

\inferrule
{	
	\sigma(|S|) = L \hspace{3mm}
    \sigma(S_L) \Downarrow a \hspace{3mm}
    \sigma(S_{L-1}) \Downarrow b \hspace{3mm}
    a = 0 \lor b = 0 \hspace{3mm}
    <\texttt{OP\_2DROP}, \sigma> \Downarrow \sigma_1 \hspace{3mm}
    <\texttt{PUSH OP\_0}, \sigma> \Downarrow \sigma' \hspace{3mm}
}
{	
	<\texttt{OP\_BOOLAND}, \sigma> \Downarrow \sigma'
}
\vspace{3mm}

% OP_BOOLOR
\subsubsection{OP\_BOOLOR}
Let the inputs be the top two items of the stack, $a$ and $b$. If either $a$ or $b$ are not 0, then push 1 onto the stack. \\

\inferrule
{	
	\sigma(|S|) = L \hspace{3mm}
    \sigma(S_L) \Downarrow a \hspace{3mm}
    \sigma(S_{L-1}) \Downarrow b \hspace{3mm}
    a \neq 0 \lor b \neq 0 \hspace{3mm}
    <\texttt{OP\_2DROP}, \sigma> \Downarrow \sigma_1 \hspace{3mm}
    <\texttt{OP\_1}, \sigma_1> \Downarrow \sigma' \hspace{3mm}
}
{	
	<\texttt{OP\_BOOLOR}, \sigma> \Downarrow \sigma'
}
\vspace{3mm}

\inferrule
{	
	\sigma(|S|) = L \hspace{3mm}
    \sigma(S_L) \Downarrow a \hspace{3mm}
    \sigma(S_{L-1}) \Downarrow b \hspace{3mm}
    a \neq 0 \land b \neq 0 \hspace{3mm}
    <\texttt{OP\_2DROP}, \sigma> \Downarrow \sigma_1 \hspace{3mm}
    <\texttt{OP\_0}, \sigma_1> \Downarrow \sigma' \hspace{3mm}
}
{	
	<\texttt{OP\_BOOLOR}, \sigma> \Downarrow \sigma'
}
\vspace{3mm}

% OP_NUMEQUAL
\subsubsection{OP\_NUMEQUAL}
Let the inputs be the top two items of the stack, $a$ and $b$. Push 1 onto the stack if $a = b$; push 0 onto the stack otherwise. \\

\inferrule
{	
    <\texttt{OP\_SUB}, \sigma> \Downarrow \sigma_1 \hspace{3mm}
    <\texttt{OP\_IF PUSH OP\_0 OP\_ELSE PUSH OP\_1}, \sigma_1> \Downarrow \sigma' \hspace{3mm}
}
{	
	<\texttt{OP\_NUMEQUAL}, \sigma> \Downarrow \sigma'
}
\vspace{3mm}

% OP_NUMEQUALVERIFY
\subsubsection{OP\_NUMEQUALVERIFY}
Let the inputs be the top two items of the stack, $a$ and $b$. Push 1 onto the stack if $a = b$; push 0 onto the stack otherwise. If 0 is pushed onto the stack, mark the transaction as invalid. \\

\inferrule
{	
    <\texttt{OP\_NUMEQUAL}, \sigma> \Downarrow \sigma_1 \hspace{3mm}
    <\texttt{OP\_VERIFY}, \sigma_1> \Downarrow \sigma' \hspace{3mm}
}
{	
	<\texttt{OP\_NUMEQUALVERIFY}, \sigma> \Downarrow \sigma'
}
\vspace{3mm}

% OP_NUMNOTEQUAL
\subsubsection{OP\_NUMNOTEQUAL}
Let the inputs be the top two items of the stack, $a$ and $b$. Push 1 if the numbers are equal, 0 otherwise. \\

\inferrule
{	
    <\texttt{OP\_SUB}, \sigma> \Downarrow \sigma_1 \hspace{3mm}
    <\texttt{OP\_IF PUSH OP\_1 OP\_ELSE PUSH OP\_0}, \sigma_1> \Downarrow \sigma' \hspace{3mm}
}
{	
	<\texttt{OP\_NUMNOTEQUAL}, \sigma> \Downarrow \sigma'
}
\vspace{3mm}

% OP_LESSTHAN
\subsubsection{OP\_LESSTHAN}
Let the inputs be the element on the top of the stack(b) and the element second to the top (a).  Push 1 onto the stack if a is less than b, otherwise push 0. \\

\inferrule
{	\sigma(|S|) = L \hspace{3mm}
	\sigma(S_{L-1}) \Downarrow a \hspace{3mm}
	\sigma(S_L) \Downarrow b \hspace{3mm}
	a \geq b \hspace{3mm}
	<\texttt{OP\_2DROP}, \sigma> \Downarrow \sigma_1 \hspace{3mm}
	<\texttt{OP\_0}, \sigma_1> \Downarrow \sigma ' \hspace{3mm}
}
{	<\texttt{OP\_LESSTHAN}, \sigma> \Downarrow \sigma '
}
\vspace{3mm}

\inferrule
{	\sigma(|S|) = L \hspace{3mm}
	\sigma(S_{L-1}) \Downarrow a \hspace{3mm}
	\sigma(S_L) \Downarrow b \hspace{3mm}
	a < b \hspace{3mm}
	<\texttt{OP\_2DROP}, \sigma> \Downarrow \sigma_1 \hspace{3mm}
	<\texttt{OP\_1}, \sigma_1> \Downarrow \sigma ' \hspace{3mm}
}
{	<\texttt{OP\_LESSTHAN}, \sigma> \Downarrow \sigma '
}
\vspace{3mm}

% OP_GREATERTHAN
\subsubsection{OP\_GREATERTHAN}
Let the inputs be the element on the top of the stack(b) and the element second to the top (a).  Push 1 onto the stack if a is greater than b, otherwise push 0. \\

\inferrule
{	\sigma(|S|) = L \hspace{3mm}
	\sigma(S_{L-1}) \Downarrow a \hspace{3mm}
	\sigma(S_L) \Downarrow b \hspace{3mm}
	a \leq b \hspace{3mm}
	<\texttt{OP\_2DROP}, \sigma> \Downarrow \sigma_1 \hspace{3mm}
	<\texttt{OP\_0}, \sigma_1> \Downarrow \sigma ' \hspace{3mm}
}
{	<\texttt{OP\_GREATERTHAN}, \sigma> \Downarrow \sigma '
}
\vspace{3mm}

\inferrule
{	\sigma(|S|) = L \hspace{3mm}
	\sigma(S_{L-1}) \Downarrow a \hspace{3mm}
	\sigma(S_L) \Downarrow b \hspace{3mm}
	a > b \hspace{3mm}
	<\texttt{OP\_2DROP}, \sigma> \Downarrow \sigma_1 \hspace{3mm}
	<\texttt{OP\_1}, \sigma_1> \Downarrow \sigma ' \hspace{3mm}
}
{	<\texttt{OP\_GREATERTHAN}, \sigma> \Downarrow \sigma '
}
\vspace{3mm}

% OP_LESSTHANOREQUAL
\subsubsection{OP\_LESSTHANOREQUAL}
Let the inputs be the element on the top of the stack(b) and the element second to the top (a).  Push 1 onto the stack if a is less than or equal to b, otherwise push 0. \\

\inferrule
{	\sigma(|S|) = L \hspace{3mm}
	\sigma(S_{L-1}) \Downarrow a \hspace{3mm}
	\sigma(S_L) \Downarrow b \hspace{3mm}
	a > b \hspace{3mm}
	<\texttt{OP\_2DROP}, \sigma> \Downarrow \sigma_1 \hspace{3mm}
	<\texttt{OP\_0}, \sigma_1> \Downarrow \sigma ' \hspace{3mm}
}
{	<\texttt{OP\_LESSTHANOREQUAL}, \sigma> \Downarrow \sigma '
}
\vspace{3mm}

\inferrule
{	\sigma(|S|) = L \hspace{3mm}
	\sigma(S_{L-1}) \Downarrow a \hspace{3mm}
	\sigma(S_L) \Downarrow b \hspace{3mm}
	a \leq b \hspace{3mm}
	<\texttt{OP\_2DROP}, \sigma> \Downarrow \sigma_1 \hspace{3mm}
	<\texttt{OP\_1}, \sigma_1> \Downarrow \sigma ' \hspace{3mm}
}
{	<\texttt{OP\_LESSTHANOREQUAL}, \sigma> \Downarrow \sigma '
}
\vspace{3mm}

% OP_GREATERTHANOREQUAL
\subsubsection{OP\_GREATERTHANOREQUAL}
Let the inputs be the element on the top of the stack(b) and the element second to the top (a).  Push 1 onto the stack if a is greater than or equal to b, otherwise push 0. \\

\inferrule
{	\sigma(|S|) = L \hspace{3mm}
	\sigma(S_{L-1}) \Downarrow a \hspace{3mm}
	\sigma(S_L) \Downarrow b \hspace{3mm}
	a < b \hspace{3mm}
	<\texttt{OP\_2DROP}, \sigma> \Downarrow \sigma_1 \hspace{3mm}
	<\texttt{OP\_0}, \sigma_1> \Downarrow \sigma ' \hspace{3mm}
}
{	<\texttt{OP\_GREATERTHANOREQUAL}, \sigma> \Downarrow \sigma '
}
\vspace{3mm}

\inferrule
{	\sigma(|S|) = L \hspace{3mm}
	\sigma(S_{L-1}) \Downarrow a \hspace{3mm}
	\sigma(S_L) \Downarrow b \hspace{3mm}
	a \geq b \hspace{3mm}
	<\texttt{OP\_2DROP}, \sigma> \Downarrow \sigma_1 \hspace{3mm}
	<\texttt{OP\_1}, \sigma_1> \Downarrow \sigma ' \hspace{3mm}
}
{	<\texttt{OP\_GREATERTHANOREQUAL}, \sigma> \Downarrow \sigma '
}
\vspace{3mm}

% OP_MIN
\subsubsection{OP\_MIN}
the larger of the two elements at the top of the stack is removed. \\

\inferrule
{
    \sigma(|S|) = L \hspace{3mm})
    \sigma(S_{L-1}) \Downarrow a \hspace{3mm}
    \sigma(S_{L}) \Downarrow b \hspace{3mm}
    a \leq b \hspace{3mm}
    <\texttt{DROP }L, \sigma> \Downarrow \sigma' \hspace{3mm}
}
{
    <\texttt{OP\_MIN}, \sigma> \Downarrow \sigma'
}
\vspace{3mm}

\inferrule
{
    \sigma(|S|) = L \hspace{3mm}
    \sigma(S_{L-1}) \Downarrow a \hspace{3mm}
    \sigma(S_{L}) \Downarrow b \hspace{3mm}
    b < a \hspace{3mm}
    <\texttt{DROP } L-1, \sigma> \Downarrow \sigma' \hspace{3mm}
}
{
    <\texttt{OP\_MIN}, \sigma> \Downarrow \sigma'
}
\vspace{3mm}

% OP_MAX
\subsubsection{OP\_MAX}
the smaller of the two elements on the top of the stack is removed. \\

\inferrule
{
    \sigma(|S|) = L \hspace{3mm}
    \sigma(S_{L-1}) \Downarrow a \hspace{3mm}
    \sigma(S_{L}) \Downarrow b \hspace{3mm}
    a \geq b \hspace{3mm}
    <\texttt{DROP }L, \sigma> \Downarrow \sigma' \hspace{3mm}
}
{
    <\texttt{OP\_MAX}, \sigma> \Downarrow \sigma'
}
\vspace{3mm}

\inferrule
{
    \sigma(|S|) = L \hspace{3mm}
    \sigma(S_{L-1}) \Downarrow a \hspace{3mm}
    \sigma(S_{L}) \Downarrow b \hspace{3mm}
    b > a \hspace{3mm}
    <\texttt{DROP } L-1, \sigma> \Downarrow \sigma' \hspace{3mm}
}
{
    <\texttt{OP\_MAX}, \sigma> \Downarrow \sigma'
}
\vspace{3mm}

% OP_WITHIN
\subsubsection{OP\_WITHIN}
the input is the 3 elements at the top of the stack x, min, max where max is the top of the stack.  push 1 if x is greater than or equal to min and less than max, otherwise push 0. \\

\inferrule
{
    \sigma(|S|) = L \hspace{3mm}
    \sigma(S_{L}) \Downarrow max \hspace{3mm}
    \sigma(S_{L-1}) \Downarrow min \hspace{3mm}
    \sigma(S_{L-2}) \Downarrow x \hspace{3mm}
    min \leq x < max \hspace{3mm}
    <\texttt{OP\_1}, \sigma> \Downarrow \sigma' \hspace{3mm}
}
{
    <\texttt{OP\_WITHIN}, \sigma> \Downarrow \sigma'
}
\vspace{3mm}

\inferrule
{
    \sigma(|S|) = L \hspace{3mm}
    \sigma(S_{L}) \Downarrow max \hspace{3mm}
    \sigma(S_{L-1}) \Downarrow min \hspace{3mm}
    \sigma(S_{L-2}) \Downarrow x \hspace{3mm}
    \lnot (min \leq x < max) \hspace{3mm}
    <\texttt{OP\_1}, \sigma> \Downarrow \sigma' \hspace{3mm}
}
{
    <\texttt{OP\_WITHIN}, \sigma> \Downarrow \sigma'
}
\vspace{3mm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Crypto}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsubsection{OP\_RIPEMD160}
The element at the top of the stack is hashed with RIPEMD-160. \\

\inferrule
{
	\sigma(|S|) = L \hspace{3mm}
    B = \textsc{RIPEMD-160}(\sigma(S_L)) \hspace{3mm}
    <\texttt{DROP }L, \sigma> \Downarrow sigma_1 \hspace{3mm}
    <\texttt{PUSH }B, \sigma_1> \Downarrow sigma' \hspace{3mm}
}
{
	<\texttt{OP\_RIPEMD-160}, \sigma> \Downarrow sigma'
}
\vspace{3mm}

\subsubsection{OP\_SHA1}
The element at the top of the stack is hashed with SHA-1. \\

\inferrule
{
	\sigma(|S|) = L \hspace{3mm}
    B = \textsc{SHA-1}(\sigma(S_L)) \hspace{3mm}
    <\texttt{DROP }L, \sigma> \Downarrow sigma_1 \hspace{3mm}
    <\texttt{PUSH }B, \sigma_1> \Downarrow sigma' \hspace{3mm}
}
{
	<\texttt{OP\_SHA1}, \sigma> \Downarrow sigma'
}
\vspace{3mm}
\subsubsection{OP\_SHA256}
The element at the top of the stack is hashed with SHA-256. \\

\inferrule
{
	\sigma(|S|) = L \hspace{3mm}
    B = \textsc{SHA-256}(\sigma(S_L)) \hspace{3mm}
    <\texttt{DROP }L, \sigma> \Downarrow sigma_1 \hspace{3mm}
    <\texttt{PUSH }B, \sigma_1> \Downarrow sigma' \hspace{3mm}
}
{
	<\texttt{OP\_SHA-256}, \sigma> \Downarrow sigma'
}
\vspace{3mm}
\subsubsection{OP\_HASH160}
The element at the top of the stack is hashed with SHA-256 and then with RIPEMD-160. \\

\inferrule
{
	\sigma(|S|) = L \hspace{3mm}
    B = \textsc{RIPEMD-160}(\textsc{SHA-256}(\sigma(S_L))) \hspace{3mm}
    <\texttt{DROP }L, \sigma> \Downarrow sigma_1 \hspace{3mm}
    <\texttt{PUSH }B, \sigma_1> \Downarrow sigma' \hspace{3mm}
}
{
	<\texttt{OP\_HASH160}, \sigma> \Downarrow sigma'
}
\vspace{3mm}

\subsubsection{OP\_HASH256}
The element at the top of the stack is hashed twice with SHA-256. \\

\inferrule
{
	\sigma(|S|) = L \hspace{3mm}
    B = \textsc{SHA-256}(\textsc{SHA-256}(\sigma(S_L))) \hspace{3mm}
    <\texttt{DROP }L, \sigma> \Downarrow sigma_1 \hspace{3mm}
    <\texttt{PUSH }B, \sigma_1> \Downarrow sigma' \hspace{3mm}
}
{
	<\texttt{OP\_HASH256}, \sigma> \Downarrow sigma'
}
\vspace{3mm}

\subsubsection{OP\_CODESEPARATOR}
\subsubsection{OP\_CHECKSIG}
\subsubsection{OP\_CHECKSIGVERIFY}
\subsubsection{OP\_CHECKMULTISIG}
\subsubsection{OP\_CHECKMULTISIGVERIFY}
\subsection{Locktime}
\subsubsection{OP\_CHECKLOCKTIMEVERIFY (previously OP\_NOP2)}
\subsubsection{OP\_CHECKSEQUENCEVERIFY (previously OP\_NOP3)}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Disabled / Reserved / Invalid}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
All disabled, reserved or otherwise invalid opcodes transition to an invalid state.

\inferrule
{\quad}
{<\texttt{dcom},\sigma> \Downarrow \sigma[V=invalid]}

\section{Operational Semantics -- Small Step Semantics}

\subsection{Local reduction rules}

[Have to explain what are atomic commands, and why we're defining them.]

[No need to add PUSHDATA commands: these can be described as atomic commands, if we remember that the PUSHDATA opcode is only part of the push commands defined in the syntax - Hugo]

\begin{align*}
&<\texttt{OP\_NOP com}, \sigma> 
	\hspace{3mm}
\longrightarrow 
	\hspace{3mm}
<\texttt{com}, \sigma>
\\
&<\texttt{atom\_com}, \sigma[S=s, V=v]> 
	\hspace{3mm}
\longrightarrow 
	\hspace{3mm}
<\texttt{OP\_NOP}, \sigma[S=s', V=v']>
\\
&<\texttt{OP\_IF com OP\_ENDIF}, \sigma[S_{|S|} \neq 0]> 
	\hspace{3mm}
\longrightarrow 
	\hspace{3mm}
<\texttt{com}, \sigma> 
\\
&<\texttt{OP\_IF com OP\_ENDIF}, \sigma[S_{|S|} = 0]> 
	\hspace{3mm}
\longrightarrow 
	\hspace{3mm}
<\texttt{OP\_NOP}, \sigma> 
\\
&<\texttt{OP\_NOTIF com OP\_ENDIF}, \sigma[S_{|S|} \neq 0]> 
	\hspace{3mm}
\longrightarrow 
	\hspace{3mm}
<\texttt{OP\_NOP}, \sigma> 
\\
&<\texttt{OP\_NOTIF com OP\_ENDIF}, \sigma[S_{|S|} = 0]> 
	\hspace{3mm}
\longrightarrow 
	\hspace{3mm}
<\texttt{com}, \sigma> 
\\
&<\texttt{OP\_IF com$_1$ OP\_ELSE com$_2$ OP\_ENDIF}, \sigma[S_{|S|} \neq 0]> 
	\hspace{3mm}
\longrightarrow 
	\hspace{3mm}
<\texttt{com$_1$}, \sigma> 
\\
&<\texttt{OP\_IF com$_1$ OP\_ELSE com$_2$ OP\_ENDIF}, \sigma[S_{|S|} = 0]> 
	\hspace{3mm}
\longrightarrow 
	\hspace{3mm}
<\texttt{com$_2$}, \sigma> 
\\
&<\texttt{OP\_NOTIF com$_1$ OP\_ELSE com$_2$ OP\_ENDIF}, \sigma[S_{|S|} \neq 0]> 
	\hspace{3mm}
\longrightarrow 
	\hspace{3mm}
<\texttt{com$_2$}, \sigma> 
\\
&<\texttt{OP\_NOTIF com$_1$ OP\_ELSE com$_2$ OP\_ENDIF}, \sigma[S_{|S|} = 0]> 
	\hspace{3mm}
\longrightarrow 
	\hspace{3mm}
<\texttt{com$_1$}, \sigma> 
\\
\end{align*}

\subsection{Global reduction rules}

Since programs in Script are evaluated strictly from left to right, there is really only global reduction rule aside from $\bullet$.

$$H ::= \bullet \mid H \hspace{1mm} \texttt{com}$$

\end{document}

### 2017-03-19 paper.tex
changed related work section

@@
 
-The best documentation for Script exists on the Bitcoin wiki [2].
-This documentation is written in plain English, and not in a logical formalism that allows mathematical reasoning.
-To our knowledge, no previous effort has been made to provide a formal specification of Script.
+To our knowledge, there has been no previous effort to provide a formal specification of the Script language.
+As a result, the existing documentation works as a rough guideline for our work but is not complete.
+Explanations of how Script works are located on the bitcoin wiki [2] and the reference implementation [1].
+
+The bitcoin wiki is an explanation of how each opcode works in plain english.
+The explanations are not rigorous, and are not in a logical formalism that allows mathematical reasoning.
+There are many situations where descriptions are vague and can be interpreted in more than one way.
+For example, the explanation of OP\_LEFT takes an argument to index a string and throws away everything to the right of the index, but it doesnt tell us if that operation is right-inclusive.
+
+The reference implementation is much more thorough that the wiki but, being c++ code, it is much harder to interpret and is not in a format that allows for logical reasoning.
 If we turn to Forth to find something related, we can find standard specifications like the Forth-83 standard (http://forth.sourceforge.net/standard/fst83/fst83-12.htm).
-However even these specifications are also given in plain English.
+However even these specifications are given in plain English.
 

### 2017-03-19 paper.tex
Extended intro

--- 2017-03-19 paper.tex
@@
 
-As Bitcoin's economic presence grows, formal specifications, abstracted away from current current implementations of the Bitcoin protocol, will become increasingly essential in mediating conflict over what Bitcoin is or should be, allowing participants in the Bitcoin community to reason formally and precisely about their opinions and proposals.
+To prevent attacks on the Bitcoin network based on exploiting a flaw in Script, the Bitcoin community agreed on a set of ``standard'' script templates that nodes will process.
+These limitations, however, are deemed temporary and subject to change in the future [3], although no consensus exists at the moment on what should replace these limitations, if anything.
+As Bitcoin's economic presence grows, formal specifications, abstracted away from current implementations of the Bitcoin protocol, will become increasingly essential in mediating conflict over what Bitcoin is or should be, allowing participants in the Bitcoin community to reason formally and precisely about their opinions and proposals.
 
-\textcolor{blue}{[Describe history of Script and Bitcoin maybe, say the language was flawed]}
-
-\textcolor{blue}{[Mention [3] when it says that "These limitations are encoded in a function called isStandard(), which defines five types of "standard" transactions.
-These limitations are temporary and might be lifted by the time you read this."
-The transaction programs being subject to change justifies allowing formal reasoning about the language.]}
+\textcolor{blue}{[anything else? can we add more examples of how a lack of formal specification was problematic?]}
 

### 2017-03-19 paper.tex
added difficulty encountered to {Application of Script Grammar}

@@
 
+\subsubsection{Difficulties encountered}
+
+1) It was our hope that we could define operational semantics for op codes of previous versions of Script that caused crashes or other issues in the past.
+Using these definitions, we would then show that the old definitions could be used to cause problems.
+Unfortunately, there is very little documentation regarding how these errors came about.
+The reference documentation itself only documents changes back to 2014 while most of the serious problems with Script were fixed in 2010.
+The only fix we were able to analyze was the change to OP\_RETURN.
+
 \section{APPLICATIONS AND RESULTS}

### 2017-03-19 paper.tex
Completed application A discussion and added relevant reference

--- 2017-03-19 paper.tex
@@
 We also know, from our description of Script's big-step semantics in Section 2 of the appendix, that every command requires a constant number of checks and operations on the stack.
+So we have that a Script program will always halt, and that the number of operations required to evaluate it is at most linear in the amount of commands it contains.
 
-So we have that a Script program will always halt, and that the number of operations required to evaluate it is at most linear in the amount of commands it contains.
+Given that the main nodes on the Bitcoin network enforce a limit on the length of Script programs one can include in a transaction, guaranteed linear execution time results in an absolute maximum number of steps for a standard Script program.
+This helps in protecting against DoS attacks based on maliciously crafted transactions.
+Furthermore, this maximum number of steps was deliberately lowered by requiring that the most computationally expensive opcodes, \texttt{OP\_CHECKSIG}, be contained exactly once in a valid standard transaction [4].
 
@@
 
+\bibitem{c4} CVE Details.
+(March 3 2012).
+\textit{Vulnerability Details : CVE-2013-2292} [Online; Accessed: 15-March-2017].
+Available: https://www.cvedetails.com/cve-details.php?t=1\&cve\_id=CVE-2013-2292.
+
 \end{thebibliography}

### 2017-03-19 paper.tex
Grammar in Related Work

@@
 
-\textcolor{blue}{COMPLETELY REVISE THIS}
-
 To our knowledge, there has been no previous effort to provide a formal specification of the Script language.
 As a result, the existing documentation works as a rough guideline for our work but is not complete.
-Explanations of how Script works are located on the bitcoin wiki [2] and the reference implementation [1].
+Explanations of how Script works are located on the Bitcoin wiki [2] and the reference implementation [1].
 
-The bitcoin wiki is an explanation of how each opcode works in plain english.
+The Bitcoin wiki is an explanation of how each opcode works in plain English.
 The explanations are not rigorous, and are not in a logical formalism that allows mathematical reasoning.
 There are many situations where descriptions are vague and can be interpreted in more than one way.
-For example, the explanation of OP\_LEFT takes an argument to index a string and throws away everything to the right of the index, but it doesnt tell us if that operation is right-inclusive.
+For example, the explanation of \texttt{OP\_LEFT} takes an argument to index a string and throws away everything to the right of the index, but it does not specify whether that operation is right-inclusive.
 
-The reference implementation is much more thorough that the wiki but, being c++ code, it is much harder to interpret and is not in a format that allows for logical reasoning.
+The reference implementation is much more thorough that the wiki but, being C++ code, it is much harder to interpret and is not in a format that allows for logical reasoning.
 If we turn to Forth to find something related, we can find standard specifications like the Forth-83 standard (http://forth.sourceforge.net/standard/fst83/fst83-12.htm).

### 2017-03-19 paper.tex
formatting in {Application of Script Grammar}
Update paper.tex
Moved Difficulties encountered section

@@
 
-\subsubsection{Difficulties encountered}
+\subsection{Application of Script Grammar}
 
-  1) The source code of a script program is just a sequence of bytes and is not human-readable.
+[Describe how grammar was used]
+
+\subsection{Difficulties encountered}
+
+$\bullet$ The source code of a script program is just a sequence of bytes and is not human-readable.
 To make documentation easier, the creators of the language defined an alias (\textit{opcode}) for each command.
@@
 
-\subsection{Application of Script Grammar}
-
-[Describe how grammar was used]
-
-\subsubsection{Difficulties encountered}
-
-1) It was our hope that we could define operational semantics for op codes of previous versions of Script that caused crashes or other issues in the past.
+$\bullet$ It was our hope that we could define operational semantics for op codes of previous versions of Script that caused crashes or other issues in the past.
 Using these definitions, we would then show that the old definitions could be used to cause problems.
@@
 The reference documentation itself only documents changes back to 2014 while most of the serious problems with Script were fixed in 2010.
-The only fix we were able to analyze was the change to OP\_RETURN.
+The only fix we were able to analyze was the change to \texttt{OP\_RETURN}.
+
+$\bullet$ \texttt{OP\_CHECKSIG} is a bitch to describe.
 

### 2017-03-20 paper.tex
Added application for OP _RETURN analysis.

@@
 %\usepackage{times} % assumes new font selection scheme installed
-%\usepackage{amsmath} % assumes amsmath package installed
-%\usepackage{amssymb}  % assumes amsmath package installed
+\usepackage{amsmath} % assumes amsmath package installed
+\usepackage{amssymb}  % assumes amsmath package installed
 \usepackage{xcolor}
@@
 
- \textcolor{red}{\textit{If appropriate, you should describe your implementation and experimental results. Explain how to interpret your numbers and results if applicable.}}
-
-\textcolor{blue}{[Motivation for the application cases we chose.]}
-
 \subsection{Programs written in the current version of Script halt within time linear in the length of the program}
 
+\subsubsection{Background}
 For the Bitcoin network to function properly, it is important that the network form a single connected component.
@@
 
+\subsubsection{Analysis}
 Using the small-step semantics defined in Section 3 of the appendix, we see that no reduction rule yields its entire input, as was the case with the \texttt{while} local reduction rule in IMP.
@@
 
-\subsection{Previous versions of Script allowed undesirable programs}
-Here we prove how previous states were problematic
+\subsection{Previous versions of Script allowed skeleton key scripts}
+
+\subsubsection{Background}
+The way an amount on Bitcoin is spent on the network is by evaluating to $\sigma[V=valid]$\footnote{See Appendix section 2 for a description of state configurations.} the concatenation $C = BA$ of a locking script A, provided by the payer, and an unlocking script B provided by the recipient.
+The intended scenario is that $A$ will contain a validation condition, which will check whether the output of $B$ is adequate.
+Typically, $B$ will have to provide a hash that proves they are the intended recipient of the money spent by the payer, and $A$ will know what to expect as a proof.
+In that way, $A$ is a guarantee to the payer that the money can only be received by the intended recipient.
+However, a careless definition of the \texttt{OP\_RETURN} opcode in previous versions of Script made it possible to program skeleton key scripts that could spend any amount of money, without proving the recipient's identity.
+
+\subsubsection{Analysis}
+A skeleton script would have the following property: 
+$$\exists B. \forall A. \hspace{1.5mm} BA \Downarrow \sigma[V=valid]$$
+
+This is equivalent to saying that either (1) $B$ is such that for any $A$, $BA$ evaluates to a valid state, or (2) $B$ evaluates to a state from which any $A$ will evaluate to a valid state.
+Case (2) was never possible, since one could write a script $A = \texttt{OP\_0}$, that pushes 0 on the stack.
+Upon normal program termination, if the top stack value is 0, then the program is deemed invalid.
+Thus for this $A$ there is no script $B$ such that $BA \Downarrow \sigma[V=valid]$.
+Case (1), however, was made possible by the semantics of \texttt{OP\_RETURN} prior to July 31 2010.
+In local reduction rule form, \texttt{OP\_RETURN} was defined as follows:
+%
+$$<\texttt{OP\_RETURN com}, \sigma> \hspace{3mm} \longrightarrow \hspace{3mm} <\texttt{OP\_NOP}, \sigma>$$
+%
+Evaluating \texttt{OP\_RETURN} results in the immediate termination of the program, \textit{without any changes to the state}.
+Consider now a script $B = \texttt{OP\_1 OP\_RETURN}$, and the evaluation of $BA$ for any arbitrary script $A$:
+
+\begin{enumerate}
+  \item \texttt{1} is pushed onto the stack
+  \item \texttt{OP\_RETURN} triggers program termination
+  \item The top stack value is nonzero, and so the program is deemed valid
+\end{enumerate}
+
+Clearly, that such a script was possible completely defeated the transaction validation process. \texttt{OP\_RETURN} semantics were subsequently changed to the following:
+%
+$$<\texttt{OP\_RETURN com}, \sigma> \hspace{1mm} \longrightarrow \hspace{1mm} <\texttt{OP\_NOP}, \sigma[V=invalid]>$$
+%
+thus ensuring that its evaluation in a script invariably resulted in a rejected transaction.
+As no other command in the language can cause early termination of the program with a valid state, we can now say that there exists no skeleton key script that could serve to unlock any arbitrary locking script.
 
 \subsection{The Script command set could be reduced without loss of expressiveness}
-Here we prove that some commands are redundant. Must exaplin why this can be desirable sometimes, but also why some of them are clearly uselessly redundant.
+Here we prove that some commands are redundant. Must explain why this can be desirable sometimes, but also why some of them are clearly uselessly redundant.
 

### 2017-03-20 paper.tex
Removed red instructions from the paper

@@
 
-\textcolor{red}{\textit{To describe and motivate the problem or your survey topic }}\\
-
 Script is what Bitcoin uses to encode transactions over the network.
@@
 \section{APPROACH}
-
-\textcolor{red}{\textit{Detailed description of the approach, and highlight the main technical difficulties and novelties. }}\\
 
@@
 
-$\bullet$ \texttt{OP\_CHECKSIG} is a bitch to describe.
+%$\bullet$ \texttt{OP\_CHECKSIG} is a bitch to describe.
 
@@
 
-\textcolor{red}{\textit{Detailed discussion of related work. Should stress how these relate to your work. Simply list and describe what other people did is not sufficient.}}\\
-
 To our knowledge, there has been no previous effort to provide a formal specification of the Script language.
@@
 \section{CONCLUSIONS}
-
-\textcolor{red}{\textit{Again summarize the work and discuss possible future work.}}\\
 

### 2017-03-20 paper.tex
Fixed typo

@@
 
+In this section, we use our grammar
+
 \subsection{Programs written in the current version of Script halt within time linear in the length of the program}
@@
 This helps in protecting against DoS attacks based on maliciously crafted transactions.
-Furthermore, this maximum number of steps was deliberately lowered by requiring that the most computationally expensive opcodes, \texttt{OP\_CHECKSIG}, be contained exactly once in a valid standard transaction [4].
+Furthermore, this maximum number of steps was deliberately lowered by requiring that the most computationally expensive opcode, \texttt{OP\_CHECKSIG}, be contained exactly once in a valid standard transaction [4].
 

### 2017-03-20 paper.tex
Removed 3rd application, won't do it.
rough rough ugly draft of 'Approach'

--- 2017-03-20 paper.tex
@@
 %\usepackage{times} % assumes new font selection scheme installed
-\usepackage{amsmath} % assumes amsmath package installed
-\usepackage{amssymb}  % assumes amsmath package installed
+%\usepackage{amsmath} % assumes amsmath package installed
+%\usepackage{amssymb}  % assumes amsmath package installed
 \usepackage{xcolor}
@@
 \subsection{Definition of Script Grammar}
+The bulk of our project consists in providing a well-formed and consistent definition of the Script language.
+This includes a description of the syntax that the text of scripts must conform to, as well descriptions of the operational semantics--big-step and small-step--by which the scripts evaluate.
+Due to its breadth, we defer the full description of this grammar to [Appendix A].
+
+The reference implementation of Bitcoin defines the network protocol.
+Script, as the language which expresses and encodes transactions, exists as a subset of the Bitcoin protocol and is therefore also defined by the reference implementation.
+
+In order to generate this grammar, it was ultimately necessary to directly turn to the source code of the reference implementation to validate the correctness of our grammar.
+Proving the consistency or equivalence of our work with the reference implementation is outside the scope of this project and is left as future work.
+In the process of forming both our understanding and therefrom
+
+we searched through several reputable sources in order to gain a working understanding of Script, eventually turning to the reference Bitcoin implementation to secure and progressively refine our understanding.
 
 \subsubsection{Syntax}
-\# Bitcoin wiki -> tables describe the format\\
-\# transaction is X Y, inputs are such and such\\
-\# looked at enums in source code to get full list of defined opcode and their hex values\\
-JOSEPH: also explain how the different syntactic groups were broken down\\
+
+First, it was necessary to define the syntax.
+We made heavy use of the [Bitcoin Wiki] to this end.
+Script is a Forth-like, stack-based language.
+It is parsed and evaluated left to right [wiki/script].
+As it is defined by it's reference implementation, in order to implement the language one must ensure that in all circumstances one's implementation interprets the binary serialization of a script in a manner compatible with the reference.
+
+It was therefore quite natural to define the syntax of the literal bytes that form programs.
+This is consistent with the internal subdivisions one may discern within a script program.
+Script defines words or opcodes with defined effects on the state of the program.
+We look to both the [Bitcoin Wiki] and the source code to retrieve a full list of defined opcodes and to infer and ultimately describe their operations.
+
+The syntax in [Appendix A] defines terms the form of a transaction, with scripts existing within transactions in defined positions.
+We provide aliases for opcodes' hex values and group commands by both their effect--arithmetic, control, crypto, etc, and their form--whether they are "atomic" or whether the opcode relies on following words or bytes to affect how it transform the state of the program and to give the opcode syntactic validity--the script \texttt{OP\_IF OP\_0 OP\_ELSE} is invalid because it has no \texttt{OP\_ENDIF}.
 
@@
 
+After providing a sufficient syntax, we moved to describe the operational semantics.
 We described the big-step semantics and the small-step semantics of the language using a combination of the reference implementation of the language [Ref] and of the Script page of the Bitcoin wiki [Ref].
+
+We provide inference rules for every opcode.
+The state of the program consists of the stack, $S$, the alt-stack, $AS$, the validity $V$, and the rest of the program to be parsed $P$.
+Script operates primarily on byte-vectors representing numbers or arbitrary data.
+We speak of a byte-vector evaluating to a number under a set of encoding rules specific to Bitcoin but not completely described.
+
+A number of op-codes have been disabled due to perceived vulnerabilities, reserved for future use, or reserved for internal use in testing.
+These are given their own syntactic category and result in an invalid state.
+
+Scripts exist in transactions and op-codes exist which directly reference their transaction context. We omit these because of the disproportionate descriptive burden they introduce.
+
+We also provide small step semantics to describe the progression of evaluation of a script program. The local and global reduction rules are compact due to the language's inherent simplicity.
 
@@
 
+ \textcolor{red}{\textit{If appropriate, you should describe your implementation and experimental results. Explain how to interpret your numbers and results if applicable.}}
+
+\textcolor{blue}{[Motivation for the application cases we chose.]}
+
 \subsection{Programs written in the current version of Script halt within time linear in the length of the program}
 
-\subsubsection{Background}
 For the Bitcoin network to function properly, it is important that the network form a single connected component.
@@
 
-\subsubsection{Analysis}
 Using the small-step semantics defined in Section 3 of the appendix, we see that no reduction rule yields its entire input, as was the case with the \texttt{while} local reduction rule in IMP.
@@
 
-\subsection{Previous versions of Script allowed skeleton key scripts}
-
-\subsubsection{Background}
-The way an amount on Bitcoin is spent on the network is by evaluating to $\sigma[V=valid]$\footnote{See Appendix section 2 for a description of state configurations.} the concatenation $C = BA$ of a locking script A, provided by the payer, and an unlocking script B provided by the recipient.
-The intended scenario is that $A$ will contain a validation condition, which will check whether the output of $B$ is adequate.
-Typically, $B$ will have to provide a hash that proves they are the intended recipient of the money spent by the payer, and $A$ will know what to expect as a proof.
-In that way, $A$ is a guarantee to the payer that the money can only be received by the intended recipient.
-However, a careless definition of the \texttt{OP\_RETURN} opcode in previous versions of Script made it possible to program skeleton key scripts that could spend any amount of money, without proving the recipient's identity.
-
-\subsubsection{Analysis}
-A skeleton script would have the following property: 
-$$\exists B. \forall A. \hspace{1.5mm} BA \Downarrow \sigma[V=valid]$$
-
-This is equivalent to saying that either (1) $B$ is such that for any $A$, $BA$ evaluates to a valid state, or (2) $B$ evaluates to a state from which any $A$ will evaluate to a valid state.
-Case (2) was never possible, since one could write a script $A = \texttt{OP\_0}$, that pushes 0 on the stack.
-Upon normal program termination, if the top stack value is 0, then the program is deemed invalid.
-Thus for this $A$ there is no script $B$ such that $BA \Downarrow \sigma[V=valid]$.
-Case (1), however, was made possible by the semantics of \texttt{OP\_RETURN} prior to July 31 2010.
-In local reduction rule form, \texttt{OP\_RETURN} was defined as follows:
-%
-$$<\texttt{OP\_RETURN com}, \sigma> \hspace{3mm} \longrightarrow \hspace{3mm} <\texttt{OP\_NOP}, \sigma>$$
-%
-Evaluating \texttt{OP\_RETURN} results in the immediate termination of the program, \textit{without any changes to the state}.
-Consider now a script $B = \texttt{OP\_1 OP\_RETURN}$, and the evaluation of $BA$ for any arbitrary script $A$:
-
-\begin{enumerate}
-  \item \texttt{1} is pushed onto the stack
-  \item \texttt{OP\_RETURN} triggers program termination
-  \item The top stack value is nonzero, and so the program is deemed valid
-\end{enumerate}
-
-Clearly, that such a script was possible completely defeated the transaction validation process. \texttt{OP\_RETURN} semantics were subsequently changed to the following:
-%
-$$<\texttt{OP\_RETURN com}, \sigma> \hspace{1mm} \longrightarrow \hspace{1mm} <\texttt{OP\_NOP}, \sigma[V=invalid]>$$
-%
-thus ensuring that its evaluation in a script invariably resulted in a rejected transaction.
-As no other command in the language can cause early termination of the program with a valid state, we can now say that there exists no skeleton key script that could serve to unlock any arbitrary locking script.
+\subsection{Previous versions of Script allowed undesirable programs}
+Here we prove how previous states were problematic
 
 \subsection{The Script command set could be reduced without loss of expressiveness}
-Here we prove that some commands are redundant. Must explain why this can be desirable sometimes, but also why some of them are clearly uselessly redundant.
+Here we prove that some commands are redundant. Must exaplin why this can be desirable sometimes, but also why some of them are clearly uselessly redundant.
 

### 2017-03-20 paper.tex

--- 2017-03-20 paper.tex
@@
 
-\subsection{The Script command set could be reduced without loss of expressiveness}
-Here we prove that some commands are redundant. Must explain why this can be desirable sometimes, but also why some of them are clearly uselessly redundant.
-
 \section{RELATED WORK}

### 2017-03-20 paper.tex
added OP_CHECKSIG to difficulties

@@
 
-%$\bullet$ \texttt{OP\_CHECKSIG} is a bitch to describe.
+$\bullet$ \texttt{OP\_CHECKSIG} and related op-codes refer to the serialized form of the surrounding transaction.
+This is important as the signature which \texttt{OP\_CHECKSIG} verifies serves dually to prove that the author of the transaction is in control of the relevant private key (corresponding to the public key which hashes to the address to which the relevant locked bitcoins were spent) and in fact authored the transaction presently under consideration.
+Due to the complexities of referencing the serialized form the transaction with big-step semantic inference rules, we did not, after much struggle, provide such rules for such op-codes.
 

### 2017-03-20 paper.tex
Beefed up the Related Work section

--- 2017-03-20 paper.tex
@@
 \subsection{Definition of Script Grammar}
-The bulk of our project consists in providing a well-formed and consistent definition of the Script language.
-This includes a description of the syntax that the text of scripts must conform to, as well descriptions of the operational semantics--big-step and small-step--by which the scripts evaluate.
-Due to its breadth, we defer the full description of this grammar to [Appendix A].
-
-The reference implementation of Bitcoin defines the network protocol.
-Script, as the language which expresses and encodes transactions, exists as a subset of the Bitcoin protocol and is therefore also defined by the reference implementation.
-
-In order to generate this grammar, it was ultimately necessary to directly turn to the source code of the reference implementation to validate the correctness of our grammar.
-Proving the consistency or equivalence of our work with the reference implementation is outside the scope of this project and is left as future work.
-In the process of forming both our understanding and therefrom
-
-we searched through several reputable sources in order to gain a working understanding of Script, eventually turning to the reference Bitcoin implementation to secure and progressively refine our understanding.
 
 \subsubsection{Syntax}
-
-First, it was necessary to define the syntax.
-We made heavy use of the [Bitcoin Wiki] to this end.
-Script is a Forth-like, stack-based language.
-It is parsed and evaluated left to right [wiki/script].
-As it is defined by it's reference implementation, in order to implement the language one must ensure that in all circumstances one's implementation interprets the binary serialization of a script in a manner compatible with the reference.
-
-It was therefore quite natural to define the syntax of the literal bytes that form programs.
-This is consistent with the internal subdivisions one may discern within a script program.
-Script defines words or opcodes with defined effects on the state of the program.
-We look to both the [Bitcoin Wiki] and the source code to retrieve a full list of defined opcodes and to infer and ultimately describe their operations.
-
-The syntax in [Appendix A] defines terms the form of a transaction, with scripts existing within transactions in defined positions.
-We provide aliases for opcodes' hex values and group commands by both their effect--arithmetic, control, crypto, etc, and their form--whether they are "atomic" or whether the opcode relies on following words or bytes to affect how it transform the state of the program and to give the opcode syntactic validity--the script \texttt{OP\_IF OP\_0 OP\_ELSE} is invalid because it has no \texttt{OP\_ENDIF}.
+\# Bitcoin wiki -> tables describe the format\\
+\# transaction is X Y, inputs are such and such\\
+\# looked at enums in source code to get full list of defined opcode and their hex values\\
+JOSEPH: also explain how the different syntactic groups were broken down\\
 
@@
 
-After providing a sufficient syntax, we moved to describe the operational semantics.
 We described the big-step semantics and the small-step semantics of the language using a combination of the reference implementation of the language [Ref] and of the Script page of the Bitcoin wiki [Ref].
-
-We provide inference rules for every opcode.
-The state of the program consists of the stack, $S$, the alt-stack, $AS$, the validity $V$, and the rest of the program to be parsed $P$.
-Script operates primarily on byte-vectors representing numbers or arbitrary data.
-We speak of a byte-vector evaluating to a number under a set of encoding rules specific to Bitcoin but not completely described.
-
-A number of op-codes have been disabled due to perceived vulnerabilities, reserved for future use, or reserved for internal use in testing.
-These are given their own syntactic category and result in an invalid state.
-
-Scripts exist in transactions and op-codes exist which directly reference their transaction context. We omit these because of the disproportionate descriptive burden they introduce.
-
-We also provide small step semantics to describe the progression of evaluation of a script program. The local and global reduction rules are compact due to the language's inherent simplicity.
 
@@
 
-To our knowledge, there has been no previous effort to provide a formal specification of the Script language.
-As a result, the existing documentation works as a rough guideline for our work but is not complete.
-Explanations of how Script works are located on the Bitcoin wiki [2] and the reference implementation [1].
+To our knowledge, no previous effort was made to provide a formal description of Script.
+The Bitcoin developers community relies mostly on a combination of the Bitcoin wiki Script article [2] and Script's C++ reference implementation [1] to learn the language or reason about it.
 
-The Bitcoin wiki is an explanation of how each opcode works in plain English.
-The explanations are not rigorous, and are not in a logical formalism that allows mathematical reasoning.
-There are many situations where descriptions are vague and can be interpreted in more than one way.
-For example, the explanation of \texttt{OP\_LEFT} takes an argument to index a string and throws away everything to the right of the index, but it does not specify whether that operation is right-inclusive.
+The Bitcoin wiki Script article contains an explanation of how each opcode works in plain English, and information about the standard script templates that main nodes of the network will accept.
+While it does provide an informal explanation for each opcode (which came in handy when writing the operational semantics), its descriptions are rather unrigourous.
+On a few occasions descriptions are vague and ambiguous.
+For example, the explanation of \texttt{OP\_LEFT} is that it takes an argument to index a string and throws away everything to the right of the index, but it does not specify whether that operation is right-inclusive.
+Furthermore, this description cannot be used as input in automated theorem proving software.
 
-The reference implementation is much more thorough that the wiki but, being C++ code, it is much harder to interpret and is not in a format that allows for logical reasoning.
-If we turn to Forth to find something related, we can find standard specifications like the Forth-83 standard (http://forth.sourceforge.net/standard/fst83/fst83-12.htm).
-However even these specifications are given in plain English.
+The C++ reference implementation, by necessity, completely specifies the semantics of the language.
+However, the implementation also specifies details that are irrelevant to the syntactic correctness or valid/invalid judgement of a Script program.
+For example, the reference implementation defines a series of error codes, one for each possible cause of early termination with invalid state.
+This level of detail is not needed, and only complexifies reasoning about the language in the way we did in the results section.
+Additionally, as is the case with the existing Bitcoin wiki resource, we cannot directly use the language's implementation as axioms and rules to be used in a theorem prover.
+
+Our proposed grammar, in constrast, unambiguously describes the behavior of Script and can (hopefully eventually) serve as basis for further analytic work.
 

### 2017-03-20 paper.tex
Some corrections in related work, added intro paragraph to results section

@@
 
-In this section, we use our grammar
+In this section, we make use of our formal specification to reason about execution time bounds of Script programs and to explain how a flaw in a prior version of the language was fixed.
 
@@
 The Bitcoin wiki Script article contains an explanation of how each opcode works in plain English, and information about the standard script templates that main nodes of the network will accept.
-While it does provide an informal explanation for each opcode (which came in handy when writing the operational semantics), its descriptions are rather unrigourous.
-On a few occasions descriptions are vague and ambiguous.
+While it does provide an informal explanation for each opcode (which came in handy when writing the operational semantics), its descriptions are rather unrigourous, and sometimes even ambiguous.
 For example, the explanation of \texttt{OP\_LEFT} is that it takes an argument to index a string and throws away everything to the right of the index, but it does not specify whether that operation is right-inclusive.
-Furthermore, this description cannot be used as input in automated theorem proving software.
+Finally, this resource cannot be used as input in automated theorem proving software.
 
@@
 For example, the reference implementation defines a series of error codes, one for each possible cause of early termination with invalid state.
-This level of detail is not needed, and only complexifies reasoning about the language in the way we did in the results section.
+These errors all cause the program execution to halt in the same way, and mainly serve to communicate the cause of an early termination to the running environment.
+Taking these various error codes into account needlessly complexifies the type of analyses we presented in our results section.
 Additionally, as is the case with the existing Bitcoin wiki resource, we cannot directly use the language's implementation as axioms and rules to be used in a theorem prover.
 
-Our proposed grammar, in constrast, unambiguously describes the behavior of Script and can (hopefully eventually) serve as basis for further analytic work.
+Our proposed grammar, in constrast, unambiguously describes the behavior of Script, is easier to reason about the C++ code, and can (hopefully eventually) serve as basis for further analytic work.
 

### 2017-03-21 paper.tex
updated \S.II.A Approach -> Definition of Script Grammar

--- 2017-03-20 paper.tex
@@
 \subsection{Definition of Script Grammar}
+The bulk of our project consists in providing a well-formed and consistent definition of the Script language.
+This includes a description of the syntax that the text of scripts must conform to as well as descriptions of the operational semantics by which the scripts evaluate.
+Due to the breadth of this component of our work, we defer exposition of the grammar to [Appendix A].
 
+As the Bitcoin protocol lacks any formal specification by which it is defined, it is well accepted that the reference implementation defines the protocol.
+We therefore turned to the reference implementation to construct our description of Script's grammar---while also heavily upon the [Bitcoin Wiki] and [Mastering Bitcoin] to gain a detailed if not authoritative conceptual understanding of the relevant concepts and constructs.
+Insomuch as our grammar is consistent with this reference, it is correct.
+Insomuch as it diverges it is incorrect.
+Formally proving the consistency or equivalence of our grammar with the reference implementation is outside the scope of this project and is left as future work.
+Indeed, our grammar is purposefully incomplete as discussed below (\S II.C) due to the disproportionate expressive burden imposed by a few op-codes.
 \subsubsection{Syntax}
-\# Bitcoin wiki -> tables describe the format\\
-\# transaction is X Y, inputs are such and such\\
-\# looked at enums in source code to get full list of defined opcode and their hex values\\
-JOSEPH: also explain how the different syntactic groups were broken down\\
+
+Script is a stack-based language, bearing some semblance to Forth.
+It is parsed and evaluated left to right [Bitcoin Wiki/script].
+The syntax of Script is defined by the binary serializations of script that reference implementation understands to be well-formed.
+We discover this by reviewing relevant literature (of which there is not much) and the source code itself, to glean such syntactic structures.
+
+The "text" of a Script program is its binary serialization.
+In most contexts, our syntax (accurately) considers this text to be tokenized into a byte vector.
+Ultimately all syntactic constructs in Script are either commands which alter the state of the program in a defined manner or contextual data which parameterize the behavior of some commands.
+
+We divide commands into "simple-word commands", termed \texttt{scom}, which are single byte commands that take no contextual program text to modify their behavior; "multiple-word commands", termed \texttt{mcom}, which rely upon contextual program text to modify their behavior---such as \texttt{OP\_IF}/\texttt{OP\_ELSE}/\texttt{OP\_ENDIF} or \texttt{OP\_PUSHDATA1}; and disabled commands, termed \texttt{dcom}, which consist of op-codes reserved for future use, disabled to remove perceived attack vectors, or reserved for testing.
+Commands are constructed from op-codes or words which are single byte literals with special meaning along with any relevant contextual data.
+
+We define our syntax sufficiently to be aware of the transactions formats in which scripts are embedded in the hope that it may be used in the future to describe the semantics of commands which require such awareness---\texttt{OP\_CHECKSIG}.
 
@@
 
-We described the big-step semantics and the small-step semantics of the language using a combination of the reference implementation of the language [Ref] and of the Script page of the Bitcoin wiki [Ref].
+We provide both big-step and small-step operational semantics.
+
+The state of the program consists entirely of the stack, $S$, the alt-stack, $AS$, and the validity $V$.
+The stack contains byte-vectors.
+The reference implementation defines the encoding used to map from byte-vectors to numbers.
+We take the existence of an inference rule describing said encoding for granted.
+
+We describe the semantic meaning of each command with one or more corresponding inference rules.
+In order to contain the size of our description, we provide that well-formed programs with no corresponding inference rule for the execution of a command in a given state transition to the state in which the script is determined to be invalid and execution halts.
+
+A number of op-codes have been disabled due to perceived vulnerabilities, reserved for future use, or reserved for internal use in testing.
+We provide inference rules for disabled op-codes, but also include them in \texttt{dcom}.
+We provide a contradictory inference rule for \texttt{dcom} which causes relevant commands to unequivocally evaluate to an invalid state.
+We provide these to allow for the modeling of the behavior of Script before these were disabled.
+If they are to be used, these op-codes must not be considered to be in \texttt{dcom}.
+
+Scripts exist in transactions and op-codes exist which directly reference their transaction context. We omit these because of the disproportionate descriptive burden they introduce.
+
+We also provide small step semantics to describe the progression of evaluation of a script program. The local and global reduction rules are compact due to the language's inherent simplicity.
 
@@
 
-To our knowledge, no previous effort was made to provide a formal description of Script.
-The Bitcoin developers community relies mostly on a combination of the Bitcoin wiki Script article [2] and Script's C++ reference implementation [1] to learn the language or reason about it.
+To our knowledge, there has been no previous effort to provide a formal specification of the Script language.
+As a result, the existing documentation works as a rough guideline for our work but is not complete.
+Explanations of how Script works are located on the Bitcoin wiki [2] and the reference implementation [1].
 
-The Bitcoin wiki Script article contains an explanation of how each opcode works in plain English, and information about the standard script templates that main nodes of the network will accept.
-While it does provide an informal explanation for each opcode (which came in handy when writing the operational semantics), its descriptions are rather unrigourous.
-On a few occasions descriptions are vague and ambiguous.
-For example, the explanation of \texttt{OP\_LEFT} is that it takes an argument to index a string and throws away everything to the right of the index, but it does not specify whether that operation is right-inclusive.
-Furthermore, this description cannot be used as input in automated theorem proving software.
+The Bitcoin wiki is an explanation of how each opcode works in plain English.
+The explanations are not rigorous, and are not in a logical formalism that allows mathematical reasoning.
+There are many situations where descriptions are vague and can be interpreted in more than one way.
+For example, the explanation of \texttt{OP\_LEFT} takes an argument to index a string and throws away everything to the right of the index, but it does not specify whether that operation is right-inclusive.
 
-The C++ reference implementation, by necessity, completely specifies the semantics of the language.
-However, the implementation also specifies details that are irrelevant to the syntactic correctness or valid/invalid judgement of a Script program.
-For example, the reference implementation defines a series of error codes, one for each possible cause of early termination with invalid state.
-This level of detail is not needed, and only complexifies reasoning about the language in the way we did in the results section.
-Additionally, as is the case with the existing Bitcoin wiki resource, we cannot directly use the language's implementation as axioms and rules to be used in a theorem prover.
-
-Our proposed grammar, in constrast, unambiguously describes the behavior of Script and can (hopefully eventually) serve as basis for further analytic work.
+The reference implementation is much more thorough that the wiki but, being C++ code, it is much harder to interpret and is not in a format that allows for logical reasoning.
+If we turn to Forth to find something related, we can find standard specifications like the Forth-83 standard (http://forth.sourceforge.net/standard/fst83/fst83-12.htm).
+However even these specifications are given in plain English.
 

### 2017-03-21 paper.tex

@@
 
-To our knowledge, there has been no previous effort to provide a formal specification of the Script language.
-As a result, the existing documentation works as a rough guideline for our work but is not complete.
-Explanations of how Script works are located on the Bitcoin wiki [2] and the reference implementation [1].
+To our knowledge, no previous effort was made to provide a formal description of Script.
+The Bitcoin developers community relies mostly on a combination of the Bitcoin wiki Script article [2] and Script's C++ reference implementation [1] to learn the language or reason about it.
 
-The Bitcoin wiki is an explanation of how each opcode works in plain English.
-The explanations are not rigorous, and are not in a logical formalism that allows mathematical reasoning.
-There are many situations where descriptions are vague and can be interpreted in more than one way.
-For example, the explanation of \texttt{OP\_LEFT} takes an argument to index a string and throws away everything to the right of the index, but it does not specify whether that operation is right-inclusive.
+The Bitcoin wiki Script article contains an explanation of how each opcode works in plain English, and information about the standard script templates that main nodes of the network will accept.
+While it does provide an informal explanation for each opcode (which came in handy when writing the operational semantics), its descriptions are rather unrigourous, and sometimes even ambiguous.
+For example, the explanation of \texttt{OP\_LEFT} is that it takes an argument to index a string and throws away everything to the right of the index, but it does not specify whether that operation is right-inclusive.
+Finally, this resource cannot be used as input in automated theorem proving software.
 
-The reference implementation is much more thorough that the wiki but, being C++ code, it is much harder to interpret and is not in a format that allows for logical reasoning.
-If we turn to Forth to find something related, we can find standard specifications like the Forth-83 standard (http://forth.sourceforge.net/standard/fst83/fst83-12.htm).
-However even these specifications are given in plain English.
+The C++ reference implementation, by necessity, completely specifies the semantics of the language.
+However, the implementation also specifies details that are irrelevant to the syntactic correctness or valid/invalid judgement of a Script program.
+For example, the reference implementation defines a series of error codes, one for each possible cause of early termination with invalid state.
+These errors all cause the program execution to halt in the same way, and mainly serve to communicate the cause of an early termination to the running environment.
+Taking these various error codes into account needlessly complexifies the type of analyses we presented in our results section.
+Additionally, as is the case with the existing Bitcoin wiki resource, we cannot directly use the language's implementation as axioms and rules to be used in a theorem prover.
+
+Our proposed grammar, in constrast, unambiguously describes the behavior of Script, is easier to reason about the C++ code, and can (hopefully eventually) serve as basis for further analytic work.
 

### 2017-03-21 paper.tex
Fixed typo

@@
 
-Our proposed grammar, in constrast, unambiguously describes the behavior of Script, is easier to reason about the C++ code, and can (hopefully eventually) serve as basis for further analytic work.
+Our proposed grammar, in constrast, unambiguously describes the behavior of Script, is easier to reason about than the C++ code, and can (hopefully eventually) serve as basis for further analytic work.
 

### 2017-03-21 paper.tex
Added missing word

@@
 As the Bitcoin protocol lacks any formal specification by which it is defined, it is well accepted that the reference implementation defines the protocol.
-We therefore turned to the reference implementation to construct our description of Script's grammar---while also heavily upon the [Bitcoin Wiki] and [Mastering Bitcoin] to gain a detailed if not authoritative conceptual understanding of the relevant concepts and constructs.
+We therefore turned to the reference implementation to construct our description of Script's grammar---while also relying heavily upon the [Bitcoin Wiki] and [Mastering Bitcoin] to gain a detailed if not authoritative conceptual understanding of the relevant concepts and constructs.
 Insomuch as our grammar is consistent with this reference, it is correct.

### 2017-03-21 paper.tex
Removed last blue comment
Added conclusion first draft.

@@
 As Bitcoin's economic presence grows, formal specifications, abstracted away from current implementations of the Bitcoin protocol, will become increasingly essential in mediating conflict over what Bitcoin is or should be, allowing participants in the Bitcoin community to reason formally and precisely about their opinions and proposals.
-
-\textcolor{blue}{[anything else? can we add more examples of how a lack of formal specification was problematic?]}
 
@@
 
+In this research project we propose the first formal description of Bitcoin's Script language.
+We describe Script's syntax and operational semantics, both big-step and small-step.
+We report difficulties encountered in the process, such as how to model byte vector representation in a compact and expressive notation, and suggest solutions to these problems.
+We then exemplify how this grammar can be used to reason about the language and its use, by showing an analysis of execution time bounds and describing how past operational semantics were flawed.
+
+One of the main advantages of formalizing the description of a language is that it allows automated formal reasoning about it.
+A logical next step to this project would be to adapt our description to make it usable by a particular formal verification engine.
+For example, it would be useful to show that certain undesirable states are unreachable using a specific subset of the language.
+Many Bitcoin developers push for changes and extensions to the capabilities of the language or to the accepted standard programs over the network.
+Others, for fear of introducing vulnerabilities in the network, argue that the community should be satisfied with the current state of affairs.
+Having a robust and automated tool to prove whether a change to the language caused unintended consequences would prove useful in mediating these discussions.
+In any case, we feel that continuing to develop and refine a formal specification of Script should help the community attain a better understanding of a language on which it relies so heavily.
+
 \addtolength{\textheight}{-12cm}   % This command serves to balance the column lengths

### 2017-03-22 paper.tex
revised Intrdouction to paper

--- 2017-03-21 paper.tex
@@
 
-Script is what Bitcoin uses to encode transactions over the network.
-It is purposefully restricted, and most notably lacks a mechanism for looping and guarantees termination.
-To properly learn the language and understand its semantics, one must read through the catalog of \textit{opcodes} (hex values used as operators) and analyze its C++ reference implementation.
+Bitcoin is a fully decentralized payment system, the first practical decentralized solution to "the double spending problem." Bitcoin has in the years since its introduction grown to prominence, with dozens if not hundreds of crypto-currencies following in its wake.
+Bitcoin is generally the simplest and most-established of these, serving as the anchor currency, against which other crypto-currencies are valued.
 
-As a postfix language written in sequences of bytes, Script is not an easy language to pick up and understand.
-It is even harder to assess the effect that proposed modifications to the language might have on the guarantees that it must provide.
+There has been a great deal of research into different aspects of Bitcoin.
+One particularly central aspect has, however, been under-examined.
+The transactions so secured by the Bitcoin network are encoded in Script, a programming language tasked with determining the validity of a transaction.
+
+Script, as with the the entirety of the Bitcoin protocol, lacks a formal specification.
+Rather, it is defined by the grammar and serialization implied by the reference implementation of Bitcoin.
+We set out to provide a formal description of the Script language and to demonstrate, with somewhat simple examples, how such a description might be useful in analyzing the properties of the language.
+
+Script is purposefully restricted, most notably lacking a mechanism for looping or recursion, guaranteeing termination.
+A a stack-based language written with byte literals as program text and no formal specification to turn to, Script is not an easy language to pick up and understand.
+It is a similarly difficult task to analyze it's behavior without a formal grammar.
+
 In the past, Script has been progressively restricted in response to discovered vulnerabilities in the language.
-The restrictions eliminated the vulnerabilities, at the cost of reducing the expressiveness of the language.
-It is not clear that all of the restrictions added were necessary to produce the desired end of securing the language.
-A formal description and eventually specification would help in such determinations, allowing more objective and precise decisions.
+The restrictions eliminated such vulnerabilities at the cost of reducing the expressiveness of the language.
+It is not clear that all of the restrictions added were entirely necessary to secure the language.
+
+This is complicated by the lack of any sufficiently precise definition of security.
+It is our hope that a formal description and eventually specification of Script will aid the construction of framework for such determinations, allowing more objective and precise analysis of both current and proposed future forms of the language.
 
@@
 
-Our proposed grammar, in constrast, unambiguously describes the behavior of Script, is easier to reason about than the C++ code, and can (hopefully eventually) serve as basis for further analytic work.
+Our proposed grammar, in constrast, unambiguously describes the behavior of Script, is easier to reason about the C++ code, and can (hopefully eventually) serve as basis for further analytic work.
 

### 2017-03-22 paper.tex

@@
 
-Our proposed grammar, in constrast, unambiguously describes the behavior of Script, is easier to reason about the C++ code, and can (hopefully eventually) serve as basis for further analytic work.
+Our proposed grammar, in constrast, unambiguously describes the behavior of Script, is easier to reason about than the C++ code, and can (hopefully eventually) serve as basis for further analytic work.
 
 \section{CONCLUSIONS}
+
+In this research project we propose the first formal description of Bitcoin's Script language.
+We describe Script's syntax and operational semantics, both big-step and small-step.
+We report difficulties encountered in the process, such as how to model byte vector representation in a compact and expressive notation, and suggest solutions to these problems.
+We then exemplify how this grammar can be used to reason about the language and its use, by showing an analysis of execution time bounds and describing how past operational semantics were flawed.
+
+One of the main advantages of formalizing the description of a language is that it allows automated formal reasoning about it.
+A logical next step to this project would be to adapt our description to make it usable by a particular formal verification engine.
+For example, it would be useful to show that certain undesirable states are unreachable using a specific subset of the language.
+Many Bitcoin developers push for changes and extensions to the capabilities of the language or to the accepted standard programs over the network.
+Others, for fear of introducing vulnerabilities in the network, argue that the community should be satisfied with the current state of affairs.
+Having a robust and automated tool to prove whether a change to the language caused unintended consequences would prove useful in mediating these discussions.
+In any case, we feel that continuing to develop and refine a formal specification of Script should help the community attain a better understanding of a language on which it relies so heavily.
 

### 2017-03-23 paper.tex

@@
 
-\subsection{Application of Script Grammar}
-
-[Describe how grammar was used]
-
 \subsection{Difficulties encountered}
@@
 The reference documentation itself only documents changes back to 2014 while most of the serious problems with Script were fixed in 2010.
+Furthermore, whenever a weakness was discovered by the Bitcoin developers themselves, they did not tell everyone exactly how to exploit the language.
+Thus it is known that there was a vulnerability involving the \texttt{OP\_LEFT} command, and that it was disabled as a result, but to our knowledge there is no information online about what the problem was exactly.
 The only fix we were able to analyze was the change to \texttt{OP\_RETURN}.
@@
 
-In this section, we make use of our formal specification to reason about execution time bounds of Script programs and to explain how a flaw in a prior version of the language was fixed.
+In this section, we make use of our formal description to reason about execution time bounds of Script programs and to explain how a flaw in a prior version of the language was fixed.
 

### 2017-03-23 APPENDIX.tex
Added appendix page

--- 2017-03-19 scriptGrammar.tex
@@
 
+\pagenumbering{gobble}
+
 \begin{document}
-
-\IfFileExists{main.pdf}{\includepdf{main.pdf}}{}
-
-\title{Script Grammar}
+\title{APPENDIX}
 \date{}
 \maketitle
-
-\section{Intdoduction}
-
-Script is a constrained stack-based language inspired by Forth.
-
-Script programs (scripts) exist within Bitcoin transactions, which eventually are finalized or irrevocably committed to the global state of the Bitcoin network by their inclusion in a block which exists in the block-chain with the most proof of work expended.
-
-Transactions consist of inputs and outputs, both of which include scripts.
-A transaction input script \textit{unlocks} bitcoins; a transaction output script \textit{locks} bitcoins.
-An input references the output of a previous transaction and provides a script which yields $\sigma_0$ when run.
-In order for the input to successfully unlock the bitcoins locked up in the referenced output, said output's script must yield $\sigma_1 = \sigma[V = valid]$ when run with it's state initialized to the state $\sigma_0$--yielded by the unlocking/input/spending script.
-
-Script program syntax is defined in terms of byte-commands.
-Each command is one byte long.
-Not all bytes are defined as valid commands and many bytes which are defined have syntactic restrictions on the following program bytes which must hold in order for the program to be well-formed.
-Script defines 173 such commands, each corresponding to a particular byte, also called \textit{words} or \textit{opcodes}.
-For example, the command that adds 1 to the value on top of the stack, is written \texttt{0x8b}.
-These opcodes have human-readable aliases to make description of the language easier.
-For example, the opcode \texttt{0x8b} is referred to as \texttt{OP\_1ADD} in the reference implementation's source code and documentation.
-
-Script is intentionally constrainted: There language does not allow for the expression of loops or recursion.
-This is desirable, guaranteeing that the runtime of a Script program is linear with the size of the program.
-
-Script is defined by the specification provided by the reference implementation. 
-Script is  as sequences of bytecodes.
-In describing its semantics, we will generally not refer to the format of the values, but some commands in the language only have meaning if we take into account the fact that any part of a Script program is a byte vector.
-For this reason, we define byte vectors as the following, and will reserve \texttt{b} and \texttt{B} to express a byte and a byte vector, respectively.
-
-\hyperlink{OPSEM}{Section 2} presents a consistent and exhaustive BNF describing Script's syntax.
-
-\hyperlink{OPSEM}{Section 3} specifies the operational semantics of the syntactic constructs defined in \hyperlink{OPSEM}{Section 2}.
-\section{Syntax}
-
-\subsection{Fundamental data types}
-\noindent
-\texttt{b ::= 0x00 | 0x01 | ... | 0xFF} \\
-\texttt{B ::= b | b B} \\
-\texttt{b$_i$ ::= b} \\
-\texttt{B$^n$ ::= b$_1$ b$_2$ ... b$_n$} \\
-\texttt{varint ::= }
-
-\subsection{Transactions}
-\noindent
-\texttt{tx ::= tx\_version tx\_numin tx\_inputs tx\_numout tx\_outputs tx\_locktime  } \\
-\texttt{tx\_version ::= B$^4$  } \\
-\texttt{tx\_numin ::= varint  } \\
-\texttt{tx\_inputs ::= txin | txin tx\_inputs   } \\
-\texttt{tx\_numout ::= varint  } \\
-\texttt{tx\_outputs ::= txout | txout tx\_outputs  } \\
-\texttt{tx\_locktime ::= B$^4$  }
-
-\subsubsection{Transaction Inputs}
-\noindent
-\texttt{txin ::= txin\_prevhash txin\_index txin\_scriptlen txin\_script txin\_seqno} \\
-\texttt{txin\_prevhash ::= B$^{256}$} \\
-\texttt{txin\_index ::= B$^4$} \\
-\texttt{txin\_scriptlen ::= varint} \\
-\texttt{txin\_script ::= script} \\
-\texttt{txin\_seqno ::= B$^4$}
-
-\subsubsection{Transaction Outputs}
-\noindent
-\texttt{txout ::= txout\_value txout\_scriptlen txout\_script  } \\
-\texttt{txout\_value ::= B$^8$} \\
-\texttt{txout\_scriptlen ::= varint} \\
-\texttt{txout\_script ::= script}
-
-\subsection{Script}
-\noindent
-\texttt{script ::= com | com script} \\
-\texttt{com ::= scom | mcom}
-
-\subsubsection{Single-Word Commands}
-\noindent \hangindent=0.7cm
-\texttt{scom ::=
-  scom\_push | scom\_control | scom\_stack | scom\_splice | scom\_bitlogic |
-  scom\_numeric | scom\_crypto | scom\_expansion | scom\_template} \\
-
-\noindent \hangindent=0.7cm
-\texttt{scom\_push ::=
-  OP\_0 | OP\_FALSE | OP\_1NEGATE | \sout{OP\_RESERVED} | OP\_1 | OP\_TRUE |
-  OP\_2 | OP\_3 | OP\_4 | OP\_5 | OP\_6 | OP\_7 | OP\_8 | OP\_9 | OP\_10 |
-  OP\_11 | OP\_12 | OP\_13 | OP\_14 | OP\_15 | OP\_16} \\
-
-\noindent \hangindent=0.7cm
-\texttt{scom\_control ::=
- OP\_NOP | \sout{OP\_VER} | \sout{OP\_VERIF} | \sout{OP\_VERNOTIF} | OP\_VERIFY
- | OP\_RETURN} \\
-
-\noindent \hangindent=0.7cm
-\texttt{scom\_stack ::=
-  OP\_TOALTSTACK | OP\_FROMALTSTACK | OP\_2DROP | OP\_2DUP | OP\_3DUP |
-  OP\_2OVER | OP\_2ROT | OP\_2SWAP | OP\_IFDUP | OP\_DEPTH | OP\_DROP | OP\_DUP
-  | OP\_NIP | OP\_OVER | OP\_PICK | OP\_ROLL | OP\_ROT | OP\_SWAP | OP\_TUCK}
-  \\
-
-\noindent \hangindent=0.7cm
-\texttt{scom\_splice ::=
-  \sout{OP\_CAT }| \sout{OP\_SUBSTR} | \sout{OP\_LEFT} | \sout{OP\_RIGHT} |
-  OP\_SIZE} \\
-
-\noindent \hangindent=0.7cm
-\texttt{scom\_bitlogic ::=
-  \sout{OP\_INVERT} | \sout{OP\_AND} | \sout{OP\_OR} | \sout{OP\_XOR} |
-  OP\_EQUAL | OP\_EQUALVERIFY | \sout{OP\_RESERVED1} | \sout{OP\_RESERVED2}} \\
-
-\noindent \hangindent=0.7cm
-\texttt{scom\_numeric ::=
-  OP\_1ADD | OP\_1SUB | \sout{OP\_2MUL} | \sout{OP\_2DIV} | OP\_NEGATE |
-  OP\_ABS | OP\_NOT | OP\_0NOTEQUAL | OP\_ADD | OP\_SUB | \sout{OP\_MUL} |
-  \sout{OP\_DIV} | \sout{OP\_MOD} | \sout{OP\_LSHIFT} | \sout{OP\_RSHIFT} |
-  OP\_BOOLAND | OP\_BOOLOR | OP\_NUMEQUAL | OP\_NUMEQUALVERIFY |
-  OP\_NUMNOTEQUAL | OP\_LESSTHAN | OP\_GREATERTHAN | OP\_LESSTHANOREQUAL |
-  OP\_GREATERTHANOREQUAL | OP\_MIN | OP\_MAX | OP\_WITHIN} \\
-
-\noindent \hangindent=0.7cm
-\texttt{scom\_crypto ::=
-  OP\_RIPEMD160 | OP\_SHA1 | OP\_SHA256 | OP\_HASH160 | OP\_HASH256 |
-  OP\_CODESEPARATOR | OP\_CHECKSIG | OP\_CHECKSIGVERIFY | OP\_CHECKMULTISIG |
-  OP\_CHECKMULTISIGVERIFY} \\
-
-\noindent \hangindent=0.7cm
-\texttt{scom\_expansion ::=
-  \sout{OP\_NOP1} | OP\_CHECKLOCKTIMEVERIFY | OP\_NOP2 |
-  OP\_CHECKSEQUENCEVERIFY | OP\_NOP3 | \sout{OP\_NOP4} | \sout{OP\_NOP5} |
-  \sout{OP\_NOP6} | \sout{OP\_NOP7} | \sout{OP\_NOP8} | \sout{OP\_NOP9} |
-  \sout{OP\_NOP10}} \\
-
-\noindent \hangindent=0.7cm
-\texttt{scom\_template ::=
-  \sout{OP\_SMALLINTEGER} | \sout{OP\_PUBKEYS} | \sout{OP\_PUBKEYHASH} |
-  \sout{OP\_PUBKEY} | \sout{OP\_INVALIDOPCODE}}
-
-\subsubsection{Multiple-Word Commands}
-\noindent \hangindent=0.7cm
-\texttt{mcom ::= mcom\_if | mcom\_notif | mcom\_push} \\
-
-% compund control commands
-\noindent
-\texttt{mcom\_if ::= OP\_IF script OP\_ENDIF | OP\_IF script mcom\_else OP\_ENDIF} \\
-\texttt{mcom\_notif ::= OP\_NOTIF script OP\_ENDIF | OP\_NOTIF script mcom\_else OP\_ENDIF} \\
-\texttt{mcom\_else ::= OP\_ELSE script | OP\_ELSE script mcom\_else } \\
-
-% compound push commands
-\noindent \hangindent=0.7cm
-\texttt{mcom\_push ::=
-  OP\_PUSHBYTES$_N$ B$^N$ | OP\_PUSHDATA1 B$^1$ B |OP\_PUSHDATA2 B$^2$ B
-  | OP\_PUSHDATA4 B$^4$ B}
-
-\subsubsection{Disabled Commands}
-\noindent \hangindent=0.7cm
-\texttt{dcom ::=
-  dcom\_push | dcom\_control | dcom\_stack | dcom\_splice | dcom\_bitlogic |
-  dcom\_numeric | dcom\_crypto | dcom\_expansion | dcom\_template} \\
-
-\noindent \hangindent=0.7cm
-\texttt{dcom\_push ::= \sout{OP\_RESERVED}} \\
-
-\noindent \hangindent=0.7cm
-\texttt{scom\_control ::=
- \sout{OP\_VER} | \sout{OP\_VERIF} | \sout{OP\_VERNOTIF}} \\
-
-\noindent \hangindent=0.7cm
-\texttt{dcom\_stack ::= } \\
-
-\noindent \hangindent=0.7cm
-\texttt{dcom\_splice ::=
-  \sout{OP\_CAT }| \sout{OP\_SUBSTR} | \sout{OP\_LEFT} | \sout{OP\_RIGHT}
-  } \\
-
-\noindent \hangindent=0.7cm
-\texttt{dcom\_bitlogic ::=
-  \sout{OP\_INVERT} | \sout{OP\_AND} | \sout{OP\_OR} | \sout{OP\_XOR} |
-  \sout{OP\_RESERVED1} | \sout{OP\_RESERVED2}} \\
-
-\noindent \hangindent=0.7cm
-\texttt{dcom\_numeric ::=
-  \sout{OP\_2MUL} | \sout{OP\_2DIV} | \sout{OP\_MUL} |
-  \sout{OP\_DIV} | \sout{OP\_MOD} | \sout{OP\_LSHIFT} | \sout{OP\_RSHIFT}} \\
-
-\noindent \hangindent=0.7cm
-\texttt{dcom\_crypto ::=} \\
-
-\noindent \hangindent=0.7cm
-\texttt{dcom\_expansion ::=
-  \sout{OP\_NOP1} | \sout{OP\_NOP4} | \sout{OP\_NOP5} |
-  \sout{OP\_NOP6} | \sout{OP\_NOP7} | \sout{OP\_NOP8} | \sout{OP\_NOP9} |
-  \sout{OP\_NOP10}} \\
-
-\noindent \hangindent=0.7cm
-\texttt{dcom\_template ::=
-  \sout{OP\_SMALLINTEGER} | \sout{OP\_PUBKEYS} | \sout{OP\_PUBKEYHASH} |
-  \sout{OP\_PUBKEY} | \sout{OP\_INVALIDOPCODE}}
-
-\pagebreak
-
-\subsubsection{OP Codes}
-
-\begin{multicols}{3}
-
-\textbf{\centering push value} \\
-\texttt{OP\_0 ::= 0x00  } \\
-\texttt{OP\_FALSE ::= OP\_0  } \\
-\texttt{{OP\_PUSHBYTES$^N$} ::= 0x01-0x4b} \\
-\texttt{OP\_PUSHDATA1 ::= 0x4c  } \\
-\texttt{OP\_PUSHDATA2 ::= 0x4d  } \\
-\texttt{OP\_PUSHDATA4 ::= 0x4e  } \\
-\texttt{OP\_1NEGATE ::= 0x4f  } \\
-\texttt{OP\_RESERVED ::= 0x50  } \\
-\texttt{OP\_1 ::= 0x51  } \\
-\texttt{OP\_TRUE::=OP\_1  } \\
-\texttt{OP\_2 ::= 0x52  } \\
-\texttt{OP\_3 ::= 0x53  } \\
-\texttt{OP\_4 ::= 0x54  } \\
-\texttt{OP\_5 ::= 0x55  } \\
-\texttt{OP\_6 ::= 0x56  } \\
-\texttt{OP\_7 ::= 0x57  } \\
-\texttt{OP\_8 ::= 0x58  } \\
-\texttt{OP\_9 ::= 0x59  } \\
-\texttt{OP\_10 ::= 0x5a  } \\
-\texttt{OP\_11 ::= 0x5b  } \\
-\texttt{OP\_12 ::= 0x5c  } \\
-\texttt{OP\_13 ::= 0x5d  } \\
-\texttt{OP\_14 ::= 0x5e  } \\
-\texttt{OP\_15 ::= 0x5f  } \\
-\texttt{OP\_16 ::= 0x60  } \\
-
-\textbf{\centering control} \\
-\texttt{OP\_NOP ::= 0x61  } \\
-\texttt{OP\_VER ::= 0x62  } \\
-\texttt{OP\_IF ::= 0x63  } \\
-\texttt{OP\_NOTIF ::= 0x64  } \\
-\texttt{OP\_VERIF ::= 0x65  } \\
-\texttt{OP\_VERNOTIF ::= 0x66  } \\
-\texttt{OP\_ELSE ::= 0x67  } \\
-\texttt{OP\_ENDIF ::= 0x68  } \\
-\texttt{OP\_VERIFY ::= 0x69  } \\
-\texttt{OP\_RETURN ::= 0x6a  } \\
-
-\textbf{\centering stack ops} \\
-\texttt{OP\_TOALTSTACK ::= 0x6b  } \\
-\texttt{OP\_FROMALTSTACK ::= 0x6c  } \\
-\texttt{OP\_2DROP ::= 0x6d  } \\
-\texttt{OP\_2DUP ::= 0x6e  } \\
-\texttt{OP\_3DUP ::= 0x6f  } \\
-\texttt{OP\_2OVER ::= 0x70  } \\
-\texttt{OP\_2ROT ::= 0x71  } \\
-\texttt{OP\_2SWAP ::= 0x72  } \\
-\texttt{OP\_IFDUP ::= 0x73  } \\
-\texttt{OP\_DEPTH ::= 0x74  } \\
-\texttt{OP\_DROP ::= 0x75  } \\
-\texttt{OP\_DUP ::= 0x76  } \\
-\texttt{OP\_NIP ::= 0x77  } \\
-\texttt{OP\_OVER ::= 0x78  } \\
-\texttt{OP\_PICK ::= 0x79  } \\
-\texttt{OP\_ROLL ::= 0x7a  } \\
-\texttt{OP\_ROT ::= 0x7b  } \\
-\texttt{OP\_SWAP ::= 0x7c  } \\
-\texttt{OP\_TUCK ::= 0x7d  } \\
-
-\textbf{\centering splice ops} \\
-\texttt{OP\_CAT ::= 0x7e  } \\
-\texttt{OP\_SUBSTR ::= 0x7f  } \\
-\texttt{OP\_LEFT ::= 0x80  } \\
-\texttt{OP\_RIGHT ::= 0x81  } \\
-\texttt{OP\_SIZE ::= 0x82  } \\
-
-\textbf{\centering bit logic} \\
-\texttt{OP\_INVERT ::= 0x83  } \\
-\texttt{OP\_AND ::= 0x84  } \\
-\texttt{OP\_OR ::= 0x85  } \\
-\texttt{OP\_XOR ::= 0x86  } \\
-\texttt{OP\_EQUAL ::= 0x87  } \\
-\texttt{OP\_EQUALVERIFY ::= 0x88  } \\
-\texttt{OP\_RESERVED1 ::= 0x89  } \\
-\texttt{OP\_RESERVED2 ::= 0x8a  } \\
-
-\textbf{\centering numeric} \\
-\texttt{OP\_1ADD ::= 0x8b  } \\
-\texttt{OP\_1SUB ::= 0x8c  } \\
-\texttt{OP\_2MUL ::= 0x8d  } \\
-\texttt{OP\_2DIV ::= 0x8e  } \\
-\texttt{OP\_NEGATE ::= 0x8f  } \\
-\texttt{OP\_ABS ::= 0x90  } \\
-\texttt{OP\_NOT ::= 0x91  } \\
-\texttt{OP\_0NOTEQUAL ::= 0x92  } \\
-\texttt{OP\_ADD ::= 0x93  } \\
-\texttt{OP\_SUB ::= 0x94  } \\
-\texttt{OP\_MUL ::= 0x95  } \\
-\texttt{OP\_DIV ::= 0x96  } \\
-\texttt{OP\_MOD ::= 0x97  } \\
-\texttt{OP\_LSHIFT ::= 0x98  } \\
-\texttt{OP\_RSHIFT ::= 0x99  } \\
-\texttt{OP\_BOOLAND ::= 0x9a  } \\
-\texttt{OP\_BOOLOR ::= 0x9b  } \\
-\texttt{OP\_NUMEQUAL ::= 0x9c  } \\
-\texttt{OP\_NUMEQUALVERIFY ::= 0x9d  } \\
-\texttt{OP\_NUMNOTEQUAL ::= 0x9e  } \\
-\texttt{OP\_LESSTHAN ::= 0x9f  } \\
-\texttt{OP\_GREATERTHAN ::= 0xa0  } \\
-\texttt{OP\_LESSTHANOREQUAL ::= 0xa1  } \\
-\texttt{OP\_GREATERTHANOREQUAL ::= 0xa2  } \\
-\texttt{OP\_MIN ::= 0xa3  } \\
-\texttt{OP\_MAX ::= 0xa4  } \\
-\texttt{OP\_WITHIN ::= 0xa5  } \\
-
-\textbf{\centering crypto} \\
-\texttt{OP\_RIPEMD160 ::= 0xa6  } \\
-\texttt{OP\_SHA1 ::= 0xa7  } \\
-\texttt{OP\_SHA256 ::= 0xa8  } \\
-\texttt{OP\_HASH160 ::= 0xa9  } \\
-\texttt{OP\_HASH256 ::= 0xaa  } \\
-\texttt{OP\_CODESEPARATOR ::= 0xab  } \\
-\texttt{OP\_CHECKSIG ::= 0xac  } \\
-\texttt{OP\_CHECKSIGVERIFY ::= 0xad  } \\
-\texttt{OP\_CHECKMULTISIG ::= 0xae  } \\
-\texttt{OP\_CHECKMULTISIGVERIFY ::= 0xaf  } \\
-
-\textbf{\centering expansion} \\
-\texttt{OP\_NOP1 ::= 0xb0  } \\
-\texttt{OP\_CHECKLOCKTIMEVERIFY ::= OP\_NOP2 ::= 0xb1  } \\
-\texttt{OP\_CHECKSEQUENCEVERIFY ::= OP\_NOP3 ::= 0xb2  } \\
-\texttt{OP\_NOP4 ::= 0xb3  } \\
-\texttt{OP\_NOP5 ::= 0xb4  } \\
-\texttt{OP\_NOP6 ::= 0xb5  } \\
-\texttt{OP\_NOP7 ::= 0xb6  } \\
-\texttt{OP\_NOP8 ::= 0xb7  } \\
-\texttt{OP\_NOP9 ::= 0xb8  } \\
-\texttt{OP\_NOP10 ::= 0xb9  } \\
-
-\textbf{\centering template matching params }\\
-\texttt{OP\_SMALLINTEGER ::= 0xfa  } \\
-\texttt{OP\_PUBKEYS ::= 0xfb  } \\
-\texttt{OP\_PUBKEYHASH ::= 0xfd  } \\
-\texttt{OP\_PUBKEY ::= 0xfe  } \\
-\texttt{OP\_INVALIDOPCODE ::= 0xff  } \\
-
-\end{multicols}
-
-\pagebreak[2]
-
-\hypertarget{OPSEM}{\section{Operational Semantics -- Big Step Semantics}}
-
-At any given point of a program's execution, the state is entirely described by the following elements:
-
-\begin{itemize}
-	\item The stack $S$, whose elements $S_i$ for $1 \leq i \leq |S|$ are indexed starting from the bottom;
-    \item The alt-stack $AS$, an auxiliary stack indexed in the same way;
-    \item The validity variable $V$, equal to \texttt{true} by default.
-\end{itemize}
-
-Any state in which $V = \texttt{false}$ causes the program to immediately terminate with a special exception, handled by the environment in which the program was being run.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Utility rules
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\hypertarget{OPGEN}{\subsection{Generic rules}}
-In order to more succinctly define the operational semantics of Script, we define generic rules not tied to any command in the language that express operations common in a stack-based language.
-
-% PUSH
-\subsubsection{PUSH}
-\inferrule
-{
-	|S| = L
-}
-{	
-	<\texttt{PUSH } B, \sigma> \Downarrow \sigma[|S| = L+1, S_{|S|} = B]
-}
-\vspace{3mm}
-
-% DROP
-\subsubsection{DROP}
-\inferrule
-{
-	\sigma(|S|) = L \hspace{3mm} 
-    B \Downarrow n \hspace{3mm} 
-	L \geq n
-}
-{
-	<\texttt{DROP } B, \sigma> \Downarrow \sigma[S_i = S_{i+1} \hspace{3mm} \forall i. \hspace{1mm} n \leq i < L]
-}
-\vspace{3mm}
-
-% PUSH_ALT
-\subsubsection{PUSH\_ALT}
-\inferrule
-{|AS| = L}
-{<\texttt{PUSH\_ALT } B, \sigma> \Downarrow \sigma[|AS| = L+1, AS_{|AS|} = B]}
-\vspace{3mm}
-
-% DROP_ALT
-\subsubsection{DROP\_ALT}
-\inferrule
-{
-	\sigma(|AS|) = L \hspace{3mm} \sigma(|AS|) \geq B
-}
-{
-	<\texttt{DROP\_ALT } B, \sigma> \Downarrow \sigma[AS_i = AS_{i+1} \hspace{3mm} \forall i. \hspace{1mm} B \leq i < L]
-}
-\vspace{3mm}
-
-% TEST
-\subsubsection{TEST}
-\inferrule
-{	
-	\sigma(|S|) \geq x \hspace{3mm}
-	\sigma(S_x) \Downarrow t \hspace{3mm} 
-    t = 0
-}
-{
-	<\texttt{TEST $x$}, \sigma> \Downarrow FALSE
-}
-\vspace{3mm}
-
-\inferrule
-{	
-	\sigma(|S|) \geq x \hspace{3mm}
-	\sigma(S_x) \Downarrow t \hspace{3mm} 
-    t \neq 0
-}
-{
-	<\texttt{TEST $x$}, \sigma> \Downarrow TRUE
-}
-\vspace{3mm}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Constants}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Constants: OP Code Definitions
-
-% OP_0, OP_FALSE
-\subsubsection{OP\_0, OP\_FALSE}
-Push the byte-vector representing 0 onto the stack. \\
-
-\inferrule
-{
-	\sigma(|S|) = L \hspace{3mm} 
-    B \Downarrow 0
-}
-{
-	<\texttt{OP\_0}, \sigma> \Downarrow \sigma[|S|=L+1, S_{L+1}=B]
-}
-
-% OP_N
-\subsubsection{OP\_N, \textit{OP\_PUSHNBYTES}}
-Push the next $N$ bytes onto the stack. \\
-
-\inferrule
-{
-	\sigma(|S|) = L
-}
-{
-	<\texttt{OP\_N $b_1 \ldots b_N$}, \sigma> \Downarrow 
-	\sigma[|S|=L+1, S_{L+1}=<b_1 \ldots b_N>]
-}
-
-% OP_PUSHDATA1
-\subsubsection{OP\_PUSHDATA1}
-The next byte specifies how many bytes to push onto the stack as a byte-vector. \\
-
-\inferrule
-{
-    B_1 \Downarrow k \hspace{3mm} 
-	\sigma(|S|) = L \hspace{3mm} 
-}
-{
-	<\texttt{OP\_PUSHDATA1 $B_1^1$ $B_2^k$}, \sigma> \Downarrow 
-	\sigma[|S|=L+1, S_{L+1}=B_2]
-}
-
-%\vspace{3mm}
-%\inferrule
-%{
-%	x = |$B$| \hspace{3mm}
-%    b_1 \Downarrow k \hspace{3mm} 
-%    x < k 
-%}
-%{	
-%	<\texttt{OP\_PUSHDATA1 $b_1$ $B$ }, \sigma> \Downarrow 
-%	\sigma[V = false]
-%}
-
-% OP_PUSHDATA2
-\subsubsection{OP\_PUSHDATA2}
-The next two bytes specify how many bytes to push onto the stack as a byte-vector. \\
-
-\inferrule
-{
-    B_1 \Downarrow k \hspace{3mm} 
-	\sigma(|S|) = L \hspace{3mm} 
-}
-{
-	<\texttt{OP\_PUSHDATA2 $B_1^2$ $B_2^k$}, \sigma> \Downarrow 
-	\sigma[|S|=L+1, S_{L+1}=B_2]
-}
-
-%\vspace{3mm}
-%\inferrule
-%{
-%	x = |$B$| \hspace{3mm}
-%    (b_1 b_2) \Downarrow k \hspace{3mm} 
-%    x < k 
-%}
-%{	
-%	<\texttt{OP\_PUSHDATA2 $b_1$ $b_2$ $B$ }, \sigma> \Downarrow 
-%	\sigma[V = false]
-%}
-
-% OP_PUSHDATA4
-\subsubsection{OP\_PUSHDATA4}
-The next four bytes specify how many bytes to push onto the stack as a byte-vector. \\
-
-\inferrule
-{
-    B_1 \Downarrow k \hspace{3mm} 
-	\sigma(|S|) = L \hspace{3mm} 
-}
-{
-	<\texttt{OP\_PUSHDATA4 $B_1^4$ $B_2^k$}, \sigma> \Downarrow
-	\sigma[|S|=L+1, S_{L+1}=B_2]
-}
-
-%\vspace{3mm}
-%\inferrule
-%{
-%	x = |$B$| \hspace{3mm}
-%    (b_1 b_2 b_3 b_4) \Downarrow k \hspace{3mm} 
-%    x < k 
-%}
-%{	
-%	<\texttt{OP\_PUSHDATA4 $b_1$ $b_2$ $b_3$ $b_4$ $B$ }, \sigma> \Downarrow 
-%	\sigma[V = false]
-%}
-
-% OP_1NEGATE
-\subsubsection{OP\_1NEGATE}
-Push the byte-vector representing -1 onto the stack.\\
-
-\inferrule
-{
-	\sigma(|S|) = L \hspace{3mm} 
-    B \Downarrow -1
-}
-{<\texttt{OP\_0}, \sigma> \Downarrow 
-	\sigma[|S|=l+1, S_{l+1}=B]}
-
-% OP_1, OP_TRUE
-\subsubsection{OP\_1, OP\_TRUE}
-Push the byte-vector representing 1 onto the stack. \\
-\inferrule
-{\sigma(|S|) = L \hspace{3mm} B \Downarrow 1}
-{<\texttt{OP\_0}, \sigma> \Downarrow 
-	\sigma[|S|=L+1, S_{L+1}=B]}
-
-% OP_2-OP_16
-\subsubsection{OP\_2-OP\_16, \textit{OP\_PUSHN}}
-Push the byte-vector representing the number specified in the word name onto the stack. \\
-
-\inferrule
-{\sigma(|S|) = L \hspace{3mm} B \Downarrow N}
-{<\texttt{OP\_PUSHN}, \sigma> \Downarrow 
-	\sigma[|S|=L+1, S_{L+1}=B]}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Flow Control}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Flow Control: OP Code Definitions
-
-% OP_NOP
-\subsubsection{OP\_NOP}
-Do nothing. \\
-
-\inferrule
-{\quad}
-{
-	<\texttt{OP\_NOP},\sigma> \Downarrow \sigma
-}
-
-% OP_IF, OP_ELSE, OP_ENDIF
-\subsubsection{OP\_IF, OP\_ELSE, OP\_ENDIF}
-Run $C_1$ if top of stack is present and evaluates to true. Otherwise run $C_2$ if stack is present and evaluates to false. \\
-
-\inferrule
-{	
-	\sigma(|S|)=L \hspace{3mm}
-	<\texttt{TEST $L$},\sigma> \Downarrow \texttt{TRUE} \hspace{3mm}
-	<\texttt{DROP $L$},\sigma> \Downarrow \sigma_1 \hspace{3mm}
-    <C_1, \sigma_1> \Downarrow \sigma '
-}
-{	
-	<\texttt{OP\_IF $C_1$ OP\_ENDIF}> \Downarrow \sigma '
-}
-\vspace{3mm}
-
-\inferrule
-{	
-	\sigma(|S|)=L \hspace{3mm}
-	<\texttt{TEST $L$},\sigma> \Downarrow \texttt{FALSE} \hspace{3mm}
-	<\texttt{DROP $L$},\sigma> \Downarrow \sigma ' \hspace{3mm}
-}
-{	
-	<\texttt{OP\_IF $C_1$ OP\_ENDIF}> \Downarrow \sigma '
-}
-\vspace{3mm}
-
-\inferrule
-{	\sigma(|S|)=L \hspace{3mm}
-	<\texttt{TEST $L$},\sigma> \Downarrow \texttt{TRUE} \hspace{3mm}
-	<\texttt{DROP $L$},\sigma> \Downarrow \sigma_1 \hspace{3mm}
-    <C_1, \sigma_1> \Downarrow \sigma '
-}
-{	<\texttt{OP\_IF $C_1$ OP\_ELSE $C_2$ OP\_ENDIF}> \Downarrow \sigma '
-}
-\vspace{3mm}
-
-\inferrule
-{	\sigma(|S|)=L \hspace{3mm}
-	<\texttt{TEST $L$},\sigma> \Downarrow \texttt{FALSE} \hspace{3mm}
-	<\texttt{DROP $L$},\sigma> \Downarrow \sigma_1 \hspace{3mm}
-    <C_2, \sigma_1> \Downarrow \sigma '
-}
-{	<\texttt{OP\_IF $C_1$ OP\_ELSE $C_2$ OP\_ENDIF}> \Downarrow \sigma '
-}
-\vspace{3mm}
-
-% OP_NOTIF, OP_ELSE, OP_ENDIF
-\subsubsection{OP\_NOTIF, OP\_ELSE, OP\_ENDIF}
-Run $C_1$ if top of stack is present and evaluates to false. Otherwise run $C_2$ if stack is present and evaluates to true. \\
-
-\inferrule
-{	\sigma(|S|)=L \hspace{3mm}
-	<\texttt{TEST $L$},\sigma> \Downarrow \texttt{FALSE} \hspace{3mm}
-	<\texttt{DROP $L$},\sigma> \Downarrow \sigma_1 \hspace{3mm}
-    <C_1, \sigma_1> \Downarrow \sigma '
-}
-{	<\texttt{OP\_NOTIF $C_1$ OP\_ENDIF}> \Downarrow \sigma '
-}
-\vspace{3mm}
-
-\inferrule
-{	\sigma(|S|)=L \hspace{3mm}
-	<\texttt{TEST $L$},\sigma> \Downarrow \texttt{TRUE} \hspace{3mm}
-	<\texttt{DROP $L$},\sigma> \Downarrow \sigma ' \hspace{3mm}
-}
-{	<\texttt{OP\_NOTIF $C_1$ OP\_ENDIF}> \Downarrow \sigma '
-}
-\vspace{3mm}
-
-\inferrule
-{	\sigma(|S|)=L \hspace{3mm}
-	<\texttt{TEST $L$},\sigma> \Downarrow \texttt{FALSE} \hspace{3mm}
-	<\texttt{DROP $L$},\sigma> \Downarrow \sigma_1 \hspace{3mm}
-    <C_1, \sigma_1> \Downarrow \sigma '
-}
-{	<\texttt{OP\_NOTIF $C_1$ OP\_ELSE $C_2$ OP\_ENDIF}> \Downarrow \sigma '
-}
-\vspace{3mm}
-
-\inferrule
-{	\sigma(|S|)=L \hspace{3mm}
-	<\texttt{TEST $L$},\sigma> \Downarrow \texttt{TRUE} \hspace{3mm}
-	<\texttt{DROP $L$},\sigma> \Downarrow \sigma_1 \hspace{3mm}
-    <C_2, \sigma_1> \Downarrow \sigma '
-}
-{	<\texttt{OP\_NOTIF $C_1$ OP\_ELSE $C_2$ OP\_ENDIF}> \Downarrow \sigma '
-}
-\vspace{3mm}
-
-% OP_VERIFY
-\subsubsection{OP\_VERIFY := 0x69}
-If top of stack is present and evaluates to true then remove top of stock and mark transaction as valid; otherwise mark transaction as invalid. \\
-
-\inferrule
-{	\sigma(|S|)=L \hspace{3mm}
-	<\texttt{TEST $L$},\sigma> \Downarrow \texttt{TRUE} \hspace{3mm}
-	<\texttt{DROP $L$},\sigma> \Downarrow \sigma ' \hspace{3mm}
-    \sigma'(V)=\texttt{invalid}
-}
-{	<\texttt{OP\_VERIFY}, \sigma> \Downarrow \sigma '
-}
-
-% OP_RETURN
-\subsubsection{OP\_RETURN := 0x6a}
-Unconditionally mark transaction as invalid. \\
-
-\inferrule
-{	\quad
-}
-{	<\texttt{OP\_RETURN}, \sigma> \Downarrow \sigma[V=\texttt{invalid}]
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Stack}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% OP_TOALTSTACK
-\subsubsection{OP\_TOALTSTACK}
-
-\inferrule
-{	
-	|S| = L \hspace{3mm}
-    S_L = x \hspace{3mm}
-    <\texttt{DROP } L, \sigma> \Downarrow \sigma_1
-    <\texttt{PUSHALT } x, \sigma_1> \Downarrow \sigma'
-}
-{
-	<\texttt{OP\_TOALTSTACK}, \sigma > \Downarrow \sigma
-}
-\vspace{3mm}
-
-% OP_FROMALTSTACK
-\subsubsection{OP\_FROMALTSTACK}
-
-\inferrule
-{	
-	|AS| = L \hspace{3mm}
-    AS_L = x \hspace{3mm}
-    <\texttt{DROPALT } L, \sigma> \Downarrow \sigma_1
-    <\texttt{PUSH } x, \sigma_1> \Downarrow \sigma'
-}
-{
-	<\texttt{OP\_FROMALTSTACK}, \sigma > \Downarrow \sigma
-}
-\vspace{3mm}
-
-%  OP_IFDUP
-\subsubsection{OP\_IFDUP}
-If top of stack is present and non-zero, duplicate the top of the stack. \\
-
-\inferrule
-{
-	\sigma(S_{|S|}) = 0
-}
-{	
-	<\texttt{OP\_IFDUP}, \sigma > \Downarrow \sigma
-}
-\vspace{3mm}
-
-\inferrule
-{
-	\sigma(S_{|S|}) \neq 0 \hspace{3mm}
-    <\texttt{PUSH } S_{|S|} \Downarrow \sigma ' >
-} 
-{
-	<\texttt{OP\_IFDUP}, \sigma > \Downarrow \sigma'
-}
-\vspace{3mm}
-
-% OP_DEPTH
-\subsubsection{OP\_DEPTH}
-Push the byte-vector representing the depth of the stack onto the stack. \\
-
-\inferrule
-{
-	\sigma(|S|) = L \hspace{3mm} 
-    < \texttt{PUSH } L, \sigma > \Downarrow \sigma '
-}
-{
-	<\texttt{OP\_DEPTH}, \sigma > \Downarrow \sigma '
-}
-\vspace{3mm}
-
-% OP_DROP
-\subsubsection{OP\_DROP}
-Remove the top of the stack. \\
-
-\inferrule
-{
-	\sigma(|S|) = L \hspace{3mm} 
-    L > 0 \hspace{3mm}
-    <\texttt{DROP } L, \sigma > \Downarrow \sigma'
-}
-{	
-	<\texttt{OP\_DROP}, \sigma > \Downarrow \sigma'
-}
-\vspace{3mm}
-
-% OP_DUP
-\subsubsection{OP\_DUP}
-Duplicate the top of the stack. \\
-
-\inferrule
-{
-	\sigma(|S|) = L \hspace{3mm} 
-    L > 0 \hspace{3mm}
-    <\texttt{PUSH }S_L, \sigma > \Downarrow \sigma'
-}
-{
-	<\texttt{OP\_DUP}, \sigma > \Downarrow \sigma'
-}
-\vspace{3mm}
-
-% OP_NIP
-\subsubsection{OP\_NIP}
-Remove the byte-vector second from the top of the stack. \\
-
-\inferrule
-{
-	\sigma(|S|) = L \hspace{3mm} 
-    L > 1 \hspace{3mm} 
-    <\texttt{DROP } L-1, \sigma > \Downarrow \sigma'
-}
-{
-	<\texttt{OP\_NIP}, \sigma > \Downarrow \sigma'
-}
-\vspace{3mm}
-
-% OP_OVER
-\subsubsection{OP\_OVER}
-Copy the byte-vector second from the top onto the top of the stack. \\
-
-\inferrule
-{
-	\sigma(|S|) = L \hspace{3mm} 
-    L > 1 \hspace{3mm} 
-    <\texttt{PUSH } S_{L-1}, \sigma > \Downarrow \sigma'
-}
-{
-	<\texttt{OP\_OVER}, \sigma > \Downarrow \sigma'
-}
-\vspace{3mm}
-
-% OP_PICK
-\subsubsection{OP\_PICK}
-Copy the byte-vector $n$ from the top onto the top of the stack, not counting the top element $n$. \\
-
-\inferrule
-{
-	\sigma(|S|) = L \hspace{3mm}
-    \sigma(S_L) = n \hspace{3mm}
-    <\texttt{DROP } L, \sigma> \Downarrow \sigma_1 \hspace{3mm}  
-    <\texttt{PUSH } S_{L-n-1}, \sigma_1> \Downarrow \sigma'
-}
-{
-	<\texttt{OP\_PICK}, \sigma > \Downarrow \sigma'
-}
-\vspace{3mm}
-
-% OP_ROLL
-\subsubsection{OP\_ROLL}
-Move the byte-vector $n$ from the top onto the top of the stack, not counting the top element $n$. \\
-
-\inferrule
-{
-	\sigma(|S|) = L \hspace{3mm}
-    \sigma(S_L) = n \hspace{3mm}
-    <\texttt{DROP } L, \sigma> \Downarrow \sigma_1 \hspace{3mm}  
-    <\texttt{PUSH } S_{L-n-1}, \sigma_1> \Downarrow \sigma_2
-    <\texttt{DROP } L-n-1, \sigma_2> \Downarrow \sigma'
-}
-{
-	<\texttt{OP\_ROLL}, \sigma > \Downarrow \sigma'
-}
-\vspace{3mm}
-
-% OP_ROT
-\subsubsection{OP\_ROT}
-The top three items on the stack are rotated. Equivalently, move the byte-vector third from the top to the top of the stack. \\
-
-\inferrule
-{
-	\sigma(|S|) = L \hspace{3mm}
-    <\texttt{PUSH } S_{L-2}, \sigma_1> \Downarrow \sigma_2
-    <\texttt{DROP } L-2, \sigma_2> \Downarrow \sigma'
-}
-{
-	<\texttt{OP\_ROT}, \sigma > \Downarrow \sigma'
-}
-\vspace{3mm}
-
-% OP_SWAP
-\subsubsection{OP\_SWAP}
-The top two items on the stack are swapped. \\
-
-\inferrule
-{
-	\sigma{|S|} = L \hspace{3mm}
-    <\texttt{PUSH } S_{L-1}, \sigma_1> \Downarrow \sigma_2
-    <\texttt{DROP } {L-1}, \sigma_2> \Downarrow \sigma'
-}
-{<\texttt{OP\_SWAP}, \sigma > \Downarrow \sigma'}
-\vspace{3mm}
-
-% OP_TUCK
-\subsubsection{OP\_TUCK}
-The item at the top of the stack is copied and inserted before the second-to-top item. \\
-
-\inferrule
-{	
-	\sigma(S_{|S|}) = x \hspace{3mm} 
-	\sigma(S_{|S|-1}) = y \hspace{3mm} 
-	<\texttt{DROP } |S|, \sigma> \Downarrow \sigma_1 \hspace{3mm}
-	<\texttt{DROP } |S|, \sigma_1> \Downarrow \sigma_2 \hspace{3mm} \\
-    <\texttt{PUSH } x, \sigma_2> \Downarrow \sigma_3 \hspace{3mm}
-	<\texttt{PUSH } y, \sigma_3> \Downarrow \sigma_4 \hspace{3mm} 
-    <\texttt{PUSH } x, \sigma_4> \Downarrow \sigma'
-}
-{	
-	<\texttt{OP\_TUCK}, \sigma > \Downarrow \sigma'
-}
-\vspace{3mm}
-
-% OP_2DROP
-\subsubsection{OP\_2DROP}
-Removes the top two stack items. \\
-
-\inferrule
-{	
-	<\texttt{DROP } |S|, \sigma> \Downarrow \sigma_1 \hspace{3mm}
-	<\texttt{DROP } |S|, \sigma_1> \Downarrow \sigma' \hspace{3mm} \\
-}
-{	
-	<\texttt{OP\_2DROP}, \sigma > \Downarrow \sigma'
-}
-\vspace{3mm}
-
-% OP_2DUP
-\subsubsection{OP\_2DUP}
-Duplicates the top two stack items. \\
-
-\inferrule
-{	
-	\sigma(|S|) = L \hspace{3mm}
-	<\texttt{PUSH } S_{L-1}, \sigma> \Downarrow \sigma_1 \hspace{3mm}
-    <\texttt{PUSH } S_L, \sigma_1> \Downarrow \sigma' \hspace{3mm}
-}
-{	
-	<\texttt{OP\_2DUP}, \sigma > \Downarrow \sigma'
-}
-\vspace{3mm}
-
-% OP_3DUP
-\subsubsection{OP\_3DUP}
-Duplicates the top three stack items. \\
-
-\inferrule
-{	
-	\sigma(|S|) = L \hspace{3mm}
-	<\texttt{PUSH } S_{L-2}, \sigma> \Downarrow \sigma_1 \hspace{3mm}
-    <\texttt{PUSH } S_{L-1}, \sigma_1> \Downarrow \sigma_2 \hspace{3mm}
-    <\texttt{PUSH } S_L, \sigma_2> \Downarrow \sigma' \hspace{3mm}
-}
-{	
-	<\texttt{OP\_3DUP}, \sigma > \Downarrow \sigma'
-}
-\vspace{3mm}
-
-% OP_2OVER
-\subsubsection{OP\_2OVER}
-Copies the pair of items two spaces back in the stack to the front. \\
-
-\inferrule
-{	
-	\sigma(|S|) = L \hspace{3mm}
-	<\texttt{PUSH } S_{L-3}, \sigma> \Downarrow \sigma_1 \hspace{3mm}
-    <\texttt{PUSH } S_{L-2}, \sigma_1> \Downarrow \sigma' \hspace{3mm}
-}
-{	
-	<\texttt{OP\_2OVER}, \sigma > \Downarrow \sigma'
-}
-\vspace{3mm}
-
-% OP_2ROT
-\subsubsection{OP\_2ROT}
-The fifth and sixth items back are moved to the top of the stack. \\
-
-\inferrule
-{	
-	\sigma(|S|) = L \hspace{3mm}
-    <\texttt{PUSH } S_{L-5}, \sigma_2> \Downarrow \sigma_3 \hspace{3mm}
-    <\texttt{PUSH } S_{L-4}, \sigma_3> \Downarrow \sigma' \hspace{3mm}
-    <\texttt{DROP } L-5, \sigma> \Downarrow \sigma_1 \hspace{3mm}
-    <\texttt{DROP } L-4, \sigma_1> \Downarrow \sigma_2 \hspace{3mm}
-}
-{	
-	<\texttt{OP\_2ROT}, \sigma > \Downarrow \sigma'
-}
-\vspace{3mm}
-
-% OP_2SWAP
-\subsubsection{OP\_2SWAP}
-The first and second items are swapped. The third and fourth items are swapped. \\
-
-\inferrule
-{	
-	\sigma(|S|) = L \hspace{3mm}
-    \sigma(S_{L}) = x \hspace{3mm}
-    \sigma(S_{L-1}) = y \hspace{3mm}
-	<\texttt{DROP } L, \sigma> \Downarrow \sigma_1 \hspace{3mm}
-    <\texttt{DROP } L-1, \sigma_1> \Downarrow \sigma_2 \hspace{3mm} \\
-    <\texttt{PUSH } S_{L-3}, \sigma_2> \Downarrow \sigma_3 \hspace{3mm}
-    <\texttt{DROP } L-3, \sigma_3> \Downarrow \sigma_4 \hspace{3mm}
-    <\texttt{PUSH } y, \sigma_4> \Downarrow \sigma_5 \hspace{3mm}
-    <\texttt{PUSH } x, \sigma_5> \Downarrow \sigma' \hspace{3mm}
-}
-{	
-	<\texttt{OP\_2SWAP}, \sigma > \Downarrow \sigma'
-}
-\vspace{3mm}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Splice}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%OP_CAT
-\subsubsection{OP\_CAT}
-\textcolor{red}{\textit{Disabled}}\\
-Concatenates two byte strings. \\
-
-\inferrule
-{	
-	\sigma(|S|) = L \hspace{3mm}
-    <\texttt{PUSH } S_{L} S_{L-1}, \sigma> \Downarrow \sigma_1 \hspace{3mm}
-    <\texttt{DROP } L, \sigma_1> \Downarrow \sigma_2 \hspace{3mm}
-    <\texttt{DROP } L-1, \sigma_2> \Downarrow \sigma' \hspace{3mm}
-}
-{	
-	<\texttt{OP\_CAT}, \sigma > \Downarrow \sigma'
-}
-\vspace{3mm}
-
-%OP_SUBSTR
-\subsubsection{OP\_SUBSTR}
-\textcolor{red}{\textit{Disabled}}\\
-Returns a subtring of the top element, starting from index given by second element, of size given by third element. \\
-
-\inferrule
-{	
-	\sigma(|S|) = L \hspace{3mm}
-    \sigma(S_L) = b_1 \ldots b_k \hspace{3mm}
-    \sigma(S_{L-1}) = start \hspace{3mm}
-    \sigma(S_{L-2}) = size \hspace{3mm}
-    start < k \hspace{3mm}
-    start + size < k \hspace{3mm} \\
-    <\texttt{PUSH } b_{start} \ldots b_{start+size}, \sigma> \Downarrow \sigma_1 \hspace{3mm}
-    <\texttt{DROP } L, \sigma_1> \Downarrow \sigma_2 \hspace{3mm}
-    <\texttt{DROP } L-1, \sigma_2> \Downarrow \sigma_3 \hspace{3mm}
-    <\texttt{DROP } L-2, \sigma_3> \Downarrow \sigma' \hspace{3mm}
-}
-{	
-	<\texttt{OP\_SUBSTR}, \sigma > \Downarrow \sigma'
-}
-\vspace{3mm}
-
-%OP_LEFT
-\subsubsection{OP\_LEFT}
-\textcolor{red}{\textit{Disabled}}\\
-Returns only character left of a specified point in top element, index given by second element. \\
-
-\inferrule
-{	
-	\sigma(|S|) = L \hspace{3mm}
-    \sigma(S_L) = b_1 \ldots b_k \hspace{3mm}
-    \sigma(S_{L-1}) = idx \hspace{3mm}
-    idx \leq k \hspace{3mm} \\
-    <\texttt{PUSH } b_1 b_{idx-1}, \sigma> \Downarrow \sigma_1 \hspace{3mm}
-    <\texttt{DROP } L, \sigma_1> \Downarrow \sigma_2 \hspace{3mm}
-    <\texttt{DROP } L-1, \sigma_2> \Downarrow \sigma' \hspace{3mm}
-}
-{	
-	<\texttt{OP\_LEFT}, \sigma > \Downarrow \sigma'
-}
-\vspace{3mm}
-
-%OP_RIGHT
-\subsubsection{OP\_RIGHT}
-\textcolor{red}{\textit{Disabled}}\\
-Returns only character right of a specified point in top element, index given by second element. \\
-
-\inferrule
-{	
-	\sigma(|S|) = L \hspace{3mm}
-    \sigma(S_L) = b_1 \ldots b_k \hspace{3mm}
-    \sigma(S_{L-1}) = idx \hspace{3mm}
-    idx \leq k \\
-    <\texttt{PUSH } b_{idx} \ldots b_k, \sigma> \Downarrow \sigma_1 \hspace{3mm}
-    <\texttt{DROP } L, \sigma_1> \Downarrow \sigma_2 \hspace{3mm}
-    <\texttt{DROP } L-1, \sigma_2> \Downarrow \sigma' \hspace{3mm}
-}
-{	
-	<\texttt{OP\_RIGHT}, \sigma > \Downarrow \sigma'
-}
-\vspace{3mm}
-
-\subsubsection{OP\_SIZE}
-Pushes the number of bytes of the top element of the stack. \\
-
-\inferrule
-{	
-	\sigma(|S|) = L \hspace{3mm}
-    \sigma(S_L) = B \hspace{3mm}
-    <\texttt{PUSH } |B|, \sigma> \Downarrow \sigma'
-}
-{	
-	<\texttt{OP\_SIZE}, \sigma > \Downarrow \sigma'
-}
-\vspace{3mm}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Bitwise Logic}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% OP_INVERT
-\subsubsection{OP\_INVERT}
-\textcolor{red}{\textit{Disabled}}\\
-Flips all bits of the top stack element. \\
-
-\inferrule
-{	
-	\sigma(|S|) = L \hspace{3mm}
-    \sigma(S_L) = b_1 \ldots b_k \hspace{3mm}
-    <\texttt{PUSH } \lnot b_1 \lnot b_2 \ldots \lnot b_k, \sigma> \Downarrow \sigma_1 \hspace{3mm}
-    <\texttt{DROP } L, \sigma_1> \Downarrow \sigma'
-}
-{	
-	<\texttt{OP\_EQUAL}, \sigma > \Downarrow \sigma'
-}
-\vspace{3mm}
-
-% OP_AND
-\subsubsection{OP\_AND}
-\textcolor{red}{\textit{Disabled}}\\
-Boolean \textit{and} between bits of first and second stack elements. \\
-
-\inferrule
-{	
-	\sigma(|S|) = L \hspace{3mm}
-    \sigma(S_L) = B_1 \hspace{3mm}
-    \sigma(S_{L-1}) = B_2 \hspace{3mm} \\
-    <\texttt{PUSH } B_1 \mathrel{\&} B_2, \sigma> \Downarrow \sigma_1 \hspace{3mm}
-    <\texttt{DROP } L, \sigma_1> \Downarrow \sigma_2 \hspace{3mm}
-    <\texttt{DROP } L-1, \sigma_2> \Downarrow \sigma'
-}
-{	
-	<\texttt{OP\_AND}, \sigma > \Downarrow \sigma'
-}
-\vspace{3mm}
-
-% OP_OR
-\subsubsection{OP\_OR}
-\textcolor{red}{\textit{Disabled}}\\
-Boolean \textit{and} between bits of first and second stack elements. \\
-
-\inferrule
-{	
-	\sigma(|S|) = L \hspace{3mm}
-    \sigma(S_L) = B_1 \hspace{3mm}
-    \sigma(S_{L-1}) = B_2 \hspace{3mm} \\
-    <\texttt{PUSH } B_1 \mathrel{|} B_2, \sigma> \Downarrow \sigma_1 \hspace{3mm}
-    <\texttt{DROP } L, \sigma_1> \Downarrow \sigma_2 \hspace{3mm}
-    <\texttt{DROP } L-1, \sigma_2> \Downarrow \sigma'
-}
-{	
-	<\texttt{OP\_OR}, \sigma > \Downarrow \sigma'
-}
-\vspace{3mm}
-
-% OP_XOR
-\subsubsection{OP\_XOR}
-\textcolor{red}{\textit{Disabled}}\\
-Boolean \textit{and} between bits of first and second stack elements. \\
-
-\inferrule
-{	
-	\sigma(|S|) = L \hspace{3mm}
-    \sigma(S_L) = B_1 \hspace{3mm}
-    \sigma(S_{L-1}) = B_2 \hspace{3mm} \\
-    <\texttt{PUSH } B_1 \hspace{1mm} ^\wedge \hspace{1mm} B_2, \sigma> \Downarrow \sigma_1 \hspace{3mm}
-    <\texttt{DROP } L, \sigma_1> \Downarrow \sigma_2 \hspace{3mm}
-    <\texttt{DROP } L-1, \sigma_2> \Downarrow \sigma'
-}
-{	
-	<\texttt{OP\_XOR}, \sigma > \Downarrow \sigma'
-}
-\vspace{3mm}
-
-% OP_EQUAL
-\subsubsection{OP\_EQUAL}
-Returns 1 if the two top elements are equal, 0 otherwise. \\
-
-\inferrule
-{	
-	\sigma(|S|) = L \hspace{3mm}
-    \sigma(S_L) = B_1 \hspace{3mm}
-    \sigma(S_{L-1}) = B_2 \hspace{3mm}
-    B_1 = B_2 \hspace{3mm}
-    <\texttt{PUSH } 1, \sigma> \Downarrow \sigma'
-}
-{	
-	<\texttt{OP\_EQUAL}, \sigma > \Downarrow \sigma'
-}
-\vspace{3mm}
-
-\inferrule
-{	
-	\sigma(|S|) = L \hspace{3mm}
-    \sigma(S_L) = B_1 \hspace{3mm}
-    \sigma(S_{L-1}) = B_2 \hspace{3mm}
-    B_1 \neq B_2 \hspace{3mm}
-    <\texttt{PUSH } 0, \sigma> \Downarrow \sigma'
-}
-{	
-	<\texttt{OP\_EQUAL}, \sigma > \Downarrow \sigma'
-}
-\vspace{3mm}
-
-% OP_EQUALVERIFY
-\subsubsection{OP\_EQUALVERIFY}
-Description. \\
-
-\inferrule
-{	
-	<\texttt{OP\_EQUAL }, \sigma> \Downarrow \sigma_1
-    <\texttt{OP\_VERIFY }, \sigma_1> \Downarrow \sigma'
-}
-{	
-	<\texttt{OP\_EQUALVERIFY}, \sigma > \Downarrow \sigma'
-}
-\vspace{3mm}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Arithmetic}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% OP_1ADD
-\subsubsection{OP\_1ADD}
-Add 1 to the top of the stack. \\
-
-\inferrule
-{	
-	\sigma(|S|) = l \hspace{3mm}
-    \sigma(S_l) \Downarrow x \hspace{3mm}
-    B \Downarrow x+1 \hspace{3mm}
-	<\texttt{DROP } l, \sigma> \Downarrow \sigma_1 \hspace{3mm}
-    <\texttt{PUSH } B, \sigma_1> \Downarrow \sigma'
-}
-{	
-	<\texttt{OP\_2SWAP}, \sigma > \Downarrow \sigma'
-}
-\vspace{3mm}
-
-% OP_1SUB
-\subsubsection{OP\_1SUB}
-Subtract 1 from the top of the stack. \\
-
-\inferrule
-{	
-	\sigma(|S|) = l \hspace{3mm}
-    \sigma(S_l) \Downarrow x \hspace{3mm}
-    B \Downarrow x-1 \hspace{3mm}
-	<\texttt{DROP } l, \sigma> \Downarrow \sigma_1 \hspace{3mm}
-    <\texttt{PUSH } B, \sigma_1> \Downarrow \sigma'
-}
-{	
-	<\texttt{OP\_1SUB}, \sigma> \Downarrow \sigma'
-}
-\vspace{3mm}
-
-% OP_ABS
-\subsubsection{OP\_ABS}
-The top of the stack is replaced with the byte-vector representing it's absolute value. \\
-
-\inferrule
-{	
-	\sigma(|S|) = l \hspace{3mm}
-    \sigma(S_l) \Downarrow x \hspace{3mm}
-    B \Downarrow |x| \hspace{3mm}
-	<\texttt{DROP } l, \sigma> \Downarrow \sigma_1 \hspace{3mm}
-    <\texttt{PUSH } B, \sigma_1> \Downarrow \sigma'
-}
-{	
-	<\texttt{OP\_1SUB}, \sigma> \Downarrow \sigma'
-}
-\vspace{3mm}
-
-% OP_NEGATE
-\subsubsection{OP\_NEGATE}
-The top of the stack is negated. \\
-
-\inferrule
-{	
-	\sigma(|S|) = l \hspace{3mm}
-    \sigma(S_l) \Downarrow x \hspace{3mm}
-    B \Downarrow -x \hspace{3mm}
-	<\texttt{DROP } l, \sigma> \Downarrow \sigma_1 \hspace{3mm}
-    <\texttt{PUSH } B, \sigma_1> \Downarrow \sigma'
-}
-{	
-	<\texttt{OP\_1SUB}, \sigma> \Downarrow \sigma'
-}
-\vspace{3mm}
-
-% OP_NOT
-\subsubsection{OP\_NOT}
-If the top of the stack is 0, replace it with the byte vector representing 1. Otherwise, replace it with the byte-vector representing 0. \\
-
-\inferrule
-{	
-    <\texttt{OP\_IF PUSH OP\_0 OP\_ELSE PUSH OP\_1 OP\_ENDIF}, \sigma> \Downarrow \sigma'
-}
-{	
-	<\texttt{OP\_NOT}, \sigma> \Downarrow \sigma'
-}
-\vspace{3mm}
-
-% OP_0NOTEQUAL
-\subsubsection{OP\_0NOTEQUAL}
-Push the byte-vector representing 0 onto the stack if the top of stack is 0. Otherwise replace the top of stack with the byte-vector representing 1. \\
-
-\inferrule
-{	
-    <\texttt{OP\_IF PUSH OP\_1 OP\_ELSE PUSH OP\_0 OP\_ENDIF}, \sigma> \Downarrow \sigma'
-}
-{	
-	<\texttt{OP\_0NOTEQUAL}, \sigma> \Downarrow \sigma'
-}
-\vspace{3mm}
-
-% OP_ADD
-\subsubsection{OP\_ADD}
-Let the inputs be the top two items of the stack, $a$ and $b$. Push the byte-vector representing $a+b$ onto the stack. \\
-
-\inferrule
-{	
-	\sigma(|S|) = l \hspace{3mm}
-    \sigma(S_l) \Downarrow x \hspace{3mm}
-    \sigma(S_{l-1}) \Downarrow y \hspace{3mm}
-    B \Downarrow x+y \hspace{3mm}
-    <\texttt{OP\_2DROP}, \sigma> \Downarrow \sigma_1 \hspace{3mm}
-    <\texttt{PUSH B}, \sigma_1> \Downarrow \sigma' \hspace{3mm}
-}
-{	
-	<\texttt{OP\_ADD}, \sigma> \Downarrow \sigma'
-}
-\vspace{3mm}
-
-% OP_SUB
-\subsubsection{OP\_SUB}
-Let the inputs be the top two items of the stack, $a$ and $b$. Push the byte-vector representing $a-b$ onto the stack. \\
-
-\inferrule
-{	
-	\sigma(|S|) = l \hspace{3mm}
-    \sigma(S_l) \Downarrow x \hspace{3mm}
-    \sigma(S_{l-1}) \Downarrow y \hspace{3mm}
-    B \Downarrow y-x \hspace{3mm}
-    <\texttt{OP\_2DROP}, \sigma> \Downarrow \sigma_1 \hspace{3mm}
-    <\texttt{PUSH B}, \sigma_1> \Downarrow \sigma' \hspace{3mm}
-}
-{	
-	<\texttt{OP\_SUB}, \sigma> \Downarrow \sigma'
-}
-\vspace{3mm}
-
-% OP_BOOLAND
-\subsubsection{OP\_BOOLAND}
-Let the inputs be the top two items of the stack, $a$ and $b$. If both $a$ and $b$ are not 0, then push 1 onto the stack. \\
-
-\inferrule
-{	
-	\sigma(|S|) = L \hspace{3mm}
-    \sigma(S_L) \Downarrow a \hspace{3mm}
-    \sigma(S_{L-1}) \Downarrow b \hspace{3mm}
-    a \neq 0 \land b \neq 0 \hspace{3mm}
-    <\texttt{OP\_2DROP}, \sigma> \Downarrow \sigma_1 \hspace{3mm}
-    <\texttt{PUSH OP\_1}, \sigma_1> \Downarrow \sigma' \hspace{3mm}
-}
-{	
-	<\texttt{OP\_BOOLAND}, \sigma> \Downarrow \sigma'
-}
-\vspace{3mm}
-
-\inferrule
-{	
-	\sigma(|S|) = L \hspace{3mm}
-    \sigma(S_L) \Downarrow a \hspace{3mm}
-    \sigma(S_{L-1}) \Downarrow b \hspace{3mm}
-    a = 0 \lor b = 0 \hspace{3mm}
-    <\texttt{OP\_2DROP}, \sigma> \Downarrow \sigma_1 \hspace{3mm}
-    <\texttt{PUSH OP\_0}, \sigma> \Downarrow \sigma' \hspace{3mm}
-}
-{	
-	<\texttt{OP\_BOOLAND}, \sigma> \Downarrow \sigma'
-}
-\vspace{3mm}
-
-% OP_BOOLOR
-\subsubsection{OP\_BOOLOR}
-Let the inputs be the top two items of the stack, $a$ and $b$. If either $a$ or $b$ are not 0, then push 1 onto the stack. \\
-
-\inferrule
-{	
-	\sigma(|S|) = L \hspace{3mm}
-    \sigma(S_L) \Downarrow a \hspace{3mm}
-    \sigma(S_{L-1}) \Downarrow b \hspace{3mm}
-    a \neq 0 \lor b \neq 0 \hspace{3mm}
-    <\texttt{OP\_2DROP}, \sigma> \Downarrow \sigma_1 \hspace{3mm}
-    <\texttt{OP\_1}, \sigma_1> \Downarrow \sigma' \hspace{3mm}
-}
-{	
-	<\texttt{OP\_BOOLOR}, \sigma> \Downarrow \sigma'
-}
-\vspace{3mm}
-
-\inferrule
-{	
-	\sigma(|S|) = L \hspace{3mm}
-    \sigma(S_L) \Downarrow a \hspace{3mm}
-    \sigma(S_{L-1}) \Downarrow b \hspace{3mm}
-    a \neq 0 \land b \neq 0 \hspace{3mm}
-    <\texttt{OP\_2DROP}, \sigma> \Downarrow \sigma_1 \hspace{3mm}
-    <\texttt{OP\_0}, \sigma_1> \Downarrow \sigma' \hspace{3mm}
-}
-{	
-	<\texttt{OP\_BOOLOR}, \sigma> \Downarrow \sigma'
-}
-\vspace{3mm}
-
-% OP_NUMEQUAL
-\subsubsection{OP\_NUMEQUAL}
-Let the inputs be the top two items of the stack, $a$ and $b$. Push 1 onto the stack if $a = b$; push 0 onto the stack otherwise. \\
-
-\inferrule
-{	
-    <\texttt{OP\_SUB}, \sigma> \Downarrow \sigma_1 \hspace{3mm}
-    <\texttt{OP\_IF PUSH OP\_0 OP\_ELSE PUSH OP\_1}, \sigma_1> \Downarrow \sigma' \hspace{3mm}
-}
-{	
-	<\texttt{OP\_NUMEQUAL}, \sigma> \Downarrow \sigma'
-}
-\vspace{3mm}
-
-% OP_NUMEQUALVERIFY
-\subsubsection{OP\_NUMEQUALVERIFY}
-Let the inputs be the top two items of the stack, $a$ and $b$. Push 1 onto the stack if $a = b$; push 0 onto the stack otherwise. If 0 is pushed onto the stack, mark the transaction as invalid. \\
-
-\inferrule
-{	
-    <\texttt{OP\_NUMEQUAL}, \sigma> \Downarrow \sigma_1 \hspace{3mm}
-    <\texttt{OP\_VERIFY}, \sigma_1> \Downarrow \sigma' \hspace{3mm}
-}
-{	
-	<\texttt{OP\_NUMEQUALVERIFY}, \sigma> \Downarrow \sigma'
-}
-\vspace{3mm}
-
-% OP_NUMNOTEQUAL
-\subsubsection{OP\_NUMNOTEQUAL}
-Let the inputs be the top two items of the stack, $a$ and $b$. Push 1 if the numbers are equal, 0 otherwise. \\
-
-\inferrule
-{	
-    <\texttt{OP\_SUB}, \sigma> \Downarrow \sigma_1 \hspace{3mm}
-    <\texttt{OP\_IF PUSH OP\_1 OP\_ELSE PUSH OP\_0}, \sigma_1> \Downarrow \sigma' \hspace{3mm}
-}
-{	
-	<\texttt{OP\_NUMNOTEQUAL}, \sigma> \Downarrow \sigma'
-}
-\vspace{3mm}
-
-% OP_LESSTHAN
-\subsubsection{OP\_LESSTHAN}
-Let the inputs be the element on the top of the stack(b) and the element second to the top (a).  Push 1 onto the stack if a is less than b, otherwise push 0. \\
-
-\inferrule
-{	\sigma(|S|) = L \hspace{3mm}
-	\sigma(S_{L-1}) \Downarrow a \hspace{3mm}
-	\sigma(S_L) \Downarrow b \hspace{3mm}
-	a \geq b \hspace{3mm}
-	<\texttt{OP\_2DROP}, \sigma> \Downarrow \sigma_1 \hspace{3mm}
-	<\texttt{OP\_0}, \sigma_1> \Downarrow \sigma ' \hspace{3mm}
-}
-{	<\texttt{OP\_LESSTHAN}, \sigma> \Downarrow \sigma '
-}
-\vspace{3mm}
-
-\inferrule
-{	\sigma(|S|) = L \hspace{3mm}
-	\sigma(S_{L-1}) \Downarrow a \hspace{3mm}
-	\sigma(S_L) \Downarrow b \hspace{3mm}
-	a < b \hspace{3mm}
-	<\texttt{OP\_2DROP}, \sigma> \Downarrow \sigma_1 \hspace{3mm}
-	<\texttt{OP\_1}, \sigma_1> \Downarrow \sigma ' \hspace{3mm}
-}
-{	<\texttt{OP\_LESSTHAN}, \sigma> \Downarrow \sigma '
-}
-\vspace{3mm}
-
-% OP_GREATERTHAN
-\subsubsection{OP\_GREATERTHAN}
-Let the inputs be the element on the top of the stack(b) and the element second to the top (a).  Push 1 onto the stack if a is greater than b, otherwise push 0. \\
-
-\inferrule
-{	\sigma(|S|) = L \hspace{3mm}
-	\sigma(S_{L-1}) \Downarrow a \hspace{3mm}
-	\sigma(S_L) \Downarrow b \hspace{3mm}
-	a \leq b \hspace{3mm}
-	<\texttt{OP\_2DROP}, \sigma> \Downarrow \sigma_1 \hspace{3mm}
-	<\texttt{OP\_0}, \sigma_1> \Downarrow \sigma ' \hspace{3mm}
-}
-{	<\texttt{OP\_GREATERTHAN}, \sigma> \Downarrow \sigma '
-}
-\vspace{3mm}
-
-\inferrule
-{	\sigma(|S|) = L \hspace{3mm}
-	\sigma(S_{L-1}) \Downarrow a \hspace{3mm}
-	\sigma(S_L) \Downarrow b \hspace{3mm}
-	a > b \hspace{3mm}
-	<\texttt{OP\_2DROP}, \sigma> \Downarrow \sigma_1 \hspace{3mm}
-	<\texttt{OP\_1}, \sigma_1> \Downarrow \sigma ' \hspace{3mm}
-}
-{	<\texttt{OP\_GREATERTHAN}, \sigma> \Downarrow \sigma '
-}
-\vspace{3mm}
-
-% OP_LESSTHANOREQUAL
-\subsubsection{OP\_LESSTHANOREQUAL}
-Let the inputs be the element on the top of the stack(b) and the element second to the top (a).  Push 1 onto the stack if a is less than or equal to b, otherwise push 0. \\
-
-\inferrule
-{	\sigma(|S|) = L \hspace{3mm}
-	\sigma(S_{L-1}) \Downarrow a \hspace{3mm}
-	\sigma(S_L) \Downarrow b \hspace{3mm}
-	a > b \hspace{3mm}
-	<\texttt{OP\_2DROP}, \sigma> \Downarrow \sigma_1 \hspace{3mm}
-	<\texttt{OP\_0}, \sigma_1> \Downarrow \sigma ' \hspace{3mm}
-}
-{	<\texttt{OP\_LESSTHANOREQUAL}, \sigma> \Downarrow \sigma '
-}
-\vspace{3mm}
-
-\inferrule
-{	\sigma(|S|) = L \hspace{3mm}
-	\sigma(S_{L-1}) \Downarrow a \hspace{3mm}
-	\sigma(S_L) \Downarrow b \hspace{3mm}
-	a \leq b \hspace{3mm}
-	<\texttt{OP\_2DROP}, \sigma> \Downarrow \sigma_1 \hspace{3mm}
-	<\texttt{OP\_1}, \sigma_1> \Downarrow \sigma ' \hspace{3mm}
-}
-{	<\texttt{OP\_LESSTHANOREQUAL}, \sigma> \Downarrow \sigma '
-}
-\vspace{3mm}
-
-% OP_GREATERTHANOREQUAL
-\subsubsection{OP\_GREATERTHANOREQUAL}
-Let the inputs be the element on the top of the stack(b) and the element second to the top (a).  Push 1 onto the stack if a is greater than or equal to b, otherwise push 0. \\
-
-\inferrule
-{	\sigma(|S|) = L \hspace{3mm}
-	\sigma(S_{L-1}) \Downarrow a \hspace{3mm}
-	\sigma(S_L) \Downarrow b \hspace{3mm}
-	a < b \hspace{3mm}
-	<\texttt{OP\_2DROP}, \sigma> \Downarrow \sigma_1 \hspace{3mm}
-	<\texttt{OP\_0}, \sigma_1> \Downarrow \sigma ' \hspace{3mm}
-}
-{	<\texttt{OP\_GREATERTHANOREQUAL}, \sigma> \Downarrow \sigma '
-}
-\vspace{3mm}
-
-\inferrule
-{	\sigma(|S|) = L \hspace{3mm}
-	\sigma(S_{L-1}) \Downarrow a \hspace{3mm}
-	\sigma(S_L) \Downarrow b \hspace{3mm}
-	a \geq b \hspace{3mm}
-	<\texttt{OP\_2DROP}, \sigma> \Downarrow \sigma_1 \hspace{3mm}
-	<\texttt{OP\_1}, \sigma_1> \Downarrow \sigma ' \hspace{3mm}
-}
-{	<\texttt{OP\_GREATERTHANOREQUAL}, \sigma> \Downarrow \sigma '
-}
-\vspace{3mm}
-
-% OP_MIN
-\subsubsection{OP\_MIN}
-the larger of the two elements at the top of the stack is removed. \\
-
-\inferrule
-{
-    \sigma(|S|) = L \hspace{3mm})
-    \sigma(S_{L-1}) \Downarrow a \hspace{3mm}
-    \sigma(S_{L}) \Downarrow b \hspace{3mm}
-    a \leq b \hspace{3mm}
-    <\texttt{DROP }L, \sigma> \Downarrow \sigma' \hspace{3mm}
-}
-{
-    <\texttt{OP\_MIN}, \sigma> \Downarrow \sigma'
-}
-\vspace{3mm}
-
-\inferrule
-{
-    \sigma(|S|) = L \hspace{3mm}
-    \sigma(S_{L-1}) \Downarrow a \hspace{3mm}
-    \sigma(S_{L}) \Downarrow b \hspace{3mm}
-    b < a \hspace{3mm}
-    <\texttt{DROP } L-1, \sigma> \Downarrow \sigma' \hspace{3mm}
-}
-{
-    <\texttt{OP\_MIN}, \sigma> \Downarrow \sigma'
-}
-\vspace{3mm}
-
-% OP_MAX
-\subsubsection{OP\_MAX}
-the smaller of the two elements on the top of the stack is removed. \\
-
-\inferrule
-{
-    \sigma(|S|) = L \hspace{3mm}
-    \sigma(S_{L-1}) \Downarrow a \hspace{3mm}
-    \sigma(S_{L}) \Downarrow b \hspace{3mm}
-    a \geq b \hspace{3mm}
-    <\texttt{DROP }L, \sigma> \Downarrow \sigma' \hspace{3mm}
-}
-{
-    <\texttt{OP\_MAX}, \sigma> \Downarrow \sigma'
-}
-\vspace{3mm}
-
-\inferrule
-{
-    \sigma(|S|) = L \hspace{3mm}
-    \sigma(S_{L-1}) \Downarrow a \hspace{3mm}
-    \sigma(S_{L}) \Downarrow b \hspace{3mm}
-    b > a \hspace{3mm}
-    <\texttt{DROP } L-1, \sigma> \Downarrow \sigma' \hspace{3mm}
-}
-{
-    <\texttt{OP\_MAX}, \sigma> \Downarrow \sigma'
-}
-\vspace{3mm}
-
-% OP_WITHIN
-\subsubsection{OP\_WITHIN}
-the input is the 3 elements at the top of the stack x, min, max where max is the top of the stack.  push 1 if x is greater than or equal to min and less than max, otherwise push 0. \\
-
-\inferrule
-{
-    \sigma(|S|) = L \hspace{3mm}
-    \sigma(S_{L}) \Downarrow max \hspace{3mm}
-    \sigma(S_{L-1}) \Downarrow min \hspace{3mm}
-    \sigma(S_{L-2}) \Downarrow x \hspace{3mm}
-    min \leq x < max \hspace{3mm}
-    <\texttt{OP\_1}, \sigma> \Downarrow \sigma' \hspace{3mm}
-}
-{
-    <\texttt{OP\_WITHIN}, \sigma> \Downarrow \sigma'
-}
-\vspace{3mm}
-
-\inferrule
-{
-    \sigma(|S|) = L \hspace{3mm}
-    \sigma(S_{L}) \Downarrow max \hspace{3mm}
-    \sigma(S_{L-1}) \Downarrow min \hspace{3mm}
-    \sigma(S_{L-2}) \Downarrow x \hspace{3mm}
-    \lnot (min \leq x < max) \hspace{3mm}
-    <\texttt{OP\_1}, \sigma> \Downarrow \sigma' \hspace{3mm}
-}
-{
-    <\texttt{OP\_WITHIN}, \sigma> \Downarrow \sigma'
-}
-\vspace{3mm}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Crypto}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsubsection{OP\_RIPEMD160}
-The element at the top of the stack is hashed with RIPEMD-160. \\
-
-\inferrule
-{
-	\sigma(|S|) = L \hspace{3mm}
-    B = \textsc{RIPEMD-160}(\sigma(S_L)) \hspace{3mm}
-    <\texttt{DROP }L, \sigma> \Downarrow sigma_1 \hspace{3mm}
-    <\texttt{PUSH }B, \sigma_1> \Downarrow sigma' \hspace{3mm}
-}
-{
-	<\texttt{OP\_RIPEMD-160}, \sigma> \Downarrow sigma'
-}
-\vspace{3mm}
-
-\subsubsection{OP\_SHA1}
-The element at the top of the stack is hashed with SHA-1. \\
-
-\inferrule
-{
-	\sigma(|S|) = L \hspace{3mm}
-    B = \textsc{SHA-1}(\sigma(S_L)) \hspace{3mm}
-    <\texttt{DROP }L, \sigma> \Downarrow sigma_1 \hspace{3mm}
-    <\texttt{PUSH }B, \sigma_1> \Downarrow sigma' \hspace{3mm}
-}
-{
-	<\texttt{OP\_SHA1}, \sigma> \Downarrow sigma'
-}
-\vspace{3mm}
-\subsubsection{OP\_SHA256}
-The element at the top of the stack is hashed with SHA-256. \\
-
-\inferrule
-{
-	\sigma(|S|) = L \hspace{3mm}
-    B = \textsc{SHA-256}(\sigma(S_L)) \hspace{3mm}
-    <\texttt{DROP }L, \sigma> \Downarrow sigma_1 \hspace{3mm}
-    <\texttt{PUSH }B, \sigma_1> \Downarrow sigma' \hspace{3mm}
-}
-{
-	<\texttt{OP\_SHA-256}, \sigma> \Downarrow sigma'
-}
-\vspace{3mm}
-\subsubsection{OP\_HASH160}
-The element at the top of the stack is hashed with SHA-256 and then with RIPEMD-160. \\
-
-\inferrule
-{
-	\sigma(|S|) = L \hspace{3mm}
-    B = \textsc{RIPEMD-160}(\textsc{SHA-256}(\sigma(S_L))) \hspace{3mm}
-    <\texttt{DROP }L, \sigma> \Downarrow sigma_1 \hspace{3mm}
-    <\texttt{PUSH }B, \sigma_1> \Downarrow sigma' \hspace{3mm}
-}
-{
-	<\texttt{OP\_HASH160}, \sigma> \Downarrow sigma'
-}
-\vspace{3mm}
-
-\subsubsection{OP\_HASH256}
-The element at the top of the stack is hashed twice with SHA-256. \\
-
-\inferrule
-{
-	\sigma(|S|) = L \hspace{3mm}
-    B = \textsc{SHA-256}(\textsc{SHA-256}(\sigma(S_L))) \hspace{3mm}
-    <\texttt{DROP }L, \sigma> \Downarrow sigma_1 \hspace{3mm}
-    <\texttt{PUSH }B, \sigma_1> \Downarrow sigma' \hspace{3mm}
-}
-{
-	<\texttt{OP\_HASH256}, \sigma> \Downarrow sigma'
-}
-\vspace{3mm}
-
-\subsubsection{OP\_CODESEPARATOR}
-\subsubsection{OP\_CHECKSIG}
-\subsubsection{OP\_CHECKSIGVERIFY}
-\subsubsection{OP\_CHECKMULTISIG}
-\subsubsection{OP\_CHECKMULTISIGVERIFY}
-\subsection{Locktime}
-\subsubsection{OP\_CHECKLOCKTIMEVERIFY (previously OP\_NOP2)}
-\subsubsection{OP\_CHECKSEQUENCEVERIFY (previously OP\_NOP3)}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Disabled / Reserved / Invalid}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-All disabled, reserved or otherwise invalid opcodes transition to an invalid state.
-
-\inferrule
-{\quad}
-{<\texttt{dcom},\sigma> \Downarrow \sigma[V=invalid]}
-
-\section{Operational Semantics -- Small Step Semantics}
-
-\subsection{Local reduction rules}
-
-[Have to explain what are atomic commands, and why we're defining them.]
-
-[No need to add PUSHDATA commands: these can be described as atomic commands, if we remember that the PUSHDATA opcode is only part of the push commands defined in the syntax - Hugo]
-
-\begin{align*}
-&<\texttt{OP\_NOP com}, \sigma> 
-	\hspace{3mm}
-\longrightarrow 
-	\hspace{3mm}
-<\texttt{com}, \sigma>
-\\
-&<\texttt{atom\_com}, \sigma[S=s, V=v]> 
-	\hspace{3mm}
-\longrightarrow 
-	\hspace{3mm}
-<\texttt{OP\_NOP}, \sigma[S=s', V=v']>
-\\
-&<\texttt{OP\_IF com OP\_ENDIF}, \sigma[S_{|S|} \neq 0]> 
-	\hspace{3mm}
-\longrightarrow 
-	\hspace{3mm}
-<\texttt{com}, \sigma> 
-\\
-&<\texttt{OP\_IF com OP\_ENDIF}, \sigma[S_{|S|} = 0]> 
-	\hspace{3mm}
-\longrightarrow 
-	\hspace{3mm}
-<\texttt{OP\_NOP}, \sigma> 
-\\
-&<\texttt{OP\_NOTIF com OP\_ENDIF}, \sigma[S_{|S|} \neq 0]> 
-	\hspace{3mm}
-\longrightarrow 
-	\hspace{3mm}
-<\texttt{OP\_NOP}, \sigma> 
-\\
-&<\texttt{OP\_NOTIF com OP\_ENDIF}, \sigma[S_{|S|} = 0]> 
-	\hspace{3mm}
-\longrightarrow 
-	\hspace{3mm}
-<\texttt{com}, \sigma> 
-\\
-&<\texttt{OP\_IF com$_1$ OP\_ELSE com$_2$ OP\_ENDIF}, \sigma[S_{|S|} \neq 0]> 
-	\hspace{3mm}
-\longrightarrow 
-	\hspace{3mm}
-<\texttt{com$_1$}, \sigma> 
-\\
-&<\texttt{OP\_IF com$_1$ OP\_ELSE com$_2$ OP\_ENDIF}, \sigma[S_{|S|} = 0]> 
-	\hspace{3mm}
-\longrightarrow 
-	\hspace{3mm}
-<\texttt{com$_2$}, \sigma> 
-\\
-&<\texttt{OP\_NOTIF com$_1$ OP\_ELSE com$_2$ OP\_ENDIF}, \sigma[S_{|S|} \neq 0]> 
-	\hspace{3mm}
-\longrightarrow 
-	\hspace{3mm}
-<\texttt{com$_2$}, \sigma> 
-\\
-&<\texttt{OP\_NOTIF com$_1$ OP\_ELSE com$_2$ OP\_ENDIF}, \sigma[S_{|S|} = 0]> 
-	\hspace{3mm}
-\longrightarrow 
-	\hspace{3mm}
-<\texttt{com$_1$}, \sigma> 
-\\
-\end{align*}
-
-\subsection{Global reduction rules}
-
-Since programs in Script are evaluated strictly from left to right, there is really only global reduction rule aside from $\bullet$.
-
-$$H ::= \bullet \mid H \hspace{1mm} \texttt{com}$$
-
 \end{document}
